---
title: "CONUS: Delineate Watersheds"
author: "Katie Willi"
date: "2026-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(mapview)
library(nngeo)
library(nhdplusTools)
library(sf)

# Sources getXYWatersheds(), add_new_locations(), and review helpers.
# getXYWatersheds() no longer contains any hardcoded site coordinates — all
# per-site delineation settings live in data/site_config.csv.
list.files("src/", full.names = TRUE) %>%
  walk(~source(.))
```

---

## One-time setup

Run `src/bootstrap_site_config.R` once (interactively or via Rscript) to create
`data/site_config.csv`. After that, edit the CSV directly to adjust settings for
specific sites. The bootstrap script only needs to be re-run if the site list
changes substantially enough to warrant reassigning location IDs from scratch —
for incremental additions, use `add_new_locations()` below.

---

## Load site config

`site_config.csv` is the single source of truth for per-location delineation
settings. It controls snap method, buffer radius, COMID corrections, and more.
Open it in Excel or any spreadsheet app to inspect or edit.

```{r load_config}
site_config <- read_csv("data/site_config.csv")

cat("Site config loaded:", nrow(site_config), "unique CONUS locations\n")
cat("  On-network (snap=TRUE): ",  sum( site_config$snap), "\n")
cat("  Off-network (snap=FALSE):", sum(!site_config$snap), "\n")
cat("  COMID overrides:         ", sum(!is.na(site_config$comid_override)), "\n")
cat("  Flagged international:   ", sum(site_config$is_international), "\n")
```

---

## Review GROW

Compare the current site list against the previous published GROWdb version.

```{r load_sites}
# NOTE: the Marshall Gulch coordinate correction is a known data quality issue
# in the source CSV. The corrected coordinates are preserved here until the
# source file is fixed upstream.
conus_sites_update <- read_csv("data/date_lat_long_grow_KKA_05.05.26.csv", show_col_types = FALSE) %>%
  mutate(
    Longitude = if_else(sample_name == "marshallgulch_2019_sw_WHONDRS-S19S_0035", -110.75784, Longitude),
    Latitude  = if_else(sample_name == "marshallgulch_2019_sw_WHONDRS-S19S_0035",   32.42735, Latitude)
  ) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  .[tigris::states() %>% filter(!STUSPS %in% c("HI", "PR", "AK")) %>% st_buffer(10000), ]
```

```{r find_new_dropped}
sites_previous <- read_csv("data/archive/GROWdb_with_vars_20230519.csv", show_col_types = FALSE) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  .[tigris::states() %>% filter(!STUSPS %in% c("HI", "PR", "AK")) %>% st_buffer(10000), ] %>%
  st_drop_geometry()

# Sites that are NEW in the latest update
new_update <- conus_sites_update %>%
  anti_join(sites_previous, by = c("sample_name" = "SampleName"))

# Sites whose coordinates changed between new and previous version
new_previous_coords_dif <- conus_sites_update %>%
  filter(sample_name %in% sites_previous$SampleName) %>%
  anti_join(sites_previous, by = c("Latitude", "Longitude"))

old_points <- read_csv("data/archive/GROWdb_with_vars_20230519.csv", show_col_types = FALSE) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  filter(SampleName %in% new_previous_coords_dif$sample_name) %>%
  select(SampleName, Longitude, Latitude)

# new_previous_coords_dif_reviewed <- review_point_pairs(new_points = new_previous_coords_dif,
#                                                        old_points = old_points)
# ... Marshall Gulch changed substantially; previous coordinates are more appropriate. Updated.

# Sites that were DROPPED from the previous GROW version
missing_previous <- sites_previous %>%
  anti_join(conus_sites_update, by = c("SampleName" = "sample_name"))
```

---

## Delineate all GROW site watersheds

```{r check_new_locations}
# Check whether any sites in the current update are missing from site_config.csv.
# New sites need a config row (with location_id and delineation settings) before
# delineation can run. If any appear here, call add_new_locations() below.
missing_from_config <- conus_sites_update %>%
  st_drop_geometry() %>%
  distinct(Latitude, Longitude) %>%
  anti_join(site_config, by = c("Latitude", "Longitude")) %>%
  # Exclude sites that are simply outside the CONUS scope
  filter(!is.na(Latitude))

if (nrow(missing_from_config) > 0) {
  cat("WARNING:", nrow(missing_from_config),
      "site location(s) are not in site_config.csv:\n")
  print(missing_from_config)
  cat("\nRun add_new_locations(missing_from_config) to register them with",
      "default settings,\nthen review site_config.csv before delineating.\n")
} else {
  cat("All site locations are registered in site_config.csv.\n")
}
```

```{r add_new_locations, eval=FALSE}
# Run this chunk if missing_from_config above is non-empty.
# It appends the new locations to site_config.csv with default settings.
# Open the CSV afterward and adjust any parameters for unusual sites.
site_config <- add_new_locations(missing_from_config)
```

```{r prep_delineation}
# International sites pass the CONUS boundary filter but are on international
# waters. They are excluded here and handled by international_workflow.Rmd.
# The is_international flag in site_config.csv controls this — no sample names
# are hardcoded in this script.
international_locs <- site_config %>%
  filter(is_international) %>%
  select(Latitude, Longitude)

# Collapse to unique locations for delineation (multiple samples at the same
# coordinates only need one watershed polygon). Join config columns so each
# row carries its own delineation settings into getXYWatersheds().
new_dupes_removed <- conus_sites_update %>%
  distinct(Longitude, Latitude, .keep_all = TRUE) %>%
  anti_join(international_locs, by = c("Latitude", "Longitude")) %>%
  left_join(site_config, by = c("Latitude", "Longitude")) %>%
  st_as_sf()

cat("Unique CONUS locations to delineate:", nrow(new_dupes_removed), "\n")
```

```{r delineate, eval=FALSE}
# Delineate each unique location. getXYWatersheds() reads its per-site settings
# (snap, buffer, COMID override, etc.) from the site_config columns on each row.
new_dupes_removed %>%
  split(seq_len(nrow(.))) %>%
  walk(function(site_row) {
    cfg <- site_row %>%
      st_drop_geometry() %>%
      select(any_of(names(site_config)))
    try(getXYWatersheds(sf        = site_row,
                        site_config = cfg,
                        file_path   = "data/grow_watersheds_all/"))
  })
```

Load all delineated watersheds:

```{r load_watersheds}
watersheds <- list.files("data/grow_watersheds_all/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x))
```

Retry any sites that errored out. A failed site produces no output file, so it
will not have a `location_id` in the loaded watersheds:

```{r retry_failed, eval=FALSE}
new_dupes_removed %>%
  filter(!location_id %in% watersheds$location_id) %>%
  split(seq_len(nrow(.))) %>%
  walk(function(site_row) {
    cfg <- site_row %>%
      st_drop_geometry() %>%
      select(any_of(names(site_config)))
    try(getXYWatersheds(sf          = site_row,
                        site_config = cfg,
                        file_path   = "data/grow_watersheds_all/"))
  })

watersheds <- list.files("data/grow_watersheds_all/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x))

# Any that still failed?
still_broken <- new_dupes_removed %>%
  filter(!location_id %in% watersheds$location_id)
print(still_broken)
```

Review the delineated watersheds — flag any that look wrong:

```{r review, eval=FALSE}
flowlines <- get_nhdplus(comid = watersheds %>% filter(!is.na(comid)) %>% pull(comid))

# # Load per-COMID flowlines individually (batching sometimes drops results)
# comids <- watersheds %>% filter(!is.na(comid)) %>% pull(comid) %>% unique()
# flowlines <- vector("list", length(comids))
# for (i in seq_along(comids)) {
#   flowlines[[i]] <- get_nhdplus(comid = comids[i]) %>% select(comid)
# }
# flowlines <- bind_rows(flowlines)

review <- review_watersheds(watersheds = watersheds, flowlines = flowlines,
                            sites = new_dupes_removed)
saveRDS(review, "data/review_all.RDS")
```

```{r inspect_review, eval=FALSE}
review <- readRDS("data/review_all.RDS")

good  <- review %>% filter(review == "Y")
small <- review %>% filter(review == "S")  # catchment too small for NHD
weird <- review %>% filter(review == "B")  # some other delineation problem
```

# This step is where you would develop the course of action for the problematic watersheds.

## Bind all watersheds

```{r bind_watersheds}
conus_watersheds <- conus_sites_update %>%
  st_drop_geometry() %>%
  inner_join(
    list.files("data/grow_watersheds_all/", full.names = TRUE) %>% map_dfr(~readRDS(.x)),
    by = c("Latitude", "Longitude")
  ) %>%
  mutate(update = if_else(sample_name %in% new_update$sample_name, "new", "previous"))

saveRDS(conus_watersheds, "data/conus_watersheds.RDS")

# Any sites in the full site list that are not in the CONUS watershed set?
# These are the international sites handled by international_workflow.Rmd.
outside_conus <- read_csv("data/date_lat_long_grow_KKA_05.05.26.csv", show_col_types = FALSE) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  filter(!sample_name %in% conus_watersheds$sample_name)

mapview(outside_conus)  # sanity check: should only show non-CONUS sites
```

---

## Add back sites dropped from previous version

```{r prep_missing}
# International locations filter (same logic as above — driven by site_config)
missing_dupes_removed <- missing_previous %>%
  distinct(Longitude, Latitude, .keep_all = TRUE) %>%
  anti_join(international_locs, by = c("Latitude", "Longitude")) %>%
  left_join(site_config, by = c("Latitude", "Longitude")) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  .[tigris::states() %>% filter(!STUSPS %in% c("HI", "PR", "AK")) %>% st_buffer(10000), ]
```

```{r delineate_missing}
missing_dupes_removed %>%
  split(seq_len(nrow(.))) %>%
  walk(function(site_row) {
    cfg <- site_row %>%
      st_drop_geometry() %>%
      select(any_of(names(site_config)))
    try(getXYWatersheds(sf          = site_row,
                        site_config = cfg,
                        file_path   = "data/grow_watersheds_missing/"))
  })

watersheds_missing <- list.files("data/grow_watersheds_missing/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x))

flowlines <- get_nhdplus(comid = watersheds_missing %>% filter(!is.na(comid)) %>% pull(comid))

review <- review_watersheds(watersheds = watersheds_missing, flowlines = flowlines,
                            sites = missing_dupes_removed)
saveRDS(review, "data/review_missing.RDS")
```

```{r bind_complete}
conus_watersheds <- list.files("data/grow_watersheds_all/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) %>%
  inner_join(conus_sites_update %>% st_drop_geometry(), by = c("Latitude", "Longitude")) %>%
  mutate(update = if_else(sample_name %in% new_update$sample_name, "new", "previous")) %>%
  st_make_valid()

conus_missing_watersheds <- list.files("data/grow_watersheds_missing/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) %>%
  inner_join(missing_previous %>% st_drop_geometry(), by = c("Latitude", "Longitude")) %>%
  mutate(update = "missing") %>%
  select(sample_name = SampleName, Collection_Date = Date,
         comid = comid.y, location_id, Latitude, Longitude, watershed, update) %>%
  st_make_valid()

conus_complete <- conus_watersheds %>%
  bind_rows(conus_missing_watersheds) %>%
  distinct() %>%
  st_make_valid()

saveRDS(conus_complete, "data/conus_watersheds.RDS")
```

---

## One-time file migration (run once after creating site_config.csv)

Existing watershed RDS files in `grow_watersheds_all/` were named by the old
session-based `rowid` (e.g., `42.RDS`). The new naming uses stable `location_id`
values from site_config (e.g., `CONUS_0042.RDS`). Run this chunk once to rename
them, then never again.

```{r migrate_file_names, eval=FALSE}
existing_files <- list.files("data/grow_watersheds_all/", full.names = TRUE,
                              pattern = "^[0-9]+\\.RDS$")

renamed <- 0L
skipped <- 0L

for (f in existing_files) {
  ws <- readRDS(f)

  match_row <- site_config %>%
    filter(abs(Latitude - ws$Latitude[1]) < 1e-4 &
           abs(Longitude - ws$Longitude[1]) < 1e-4)

  if (nrow(match_row) != 1) {
    message("No unique config match for ", basename(f), " — skipped")
    skipped <- skipped + 1L
    next
  }

  new_path <- file.path("data/grow_watersheds_all/",
                        paste0(match_row$location_id, ".RDS"))

  if (!file.exists(new_path)) {
    # Add location_id to the stored object before renaming
    ws$location_id <- match_row$location_id
    saveRDS(ws, new_path)
    file.remove(f)
    renamed <- renamed + 1L
  } else {
    skipped <- skipped + 1L
  }
}

cat(sprintf("Renamed: %d  Skipped: %d\n", renamed, skipped))
```
