---
title: "Global: Delineate Watersheds"
output: html_document
date: "2026-06-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = FALSE)
```

```{r libraries}
library(tidyverse)
library(sf)
library(terra)
library(elevatr)
library(lwgeom)
library(nngeo)
library(mapview)
library(leaflet)
library(units)
library(geodata)
library(arrow)
library(exactextractr)

list.files("src", full.names = TRUE) %>%
  walk(~source(.))

sf_use_s2(FALSE)
```

# 1. Setup

```{r paths}
site_csv <- "data/date_lat_long_grow_KKA_05.05.26.csv"
conus_watershed_rds <- "data/conus_watersheds.RDS"

hydrobasins_dir <- "data/hydrobasins"
hydrorivers_dir <- "data/hydrorivers"
```

# 2. Identify international sites

All GROWdb samples not represented in the CONUS watershed object are routed through this international workflow.

```{r identify-international-sites}
all_sites <- readr::read_csv(site_csv, show_col_types = FALSE)
conus_watersheds <- readRDS(conus_watershed_rds)

# If conus_watersheds has sample_name, use that directly. If not, fall back to
# filtering by coordinates already present in the CONUS watershed table.
if ("sample_name" %in% names(conus_watersheds)) {
  intl_sites_all <- all_sites %>%
    filter(!sample_name %in% conus_watersheds$sample_name)
} else {
  conus_coords <- conus_watersheds %>%
    sf::st_drop_geometry() %>%
    distinct(Latitude, Longitude) %>%
    mutate(coord_key = paste(round(Latitude, 6), round(Longitude, 6)))
  
  intl_sites_all <- all_sites %>%
    mutate(coord_key = paste(round(Latitude, 6), round(Longitude, 6))) %>%
    filter(!coord_key %in% conus_coords$coord_key) %>%
    select(-coord_key)
}

# Unique pour-point locations only. All samples sharing the same coordinate are
# joined back after watershed delineation/review.
intl_unique <- intl_sites_all %>%
  distinct(Latitude, Longitude) %>%
  arrange(desc(Latitude), Longitude) %>%
  rowid_to_column("rowid") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE)

cat("International samples:", nrow(intl_sites_all), "\n")
cat("Unique international locations:", nrow(intl_unique), "\n")

mapview(intl_unique, col.regions = "red", cex = 4)
```
# 3. Load HydroBASINS level 12

HydroBASINS is used to get the full upstream watershed for each international pour point.
Level 12 is the finest HydroBASINS level and is the closest global analogue to a local watershed unit.

Download HydroBASINS from HydroSHEDS and unzip all continent-level files needed for the international sites.
Expected shapefile names usually look like `hybas_na_lev12_v1c.shp`, `hybas_sa_lev12_v1c.shp`, etc.

```{r load-hydrobasins}
basin_files <- list.files(hydrobasins_dir,
                          pattern = "hybas.*lev12.*\\.shp$",
                          full.names = TRUE,
                          recursive = TRUE)

#stopifnot(length(basin_files) > 0)

basins <- purrr::map(basin_files, ~sf::st_read(.x, quiet = TRUE)) %>%
  dplyr::bind_rows() %>%
  sf::st_transform(4326) %>%
  sf::st_make_valid()
```

# 4. Load HydroRIVERS

HydroRIVERS is used in two ways:

1. As a review-map overlay.
2. To split the local HydroBASINS polygon at the pour point and remove the downstream area below the sample location.

Download HydroRIVERS from HydroSHEDS and unzip all continent-level files needed for the international sites.
Expected shapefile names usually look like `HydroRIVERS_v10_na.shp`, `HydroRIVERS_v10_sa.shp`, etc.

```{r load-hydrorivers}
hydro_rivers <- st_read("data/hydrorivers/HydroRIVERS_v10.gdb/HydroRIVERS_v10.gdb")
```

Keep only HydroRIVERS segments near the international sites for faster maps and split calls.

```{r trim-hydrorivers}
intl_sites_buf <- intl_unique %>%
  sf::st_transform(3857) %>%
  sf::st_buffer(50000) %>%
  sf::st_transform(4326)

hydro_rivers_review <- hydro_rivers[intl_sites_buf, ]
```

```{r}
flowline_selections <- select_flowline_global(flowlines = hydro_rivers_review,
                                              sites = intl_unique,
                                              river_buffer_m = 20000,
                                              id_col = "rowid")

saveRDS(flowline_selections, "data/flowline_selections.RDS")
```

```{r, eval = FALSE}
delineate_ws(sites = intl_unique %>% filter(!rowid %in% c(to_redo_big$rowid, to_redo$rowid)),
             flowline_selections = flowline_selections,
             flowlines_network   = hydro_rivers,
             basins = basins,
             flow_acc_raster     = terra::rast("data/hyd_glo_acc_15s/hyd_glo_acc_15s.tif"),
             flow_dir_raster     = terra::rast("data/hyd_glo_dir_15s/hyd_glo_dir_15s.tif"),
             flowline_id_col     = "HYRIV_ID",
             downstream_col      = "NEXT_DOWN",
             snap_method         = "point",
             local_basin_pct = 0.80,
             id_col = "rowid",
             hybas_id_co = "HYBAS_ID",
             file_path = "data/intl_watersheds_refined/",
             use_s2 = FALSE)
```

```{r}
sites <- intl_unique %>%
  select(rowid)
```

```{r}
watersheds <- list.files("data/intl_watersheds_refined/", full.names = TRUE) %>%
  purrr::map_dfr(~readRDS(.x) %>%
                   dplyr::mutate(rowid = as.integer(tools::file_path_sans_ext(basename(.x)))) %>%
                   dplyr::select(rowid) %>%
                   nngeo::st_remove_holes())

basins_review <- basins %>%
  .[intl_unique,] %>%
  st_join(., sites, left = FALSE) %>%
  select(HYBAS_ID, rowid)

rivers_review <- hydro_rivers_review %>%
  .[sites %>% 
      sf::st_transform(3857) %>%
      sf::st_buffer(200000) %>%
      sf::st_transform(4326),] %>%
  st_join(., sites %>% 
            sf::st_transform(3857) %>%
            sf::st_buffer(200000) %>%
            sf::st_transform(4326), left = FALSE) %>%
  select(HYRIV_ID, rowid)

reviewed_watersheds <- review_watersheds_global(watersheds = watersheds,
                                                sites = sites,
                                                basins = basins_review,
                                                rivers = rivers_review,
                                                site_buffer_m = 2500,
                                                id_col = "rowid",
                                                review_col = "review")
```

What are the watersheds that weren't delineated correctly?

```{r}
bad <- reviewed_watersheds %>%
  filter(review != "Y")

# what's wrong with them? List out...
review_watersheds_global(watersheds = watersheds %>% filter(rowid %in% bad$rowid),
                         sites = sites %>% filter(rowid %in% bad$rowid),
                         basins = basins_review,
                         rivers = rivers_review,
                         site_buffer_m = 2500,
                         id_col = "rowid",
                         review_col = "review")

```

Fix those watersheds...


```{r}
r1 <- readRDS("data/intl_watersheds_refined/1.RDS")

big_basins <- purrr::map(list.files(hydrobasins_dir,
                          pattern = "hybas.*lev08.*\\.shp$",
                          full.names = TRUE,
                          recursive = TRUE), ~sf::st_read(.x, quiet = TRUE)) %>%
  dplyr::bind_rows() %>%
  sf::st_transform(4326) %>%
  sf::st_make_valid() %>%
  .[r1,] %>%
  .[c(143),] %>%
  select(geometry)

mapview(r1) + big_basins + intl_unique %>% filter(rowid == 1)

final_r1 <- bind_rows(big_basins, r1) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(final_r1) + intl_unique %>% filter(rowid == 1)

saveRDS(final_r1, "data/intl_watersheds_refined/1.RDS")
```


```{r}
# ---------------------------------------------------------------------------
# Rio Grande not quite right... 32
# ---------------------------------------------------------------------------
rio <- readRDS("data/intl_watersheds_refined/32.RDS")

big_basin_files <- list.files(hydrobasins_dir,
                          pattern = "hybas.*lev05.*\\.shp$",
                          full.names = TRUE,
                          recursive = TRUE)

big_basins <- purrr::map(big_basin_files, ~sf::st_read(.x, quiet = TRUE)) %>%
  dplyr::bind_rows() %>%
  sf::st_transform(4326) %>%
  sf::st_make_valid() %>%
  .[rio,] %>%
  .[c(26,24,23),] %>%
  select(geometry)

mapview(rio) + big_basins

final_rio <- bind_rows(big_basins, rio) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(final_rio)

saveRDS(final_rio, "data/intl_watersheds_refined/32.RDS")
```

```{r}
r2 <- readRDS("data/intl_watersheds_refined/2.RDS")

big_basins <- purrr::map(list.files(hydrobasins_dir,
                          pattern = "hybas.*lev08.*\\.shp$",
                          full.names = TRUE,
                          recursive = TRUE), ~sf::st_read(.x, quiet = TRUE)) %>%
  dplyr::bind_rows() %>%
  sf::st_transform(4326) %>%
  sf::st_make_valid() %>%
  .[r2,] %>%
  .[c(66),] %>%
  select(geometry)

mapview(r2) + big_basins + intl_unique %>% filter(rowid == 2)

final_r2 <- bind_rows(big_basins, r2) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(final_r2)

saveRDS(final_r2, "data/intl_watersheds_refined/2.RDS")
```

```{r}
r24 <- readRDS("data/intl_watersheds_refined/24.RDS")

big_basins <- purrr::map(list.files(hydrobasins_dir,
                          pattern = "hybas.*lev05.*\\.shp$",
                          full.names = TRUE,
                          recursive = TRUE), ~sf::st_read(.x, quiet = TRUE)) %>%
  dplyr::bind_rows() %>%
  .[r24,] %>%
  sf::st_transform(4326) %>%
  sf::st_make_valid() %>%
  .[c(7),] %>%
  select(geometry)

mapview(r24) + big_basins + intl_unique %>% filter(rowid == 24)

final_r24 <- bind_rows(big_basins, r24) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(final_r24)

saveRDS(final_r24, "data/intl_watersheds_refined/24.RDS")
```

```{r}
r7 <- readRDS("data/intl_watersheds_refined/7.RDS")

big_basins <- purrr::map(list.files(hydrobasins_dir,
                          pattern = "hybas.*lev10.*\\.shp$",
                          full.names = TRUE,
                          recursive = TRUE), ~sf::st_read(.x, quiet = TRUE)) %>%
  dplyr::bind_rows() %>%
  .[r7,] %>%
  sf::st_transform(4326) %>%
  sf::st_make_valid() %>%
  .[c(12,48),] %>%
  select(geometry)

mapview(r7) + big_basins + intl_unique %>% filter(rowid == 7)

final_r7 <- bind_rows(big_basins, r7) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(final_r7)

saveRDS(final_r7, "data/intl_watersheds_refined/7.RDS")
```

```{r}
# ---------------------------------------------------------------------------
# wrong flowline selected: 41
# ---------------------------------------------------------------------------
r41_flowline <- select_flowline_global(
  sites         = intl_unique %>% filter(rowid == 41),
  rivers        = hydro_rivers_review,
  river_buffer_m = 10000,
  id_col        = "rowid"
)

watersheds <- delineate_ws(
  sites               = intl_unique %>% filter(rowid == 41),
  flowline_selections = r41_flowline,
  flowlines_network   = hydro_rivers,
  basins              = basins,
  flow_acc_raster     = hyd_glo_acc_15s,
  flow_dir_raster     = hyd_glo_dir_15s,
  flowline_id_col     = "HYRIV_ID",
  downstream_col      = "NEXT_DOWN",
  snap_method         = "point",
  local_basin_pct     = 0.80,
  id_col              = "rowid",
  hybas_id_col        = "HYBAS_ID",
  file_path = "data/intl_watersheds_refined/",
  use_s2 = FALSE
)

mapview(readRDS("data/intl_watersheds_refined/41.RDS")) + intl_unique %>% filter(rowid == 41)

# ---------------------------------------------------------------------------
#  too small: rowid 13, 19, 21, 33
# ---------------------------------------------------------------------------
delineate_neighborhood <- function(site,
                                   flow_acc_raster,
                                   flow_dir_raster,
                                   search_buffer_m = 500000,
                                   id_col          = "rowid") {
  
  stopifnot(inherits(site,            "sf"))
  stopifnot(inherits(flow_acc_raster, "SpatRaster"))
  stopifnot(inherits(flow_dir_raster, "SpatRaster"))
  stopifnot(id_col %in% names(site))
  
  site      <- sf::st_transform(site, 4326)
  out_rowid <- site[[id_col]][1]
  acc_crs   <- sf::st_crs(flow_acc_raster)
  
  message(sprintf("[%s] Delineating 9-cell neighborhood watersheds", out_rowid))
  
  # ---------------------------------------------------------------------------
  # 1. Crop rasters to buffered bbox around site
  # ---------------------------------------------------------------------------
  site_buf_vect <- site %>%
    sf::st_transform(3857) %>%
    sf::st_buffer(search_buffer_m) %>%
    sf::st_transform(acc_crs) %>%
    terra::vect()
  
  acc_local <- terra::crop(flow_acc_raster, site_buf_vect)
  fdr_local <- terra::crop(flow_dir_raster, site_buf_vect)
  acc_col   <- names(acc_local)[1]
  
  message(sprintf("  Local raster: %d x %d cells",
                  terra::nrow(fdr_local), terra::ncol(fdr_local)))
  
  # ---------------------------------------------------------------------------
  # 2. Find the cell the site falls in, then get all 9 neighborhood cells
  #    (the site cell + its 8 immediate neighbours)
  # ---------------------------------------------------------------------------
  site_proj  <- sf::st_transform(site, acc_crs)
  site_vect  <- terra::vect(site_proj)
  center_cell <- terra::cellFromXY(acc_local, terra::crds(site_vect))
  
  if (is.na(center_cell))
    stop(sprintf("[%s] Site falls outside raster extent.", out_rowid))
  
  center_rc <- terra::rowColFromCell(acc_local, center_cell)
  cr        <- center_rc[1]
  cc        <- center_rc[2]
  nrows     <- terra::nrow(acc_local)
  ncols     <- terra::ncol(acc_local)
  
  # All 9 cells: center + 8 neighbours in row/col offsets
  offsets <- expand.grid(dr = -1L:1L, dc = -1L:1L)
  
  neighborhood <- purrr::pmap_dfr(offsets, function(dr, dc) {
    nr <- cr + dr
    nc <- cc + dc
    if (nr < 1 || nr > nrows || nc < 1 || nc > ncols) return(NULL)
    cell  <- terra::cellFromRowCol(acc_local, nr, nc)
    xy    <- terra::xyFromCell(acc_local, cell)
    acc   <- terra::extract(acc_local, cell)[[acc_col]][1]
    tibble::tibble(
      cell_id  = paste0("r", dr + 2L, "c", dc + 2L),  # label: r1c1 to r3c3
      row      = nr,
      col      = nc,
      cell_idx = cell,
      x        = xy[1],
      y        = xy[2],
      acc_val  = acc,
      dr       = dr,
      dc       = dc
    )
  })
  
  message(sprintf("  Neighborhood cells: %d", nrow(neighborhood)))
  message(sprintf("  Accumulation values range: %.0f to %.0f",
                  min(neighborhood$acc_val, na.rm = TRUE),
                  max(neighborhood$acc_val, na.rm = TRUE)))
  
  # ---------------------------------------------------------------------------
  # 3. Pre-load flow direction matrix once for BFS reuse
  # ---------------------------------------------------------------------------
  fdr_matrix <- terra::as.matrix(fdr_local, wide = TRUE)
  nrows_fdr  <- nrow(fdr_matrix)
  ncols_fdr  <- ncol(fdr_matrix)
  
  drains_into <- c(
    "1"   = 16L, "2"   = 32L, "4"   = 64L,  "8"   = 128L,
    "16"  = 1L,  "32"  = 2L,  "64"  = 4L,   "128" = 8L
  )
  
  neighbour_offsets <- list(
    "1"   = c( 0L,  1L), "2"   = c( 1L,  1L),
    "4"   = c( 1L,  0L), "8"   = c( 1L, -1L),
    "16"  = c( 0L, -1L), "32"  = c(-1L, -1L),
    "64"  = c(-1L,  0L), "128" = c(-1L,  1L)
  )
  
  # ---------------------------------------------------------------------------
  # 4. BFS helper — delineates upstream from a single pour point cell
  # ---------------------------------------------------------------------------
  bfs_upstream <- function(pour_row, pour_col) {
    
    visited      <- matrix(FALSE, nrow = nrows_fdr, ncol = ncols_fdr)
    visited[pour_row, pour_col] <- TRUE
    
    max_cells <- nrows_fdr * ncols_fdr
    up_r      <- integer(max_cells)
    up_c      <- integer(max_cells)
    up_r[1]   <- pour_row
    up_c[1]   <- pour_col
    cell_ptr  <- 1L
    frontier  <- matrix(c(pour_row, pour_col), nrow = 1)
    
    repeat {
      new_r <- integer(0)
      new_c <- integer(0)
      
      for (fi in seq_len(nrow(frontier))) {
        r <- frontier[fi, 1]
        c <- frontier[fi, 2]
        
        for (dir_name in names(neighbour_offsets)) {
          off <- neighbour_offsets[[dir_name]]
          nr  <- r + off[1]
          nc  <- c + off[2]
          if (nr < 1 || nr > nrows_fdr || nc < 1 || nc > ncols_fdr) next
          if (visited[nr, nc]) next
          nbr_fdr <- fdr_matrix[nr, nc]
          if (is.na(nbr_fdr)) next
          if (nbr_fdr == drains_into[dir_name]) {
            visited[nr, nc] <- TRUE
            new_r <- c(new_r, nr)
            new_c <- c(new_c, nc)
          }
        }
      }
      
      if (length(new_r) == 0) break
      
      n_new    <- length(new_r)
      up_r[(cell_ptr + 1L):(cell_ptr + n_new)] <- new_r
      up_c[(cell_ptr + 1L):(cell_ptr + n_new)] <- new_c
      cell_ptr <- cell_ptr + n_new
      frontier <- matrix(c(new_r, new_c), ncol = 2)
    }
    
    list(
      up_r     = up_r[seq_len(cell_ptr)],
      up_c     = up_c[seq_len(cell_ptr)],
      n_cells  = cell_ptr
    )
  }
  
  # ---------------------------------------------------------------------------
  # 5. Vectorize BFS result to polygon
  # ---------------------------------------------------------------------------
  cells_to_polygon <- function(up_r, up_c) {
    cells <- terra::cellFromRowCol(fdr_local, up_r, up_c)
    ws_rast       <- terra::rast(fdr_local)
    ws_rast[]     <- NA
    ws_rast[cells] <- 1L
    
    tryCatch({
      terra::as.polygons(ws_rast) %>%
        sf::st_as_sf() %>%
        sf::st_set_crs(acc_crs) %>%
        sf::st_transform(4326) %>%
        sf::st_make_valid() %>%
        sf::st_union() %>%
        sf::st_as_sf() %>%
        dplyr::rename(geometry = 1) %>%
        sf::st_set_geometry("geometry")
    }, error = function(e) {
      message(sprintf("    Vectorization failed: %s", conditionMessage(e)))
      NULL
    })
  }
  
  # ---------------------------------------------------------------------------
  # 6. Delineate watershed for each of the 9 neighborhood cells
  # ---------------------------------------------------------------------------
  
  ws_list <- vector("list", nrow(neighborhood))
  
  for (i in seq_len(nrow(neighborhood))) {
    
    cell_info <- neighborhood[i, ]
    
    message(sprintf("  [%d/9] Cell %s | row=%d col=%d | acc=%.0f",
                    i, cell_info$cell_id,
                    cell_info$row, cell_info$col,
                    cell_info$acc_val))
    
    bfs_result <- bfs_upstream(cell_info$row, cell_info$col)
    
    message(sprintf("    Upstream cells: %d", bfs_result$n_cells))
    
    ws_poly <- cells_to_polygon(bfs_result$up_r, bfs_result$up_c)
    
    if (is.null(ws_poly)) next
    
    ws_poly <- ws_poly %>%
      dplyr::mutate(
        "{id_col}"    := out_rowid,
        cell_id       = cell_info$cell_id,
        cell_row      = cell_info$row,
        cell_col      = cell_info$col,
        pour_x        = cell_info$x,
        pour_y        = cell_info$y,
        acc_val       = cell_info$acc_val,
        dr            = cell_info$dr,
        dc            = cell_info$dc,
        is_center     = (cell_info$dr == 0 & cell_info$dc == 0),
        area_km2      = as.numeric(sf::st_area(.)) / 1e6
      )
    
    ws_list[[i]] <- ws_poly
    
    message(sprintf("    Area: %.1f km²", ws_poly$area_km2[1]))
  }
  
  # ---------------------------------------------------------------------------
  # 7. Bind all 9 watersheds and save
  # ---------------------------------------------------------------------------
  out <- dplyr::bind_rows(Filter(Negate(is.null), ws_list))
  
  return(out)
  
}

hyd_glo_acc_15s <- terra::rast("data/hyd_glo_acc_15s/hyd_glo_acc_15s.tif")

hyd_glo_dir_15s <- terra::rast("data/hyd_glo_dir_15s/hyd_glo_dir_15s.tif")

r13 <- delineate_neighborhood(site = intl_unique %>% dplyr::filter(rowid == 13),
                              flow_acc_raster = hyd_glo_acc_15s,
                              flow_dir_raster = hyd_glo_dir_15s,
                              search_buffer_m = 500000,
                              id_col          = "rowid") %>%
  rowid_to_column("index")

m <- mapview::mapview(intl_unique %>% dplyr::filter(rowid == 13),
                      col.regions = "red",
                      cex         = 8,
                      layer.name  = "Site")

for (i in seq_len(nrow(r13))) {
  row  <- r13[i, ]
  name <- sprintf("Cell %s | acc=%.0f | area=%.0f km²",
                  row$cell_id, row$acc_val, row$area_km2)
  m <- m + mapview::mapview(row, layer.name = name)
}

m

r13 <- r13 %>%
  .[c(3, 5, 6, 8),] %>%
  summarize() %>%
  mutate(rowid = 13)

mapview(r13) + intl_unique %>% dplyr::filter(rowid == 13)

saveRDS(r13, "data/intl_watersheds_refined/13.RDS")


r19 <- delineate_neighborhood(site = intl_unique %>% dplyr::filter(rowid == 19),
                              flow_acc_raster = hyd_glo_acc_15s,
                              flow_dir_raster = hyd_glo_dir_15s,
                              search_buffer_m = 500000,
                              id_col          = "rowid") %>%
  rowid_to_column("index")

m <- mapview::mapview(intl_unique %>% dplyr::filter(rowid == 19),
                      col.regions = "red",
                      cex         = 8,
                      layer.name  = "Site")

for (i in seq_len(nrow(r19))) {
  row  <- r19[i, ]
  name <- sprintf("Cell %s | acc=%.0f | area=%.0f km²",
                  row$cell_id, row$acc_val, row$area_km2)
  m <- m + mapview::mapview(row, layer.name = name)
}

m

r19_ <- r19 %>%
  .[c(7,8),] %>%
  summarize() %>%
  mutate(rowid = 19)

mapview(r19_) + intl_unique %>% dplyr::filter(rowid == 19)

saveRDS(r19_, "data/intl_watersheds_refined/19.RDS")


r21 <- delineate_neighborhood(site = intl_unique %>% dplyr::filter(rowid == 21),
                              flow_acc_raster = hyd_glo_acc_15s,
                              flow_dir_raster = hyd_glo_dir_15s,
                              search_buffer_m = 500000,
                              id_col          = "rowid") %>%
  rowid_to_column("index")

m <- mapview::mapview(intl_unique %>% dplyr::filter(rowid == 21),
                      col.regions = "red",
                      cex         = 8,
                      layer.name  = "Site")

for (i in seq_len(nrow(r21))) {
  row  <- r21[i, ]
  name <- sprintf("Cell %s | acc=%.0f | area=%.0f km²",
                  row$cell_id, row$acc_val, row$area_km2)
  m <- m + mapview::mapview(row, layer.name = name)
}

m

r21_ <- r21 %>%
  .[c(8),] %>%
  summarize() %>%
  mutate(rowid = 21)

mapview(r21_) + intl_unique %>% dplyr::filter(rowid == 21)

saveRDS(r21_, "data/intl_watersheds_refined/21.RDS")


r33 <- delineate_neighborhood(site = intl_unique %>% dplyr::filter(rowid == 33),
                              flow_acc_raster = hyd_glo_acc_15s,
                              flow_dir_raster = hyd_glo_dir_15s,
                              search_buffer_m = 500000,
                              id_col          = "rowid") %>%
  rowid_to_column("index")

m <- mapview::mapview(intl_unique %>% dplyr::filter(rowid == 33),
                      col.regions = "red",
                      cex         = 8,
                      layer.name  = "Site")

for (i in seq_len(nrow(r33))) {
  row  <- r33[i, ]
  name <- sprintf("Cell %s | acc=%.0f | area=%.0f km²",
                  row$cell_id, row$acc_val, row$area_km2)
  m <- m + mapview::mapview(row, layer.name = name)
}

m

r33_ <- r33 %>%
  .[c(5),] %>%
  summarize() %>%
  mutate(rowid = 33)

mapview(r33_) + intl_unique %>% dplyr::filter(rowid == 33)

saveRDS(r33_, "data/intl_watersheds_refined/33.RDS")
# ---------------------------------------------------------------------------
#  needs additional flowlines delineated/review associated flowline: 30, 22, 27, 40, 45, 49, 7
# ---------------------------------------------------------------------------
wonky_sites_update_a <- select_flowline_global(sites = intl_unique %>% 
                                                 filter(rowid %in% c(30, 22, 27, 40, 45, 49, 7)),
                                               rivers = hydro_rivers_review,
                                               river_buffer_m = 10000,
                                               id_col = "rowid")

still_weird <- still_weird %>%
  filter(review != "Y") %>%
  pull(rowid)

still_weird


r30 <- readRDS("data/intl_watersheds_refined/30.RDS") %>%
  select(rowid)
r30_basins <- basins %>%
  .[r30,] %>%
  .[c(222),]
r30 <- r30 %>%
  bind_rows(r30_basins) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(r30) +  mapview(r30_basins, col.region = "red") + intl_unique %>% filter(rowid == 30) + hydro_rivers_review

saveRDS(r30, "data/intl_watersheds_refined/30.RDS")

r22 <- readRDS("data/intl_watersheds_refined/22.RDS") %>%
  select(rowid)
r22_basins <- basins %>%
  .[r22,] %>%
  .[c(63, 64, 65, 70),]
r22 <- r22 %>%
  bind_rows(r22_basins) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(r22) +  mapview(r22_basins, col.region = "red") + intl_unique %>% filter(rowid == 22) #hydro_rivers_review

saveRDS(r22, "data/intl_watersheds_refined/22.RDS")

r27 <- readRDS("data/intl_watersheds_refined/27.RDS") %>%
  select(rowid)

r27_basins <- basins %>%
  .[r27,] %>%
  .[c(12, 11),]

r27 <- r27 %>%
  bind_rows(r27_basins) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(r27) +  mapview(r27_basins, col.region = "red") + intl_unique %>% filter(rowid == 27) + hydro_rivers_review
saveRDS(r27, "data/intl_watersheds_refined/27.RDS")

r40 <- readRDS("data/intl_watersheds_refined/40.RDS") %>%
  select(rowid)

r40_basins <- basins %>%
  .[r40,] %>%
  .[c(8),]

r40 <- r40 %>%
  bind_rows(r40_basins) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(r40) +  mapview(r40_basins, col.region = "red") + intl_unique %>% filter(rowid == 40) + hydro_rivers_review

saveRDS(r40, "data/intl_watersheds_refined/40.RDS")

r45 <- readRDS("data/intl_watersheds_refined/45.RDS") %>%
  select(rowid)

r45_basins <- basins %>%
  .[r45,] %>%
  .[c(17),]

r45 <- r45 %>%
  bind_rows(r45_basins) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(r45) +  mapview(r45_basins, col.region = "red") + intl_unique %>% filter(rowid == 45) + hydro_rivers_review

saveRDS(r45, "data/intl_watersheds_refined/45.RDS")

r49 <- readRDS("data/intl_watersheds_refined/49.RDS") %>%
  select(rowid)

r49_basins <- basins %>%
  .[r49,] %>%
  .[c(250, 254),]

r49 <- r49 %>%
  bind_rows(r49_basins) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(r49) +  mapview(r49_basins, col.region = "red") + intl_unique %>% filter(rowid == 49) + hydro_rivers_review

saveRDS(r49, "data/intl_watersheds_refined/49.RDS")


r7 <- readRDS("data/intl_watersheds_refined/7.RDS") %>%
  select(rowid)

r7_basins <- basins %>%
  .[r7,] %>%
  .[c(14),]

r7 <- r7 %>%
  bind_rows(r7_basins) %>%
  summarize() %>%
  nngeo::st_remove_holes()

mapview(r7) +  mapview(r7_basins, col.region = "red") + intl_unique %>% filter(rowid == 7) + hydro_rivers_review

saveRDS(r7, "data/intl_watersheds_refined/7.RDS")

# ---------------------------------------------------------------------------
#  some downstream bits still showing up: 23
# ---------------------------------------------------------------------------
r23_flowline <- select_flowline_global(sites = intl_unique %>% filter(rowid ==23),
                                       rivers = hydro_rivers_review,
                                       river_buffer_m = 10000,
                                       id_col = "rowid")

delineate_ws(sites = intl_unique %>% filter(rowid == 23),
             flowline_selections = r23_flowline,
             flowlines_network = hydro_rivers,
             basins = basins,
             flow_acc_raster = hyd_glo_acc_30s,
             flow_dir_raster = hyd_glo_dir_30s,
             flowline_id_col = "HYRIV_ID",
             downstream_col = "NEXT_DOWN",
             snap_method = "point",
             local_basin_pct = 0.80,
             id_col = "rowid",
             hybas_id_col = "HYBAS_ID",
             file_path = "data/intl_watersheds_refined/",
             use_s2 = FALSE)

mapview(readRDS("data/intl_watersheds_refined/23.RDS")) + intl_unique %>% filter(rowid == 23)


# ---------------------------------------------------------------------------
# huge rivers -- estuarine hydro approach?: 43, 44, 
# ---------------------------------------------------------------------------

big_basin_1 <- basins %>%
  .[intl_unique %>% 
      filter(rowid == 43) %>% 
      sf::st_transform(3857) %>%
      sf::st_buffer(200000) %>%
      sf::st_transform(4326),]
mapview(big_basin_1) + intl_unique %>% filter(rowid == 43)

basins_crs <- sf::st_crs(big_basin_1)

# Work in meters for buffering / sliver handling
site_m <- intl_unique %>% 
  filter(rowid == 43) %>%
  sf::st_transform(3857)

basins_m <- big_basin_1 %>%
  sf::st_transform(3857) %>%
  sf::st_make_valid()

# Local search area around the point
search_area_m <- site_m %>%
  sf::st_buffer(100000) %>%
  sf::st_make_valid()

# Only keep basins near the point
local_basins_m <- basins_m[search_area_m, ]

# Make one dissolved polygon of all local basins
local_basin_union_m <- local_basins_m %>%
  sf::st_union() %>%
  sf::st_make_valid()

# The "empty spaces" are everything inside the search window
# that is not covered by HydroBASINS polygons.
gaps_m <- sf::st_difference(sf::st_geometry(search_area_m),
                            local_basin_union_m) %>%
  sf::st_collection_extract("POLYGON") %>%
  sf::st_as_sf() %>%
  sf::st_make_valid()


site_xy <- sf::st_coordinates(site_m)[1, ]

x0 <- site_xy[["X"]]
y0 <- site_xy[["Y"]]

# Make this longer than your search buffer if needed
clip_dist_m <- 500000

# Wider wedge to keep adjacent basins near the point
wedge_half_angle_deg <- 70

# SW direction
center_angle_deg <- 225

angle1 <- (center_angle_deg - wedge_half_angle_deg) * pi / 180
angle2 <- (center_angle_deg + wedge_half_angle_deg) * pi / 180

sw_wedge_coords <- matrix(c(x0, y0,
                            x0 + clip_dist_m * cos(angle1), y0 + clip_dist_m * sin(angle1),
                            x0 + clip_dist_m * cos(angle2), y0 + clip_dist_m * sin(angle2),
                            x0, y0),
                          ncol = 2,
                          byrow = TRUE)

sw_wedge_m <- sf::st_sf(geometry = sf::st_sfc(sf::st_polygon(list(sw_wedge_coords)),
                                              crs = sf::st_crs(site_m)))

site_keep_m <- site_m %>%
  sf::st_buffer(8000)

sw_clip_m <- rbind(sw_wedge_m,
                   site_keep_m %>% dplyr::select(geometry)) %>%
  sf::st_union() %>%
  sf::st_as_sf() %>%
  sf::st_make_valid()

gaps_sw_m <- sf::st_intersection(
  gaps_m,
  sw_clip_m) %>%
  sf::st_make_valid() %>%
  st_cast("POLYGON") %>%
  .[-c(13:16),] 

mapview(gaps_sw_m, col.region = "red") + site_m + big_basin_1 + hydro_rivers %>% .[big_basin_1 %>% st_buffer(1),]

names(gaps_sw_m)[names(gaps_sw_m) == "x"] <- "geometry"
st_geometry(gaps_sw_m) <- "geometry"

touching_basins_m <- big_basin_1[gaps_sw_m %>% st_buffer(800) %>% st_transform(4326),]

mapview(touching_basins_m, col.region = "red") + mapview(gaps_sw_m, col.region = "blue") + site_m

touching_basins = touching_basins_m %>%
  sf::st_transform(basins_crs)

touching_flowlines <- hydro_rivers %>%
  .[touching_basins,] %>%
  # Amazonas
  bind_rows(hydro_rivers %>%
              dplyr::filter(.data$HYRIV_ID  == 60496312))

# 2. Trace upstream HydroRIVERS flowlines
get_upstream_flowline_ids <- function(target_flowline_ids,
                                      rivers,
                                      id_col = "HYRIV_ID",
                                      down_col = "NEXT_DOWN") {
  
  river_links <- rivers %>%
    sf::st_drop_geometry() %>%
    dplyr::transmute(
      river_id = .data[[id_col]],
      downstream_id = .data[[down_col]]
    ) %>%
    dplyr::filter(!is.na(.data$river_id))
  
  upstream_ids <- unique(target_flowline_ids)
  
  repeat {
    
    next_ids <- river_links %>%
      dplyr::filter(.data$downstream_id %in% upstream_ids) %>%
      dplyr::pull(.data$river_id) %>%
      unique()
    
    next_ids <- setdiff(next_ids, upstream_ids)
    
    if (length(next_ids) == 0) {
      break
    }
    
    upstream_ids <- c(upstream_ids, next_ids)
  }
  
  unique(upstream_ids)
}

target_flowline_ids <- touching_flowlines %>%
  sf::st_drop_geometry() %>%
  dplyr::pull(HYRIV_ID) %>%
  unique() %>%
  stats::na.omit()

upstream_flowline_ids_43 <- get_upstream_flowline_ids(target_flowline_ids = target_flowline_ids,
                                                      rivers = hydro_rivers,
                                                      id_col = "HYRIV_ID",
                                                      down_col = "NEXT_DOWN")


upstream_flowlines_43 <- hydro_rivers %>%
  dplyr::filter(.data$HYRIV_ID %in% upstream_flowline_ids_43) %>%
  sf::st_make_valid()

mapview(upstream_flowlines_43)

# Grab all HydroBASINS touched by upstream flowlines
upstream_flowlines_buffer_43 <- upstream_flowlines_43 %>%
  sf::st_union() %>%
  sf::st_as_sf() %>%
  sf::st_make_valid()

sel_43 <- basins[upstream_flowlines_43, ] %>%
  dplyr::bind_rows(gaps_sw_m %>% sf::st_transform(4326))

upstream_basins_43 <- terra::vect(sel_43) %>%
  terra::aggregate() %>%  
  sf::st_as_sf() %>%
  nngeo::st_remove_holes() 

mapview::mapview(upstream_basins_43, color = "orange", alpha.regions = 0.25) +
  # mapview::mapview(upstream_flowlines_43, color = "blue", lwd = 2) +
  mapview::mapview(intl_unique %>% dplyr::filter(rowid == 43),
                   col.regions = "black", cex = 8)

saveRDS(upstream_basins_43 %>% mutate(rowid = 43), "data/intl_watersheds_refined/43.RDS")

# NEXT, 44, 

mapview(big_basin_1) + intl_unique %>% filter(rowid %in% c(44, 43))

basins_crs <- sf::st_crs(big_basin_1)


# Work in meters for buffering / sliver handling
site_m <- intl_unique %>% 
  filter(rowid == 44) %>%
  sf::st_transform(3857)

basins_m <- big_basin_1 %>%
  sf::st_transform(3857) %>%
  sf::st_make_valid()

# Local search area around the point
search_area_m <- site_m %>%
  sf::st_buffer(80000) %>%
  sf::st_make_valid()

# Only keep basins near the point
local_basins_m <- basins_m[search_area_m, ]

# Make one dissolved polygon of all local basins
local_basin_union_m <- local_basins_m %>%
  sf::st_union() %>%
  sf::st_make_valid()

# The "empty spaces" are everything inside the search window
# that is not covered by HydroBASINS polygons.
gaps_m <- sf::st_difference(sf::st_geometry(search_area_m),
                            local_basin_union_m) %>%
  sf::st_collection_extract("POLYGON") %>%
  sf::st_as_sf() %>%
  sf::st_make_valid()


site_xy <- sf::st_coordinates(site_m)[1, ]

x0 <- site_xy[["X"]]
y0 <- site_xy[["Y"]]

# Make this longer than your search buffer if needed
clip_dist_m <- 200000

# Wider wedge to keep adjacent basins near the point
wedge_half_angle_deg <- 70

# SW direction
center_angle_deg <- 292

angle1 <- (center_angle_deg - wedge_half_angle_deg) * pi / 180
angle2 <- (center_angle_deg + wedge_half_angle_deg) * pi / 180

sw_wedge_coords <- matrix(
  c(
    x0, y0,
    x0 + clip_dist_m * cos(angle1), y0 + clip_dist_m * sin(angle1),
    x0 + clip_dist_m * cos(angle2), y0 + clip_dist_m * sin(angle2),
    x0, y0
  ),
  ncol = 2,
  byrow = TRUE
)

sw_wedge_m <- sf::st_sf(
  geometry = sf::st_sfc(
    sf::st_polygon(list(sw_wedge_coords)),
    crs = sf::st_crs(site_m)
  )
)

site_keep_m <- site_m %>%
  sf::st_buffer(9000)

sw_clip_m <- rbind(
  sw_wedge_m,
  site_keep_m %>% dplyr::select(geometry)
) %>%
  sf::st_union() %>%
  sf::st_as_sf() %>%
  sf::st_make_valid()

gaps_sw_m <- sf::st_intersection(
  gaps_m,
  sw_clip_m
) %>%
  sf::st_make_valid() %>%
  st_cast("POLYGON") %>%
  .[-c(1, 8:13),] 

mapview(gaps_sw_m, col.region = "red") + site_m + big_basin_1

names(gaps_sw_m)[names(gaps_sw_m) == "x"] <- "geometry"
st_geometry(gaps_sw_m) <- "geometry"

touching_basins_m <- big_basin_1[gaps_sw_m %>% st_buffer(100) %>% st_transform(4326),]

mapview(touching_basins_m, col.region = "red") + mapview(gaps_sw_m, col.region = "blue") + site_m

touching_basins = touching_basins_m %>%
  sf::st_transform(basins_crs)

touching_flowlines <- hydro_rivers %>%
  .[touching_basins,] %>%
  # Amazonas
  bind_rows(hydro_rivers %>%
              dplyr::filter(.data$HYRIV_ID == 60496312))

# 2. Trace upstream HydroRIVERS flowlines
get_upstream_flowline_ids <- function(target_flowline_ids,
                                      rivers,
                                      id_col = "HYRIV_ID",
                                      down_col = "NEXT_DOWN") {
  
  river_links <- rivers %>%
    sf::st_drop_geometry() %>%
    dplyr::transmute(river_id = .data[[id_col]],
      downstream_id = .data[[down_col]]) %>%
    dplyr::filter(!is.na(.data$river_id))
  
  upstream_ids <- unique(target_flowline_ids)
  
  repeat {
    
    next_ids <- river_links %>%
      dplyr::filter(.data$downstream_id %in% upstream_ids) %>%
      dplyr::pull(.data$river_id) %>%
      unique()
    
    next_ids <- setdiff(next_ids, upstream_ids)
    
    if (length(next_ids) == 0) {
      break
    }
    
    upstream_ids <- c(upstream_ids, next_ids)
  }
  
  unique(upstream_ids)
}

target_flowline_ids <- touching_flowlines %>%
  sf::st_drop_geometry() %>%
  dplyr::pull(HYRIV_ID) %>%
  unique() %>%
  stats::na.omit()

upstream_flowline_ids_44 <- get_upstream_flowline_ids(target_flowline_ids = target_flowline_ids,
                                                      rivers = hydro_rivers,
                                                      id_col = "HYRIV_ID",
                                                      down_col = "NEXT_DOWN")


upstream_flowlines_44 <- hydro_rivers %>%
  dplyr::filter(.data$HYRIV_ID %in% upstream_flowline_ids_43) %>%
  sf::st_make_valid()

mapview(upstream_flowlines_44)

# 3. Grab all HydroBASINS touched by upstream flowlines
upstream_flowlines_buffer_44 <- upstream_flowlines_44 %>%
  sf::st_union() %>%
  sf::st_as_sf() %>%
  sf::st_make_valid()

sel_44 <- basins[upstream_flowlines_44, ] %>%
  dplyr::bind_rows(gaps_sw_m %>% sf::st_transform(4326))

upstream_basins_44 <- terra::vect(sel_44) %>%
  terra::aggregate() %>%  
  sf::st_as_sf() %>%
  nngeo::st_remove_holes() 

mapview::mapview(upstream_basins_44, color = "orange", alpha.regions = 0.25) +
  mapview::mapview(upstream_flowlines_44, color = "blue", lwd = 2) +
  mapview::mapview(intl_unique %>% dplyr::filter(rowid == 43),
                   col.regions = "black", cex = 8)

saveRDS(upstream_basins_44 %>% mutate(rowid = 44), "data/intl_watersheds_refined/44.RDS")


# ---------------------------------------------------------------------------
# "holes" need filled: 47
# ---------------------------------------------------------------------------
r47 <- readRDS("data/intl_watersheds_refined/47.RDS") %>%
  sf::st_transform(3857) %>%
  sf::st_make_valid() %>%
  sf::st_buffer(1000) %>%
  nngeo::st_remove_holes() %>%
  sf::st_buffer(-1000) %>%
  sf::st_make_valid() %>%
  sf::st_transform(4326)

saveRDS(r47, "data/intl_watersheds_refined/47.RDS")

mapview(r47) + intl_unique %>% filter(rowid == 47)

# ---------------------------------------------------------------------------
# needs to be run with flow dir and flow acc: 15, 34
# --------------------------------------------------------------------------- 
delineate_ws(sites = intl_unique %>% filter(rowid %in% c(15, 34)),
             flowline_selections = flowline_selections,
             flowlines_network = hydro_rivers,
             basins = basins,
             flow_acc_raster = hyd_glo_acc_30s,
             flow_dir_raster = hyd_glo_dir_30s,
             flowline_id_col = "HYRIV_ID",
             downstream_col = "NEXT_DOWN",
             snap_method = "point",
             local_basin_pct = 0.80,
             id_col = "rowid",
             hybas_id_col = "HYBAS_ID",
             file_path = "data/intl_watersheds_refined/",
             use_s2 = FALSE)

mapview(readRDS("data/intl_watersheds_refined/15.RDS")) + intl_unique %>% filter(rowid == 15)
mapview(readRDS("data/intl_watersheds_refined/34.RDS")) + intl_unique %>% filter(rowid == 34)

# ---------------------------------------------------------------------------
# yojoa: 35, 36, 37, 38, 39
# --------------------------------------------------------------------------- 
yojoa_basins <- basins %>%
  .[intl_unique %>% 
      filter(rowid %in% 35:39) %>% 
      sf::st_transform(3857) %>%
      sf::st_buffer(2000) %>%
      sf::st_transform(4326),] %>%
  summarize()

yojoa_basins_by_rowid <- sf::st_sf(
  rowid = 35:39,
  geometry = sf::st_geometry(yojoa_basins)[rep(1, length(35:39))]
)

output_dir <- "data/intl_watersheds_refined"

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

yojoa_basins_by_rowid %>%
  split(.$rowid) %>%
  purrr::iwalk(~{
    saveRDS(
      .x,
      file = file.path(output_dir, paste0(.y, ".rds"))
    )
  })

```

Re-review:
```{r}
watersheds <- list.files("data/intl_watersheds_refined/", full.names = TRUE) %>%
  purrr::map_dfr(~readRDS(.x) %>%
                   dplyr::mutate(rowid = as.integer(tools::file_path_sans_ext(basename(.x)))) %>%
                   dplyr::select(rowid) %>%
                   nngeo::st_remove_holes())

big_basin_files <- list.files(hydrobasins_dir,
                          pattern = "hybas.*lev03.*\\.shp$",
                          full.names = TRUE,
                          recursive = TRUE)

big_basins <- purrr::map(big_basin_files, ~sf::st_read(.x, quiet = TRUE)) %>%
  dplyr::bind_rows() %>%
  sf::st_transform(4326) %>%
  sf::st_make_valid()

big_basins_review <- big_basins %>%
  st_join(., watersheds, left = FALSE) %>%
  select(HYBAS_ID, rowid)


review_watersheds_global_final <- function(watersheds,
                                     sites,
                                     basins = NULL,
                                     rivers = NULL,
                                     site_buffer_m = 2500,
                                     id_col = "rowid",
                                     review_col = "review") {

  stopifnot(inherits(watersheds, "sf"))
  stopifnot(inherits(sites, "sf"))
  stopifnot(id_col %in% names(watersheds))
  stopifnot(id_col %in% names(sites))

  if (!is.null(basins)) {
    stopifnot(inherits(basins, "sf"))
  }

  if (!is.null(rivers)) {
    stopifnot(inherits(rivers, "sf"))
    stopifnot(id_col %in% names(rivers))
  }

  if (!review_col %in% names(watersheds)) {
    watersheds[[review_col]] <- NA_character_
  }

  remaining <- which(is.na(watersheds[[review_col]]))

  if (length(remaining) == 0) {
    message("All international watersheds have been reviewed!")
    return(watersheds)
  }

  message(sprintf("Starting review. %d of %d international watersheds remaining.\n",
                  length(remaining), nrow(watersheds)))
  message("Workflow:")
  message("  Single map: upstream watershed + HydroBASINS overlapping the watershed extent + HydroRIVERS + site, zoomed to the rivers extent")
  message("Commands: y = Yes | s = Small | b = Big | m = Maybe/manual | q = Quit & save\n")

  # ---------------------------------------------------------------------------
  # Pre-project all layers once to WGS84 before the loop
  # ---------------------------------------------------------------------------
  watersheds_wgs <- sf::st_transform(watersheds, 4326)
  sites_wgs      <- sf::st_transform(sites, 4326)
  basins_wgs     <- if (!is.null(basins)) sf::st_transform(basins, 4326) else NULL
  rivers_wgs     <- if (!is.null(rivers)) sf::st_transform(rivers, 4326) else NULL

  for (i in remaining) {

    watershed <- watersheds_wgs[i, ]
    site_id   <- watershed[[id_col]][1]

    site <- sites_wgs %>%
      dplyr::filter(.data[[id_col]] == site_id)

    # -------------------------------------------------------------------------
    # Basins = all HydroBASINS polygons overlapping the watershed extent (bbox).
    # -------------------------------------------------------------------------
    local_basin <- NULL
    if (!is.null(basins_wgs)) {
      ws_extent <- sf::st_as_sfc(sf::st_bbox(watershed))
      hits <- lengths(sf::st_intersects(basins_wgs, ws_extent)) > 0
      local_basin <- basins_wgs[hits, ]
      # Drop duplicate basin polygons (the same basin can appear once per
      # rowid/site it maps to) so each distinct basin is drawn only once.
      local_basin <- local_basin[!duplicated(sf::st_geometry(local_basin)), ]
      if (nrow(local_basin) == 0) local_basin <- NULL
    }

    nearby_rivers <- NULL
    if (!is.null(rivers_wgs)) {
      nearby_rivers <- rivers_wgs %>%
        dplyr::filter(.data[[id_col]] == site_id)

      if (nrow(nearby_rivers) == 0) {
        nearby_rivers <- NULL
      }
    }

    # -------------------------------------------------------------------------
    # Readable lat/lon from the site geometry if available
    # -------------------------------------------------------------------------
    if (nrow(site) > 0) {
      site_coords <- sf::st_coordinates(sf::st_geometry(site))[1, ]
      site_lon <- round(site_coords[["X"]], 6)
      site_lat <- round(site_coords[["Y"]], 6)
    } else {
      site_lon <- NA_real_
      site_lat <- NA_real_
    }

    # -------------------------------------------------------------------------
    # Bounding box = rivers extent (with sensible fallbacks if no rivers match)
    # -------------------------------------------------------------------------
    if (!is.null(nearby_rivers) && nrow(nearby_rivers) > 0) {
      bb <- sf::st_bbox(nearby_rivers)
    } else if (!is.null(local_basin) && nrow(local_basin) > 0) {
      bb <- sf::st_bbox(local_basin)
    } else if (nrow(site) > 0) {
      bb <- site %>%
        sf::st_transform(3857) %>%
        sf::st_buffer(site_buffer_m) %>%
        sf::st_transform(4326) %>%
        sf::st_bbox()
    } else {
      bb <- sf::st_bbox(watershed)
    }

    # -------------------------------------------------------------------------
    # Single map
    # -------------------------------------------------------------------------
    m <- mapview::mapview(
      watershed,
      alpha.regions = 0.2,
      color = "orange",
      layer.name = "Full upstream watershed"
    )

    if (!is.null(local_basin) && nrow(local_basin) > 0) {
      m <- m + mapview::mapview(
        local_basin,
        alpha.regions = 0.15,
        color = "purple",
        layer.name = "Overlapping HydroBASINS"
      )
    }

    if (!is.null(nearby_rivers) && nrow(nearby_rivers) > 0) {
      m <- m + mapview::mapview(
        nearby_rivers,
        color = "steelblue",
        lwd = 3,
        layer.name = "rowid-matched HydroRIVERS"
      )
    }

    if (nrow(site) > 0) {
      m <- m + mapview::mapview(
        site,
        col.regions = "red",
        cex = 8,
        layer.name = "Sample location"
      )
    }

    m@map <- m@map %>%
      leaflet::fitBounds(
        lng1 = bb[["xmin"]], lat1 = bb[["ymin"]],
        lng2 = bb[["xmax"]], lat2 = bb[["ymax"]]
      )

    print(m)

    message(sprintf("[%d / %d] %s: %s | lat/lon: %s, %s",
                    match(i, remaining), length(remaining),
                    id_col, site_id,
                    site_lat, site_lon))

    if (!is.null(local_basin) && nrow(local_basin) > 0) {
      message(sprintf("HydroBASINS polygons shown: %d", nrow(local_basin)))
    }

    if (!is.null(nearby_rivers) && nrow(nearby_rivers) > 0) {
      message(sprintf("HydroRIVERS segments shown: %d", nrow(nearby_rivers)))
    }

    response <- ""
    while (!response %in% c("y", "s", "b", "m", "q")) {
      response <- tolower(trimws(readline("Accept? [y/s/b/m/q]: ")))
    }

    if (response == "q") {
      message("Quitting. Progress saved.")
      break
    }

    watersheds[[review_col]][i] <- toupper(response)
    message(sprintf("  Saved: %s\n", toupper(response)))
  }

  reviewed  <- sum(!is.na(watersheds[[review_col]]))
  remaining <- sum(is.na(watersheds[[review_col]]))
  message(sprintf("Done. %d reviewed, %d remaining.", reviewed, remaining))

  watersheds
}

final_review <- review_watersheds_global_final(watersheds = watersheds,
                                                sites = intl_unique,
                                                basins = big_basins_review,
                                                rivers = hydro_rivers_review %>% st_join(watersheds, left = FALSE),
                                                site_buffer_m = 2500,
                                                id_col = "rowid",
                                                review_col = "review")
```

... look good now!

Bind all together and save:
```{r}
global_watersheds <- list.files("data/intl_watersheds_refined/", full.names = TRUE) %>%
  purrr::map_dfr(~readRDS(.x) %>%
                   dplyr::mutate(rowid = as.integer(tools::file_path_sans_ext(basename(.x)))) %>%
                   dplyr::select(rowid) %>%
                   nngeo::st_remove_holes()) %>%
  left_join(., intl_unique %>% st_drop_geometry(), by = "rowid")

global_watersheds_all <- global_watersheds %>%
  right_join(.,
             intl_sites_all, by = c("Longitude", "Latitude"))

saveRDS(global_watersheds_all, "data/global_watersheds.RDS")
```
