---
title: "Global: Grabbing Geospatial Variables"
output: html_document
date: "2026-06-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = FALSE)
```

```{r libraries}
library(tidyverse)
library(sf)
library(terra)
library(elevatr)
library(lwgeom)
library(nngeo)
library(mapview)
library(leaflet)
library(units)
library(geodata)
library(arrow)
library(exactextractr)

list.files("src", full.names = TRUE) %>%
  walk(~source(.))

sf_use_s2(FALSE)
```

# 5. Grab geospatial

This section attaches all geospatial attributes to the GROWdb sites. It runs *after*
steps 1–4, which produce:

* `data/global_watersheds.RDS` and `data/conus_watersheds.RDS` (delineated watersheds),
* the reviewed/refined watersheds in `data/intl_watersheds_refined/`,
* the helper functions in `src/` (e.g. `select_flowline_global()`).

Everything below is organized as:

1. **5.1** Build the unified watershed polygons and site points used for extraction.
2. **5.2** Selected-flowline (HydroRIVERS) attributes per site.
3. **5.3** Local-basin (HydroBASINS) attributes per site.
4. **5.4** Load all geospatial source layers (rasters + vectors).
5. **5.5** Extraction helper functions.
6. **5.6** Watershed-scale extractions (`*Ws`).
7. **5.7** Site-scale / point extractions (`*Site`).
8. **5.8** Assemble the final per-sample master table.

```{r config}
site_csv             <- "data/date_lat_long_grow_KKA_05.05.26.csv"
conus_watershed_rds  <- "data/conus_watersheds.RDS"
hydrobasins_dir      <- "data/hydrobasins"
hydrorivers_gdb      <- "data/hydrorivers/HydroRIVERS_v10.gdb/HydroRIVERS_v10.gdb"

out_dir     <- "data/global_watershed_geospatial"
raster_dir  <- file.path(out_dir, "rasters")
extract_dir <- file.path(out_dir, "extracted")

walk(c(out_dir, raster_dir, extract_dir),
     ~dir.create(.x, recursive = TRUE, showWarnings = FALSE))

# Single id column used throughout extraction.
id_col <- "rowid"

# Original sample table (one row per sample). Everything is ultimately joined
# back to this so the master table is sample-level.
all_sites <- readr::read_csv(site_csv, show_col_types = FALSE)
```

Steps 1–3 load `basins` (HydroBASINS lvl 12) and `hydro_rivers` (HydroRIVERS).
Because this section can be run on its own, the guard below re-loads them only if
they are not already in the environment.

```{r load-prerequisites}
if (!exists("basins")) {
  basin_files <- list.files(hydrobasins_dir,
                            pattern   = "hybas.*lev12.*\\.shp$",
                            full.names = TRUE,
                            recursive  = TRUE)
  basins <- purrr::map(basin_files, ~sf::st_read(.x, quiet = TRUE)) %>%
    dplyr::bind_rows() %>%
    sf::st_transform(4326) %>%
    sf::st_make_valid()
}

if (!exists("hydro_rivers")) {
  hydro_rivers <- sf::st_read(hydrorivers_gdb, quiet = TRUE)
}
```

## 5.1 Unified watersheds and site points

Combine the CONUS and global watershed polygons into a single object keyed by a
fresh `rowid`. Each unique pour-point location (lat/long) appears once. The same
`rowid` keys every watershed-scale extraction below.

```{r build-watersheds}
watersheds <- readRDS("data/global_watersheds.RDS") %>%
  dplyr::select(Latitude, Longitude) %>%
  dplyr::bind_rows(readRDS(conus_watershed_rds) %>% sf::st_transform(4326) %>% dplyr::select(Latitude, Longitude)) %>%
  dplyr::distinct(Latitude, Longitude, .keep_all = TRUE) %>%
  tibble::rowid_to_column(id_col) %>%
  # Clean the watershed geometries once, up front, so every extraction below uses the
  # same valid, dateline-wrapped polygons. The cleaned object is saved for reuse.
  sf::st_make_valid() %>%
  dplyr::distinct(rowid, .keep_all = TRUE) %>%
  # claude suggestion
  sf::st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"),
                       quiet   = TRUE) %>%
  sf::st_make_valid()

saveRDS(watersheds, file.path(out_dir, "watersheds_cleaned_for_extraction.RDS"))
```

Build the matching point object: one point per unique site location, keyed by its
own `rowid`. This is the object used for every site-scale (i.e., point) extraction.

```{r build-site-points}
sites_pts <- all_sites %>%
  dplyr::distinct(Latitude, Longitude) %>%
  dplyr::arrange(dplyr::desc(Latitude), Longitude) %>%
  tibble::rowid_to_column(id_col) %>%
  sf::st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE)
```

## 5.2 Selected-flowline (HydroRIVERS) attributes

Trim HydroRIVERS to a buffer around the sites, then snap each site to its
representative flowline.

```{r select-flowlines}
sites_buf <- sites_pts %>%
  sf::st_transform(3857) %>%
  sf::st_buffer(50000) %>%
  sf::st_transform(4326)

hydro_rivers_review <- hydro_rivers[sites_buf, ]

# selected_flowlines <- select_flowline_global(flowlines = hydro_rivers_review,
#                                              sites = sites_pts)
# 
# # Drop geometry, re-attach the canonical lat/long, and save the tabular flowline
# # data. `river_attrs` is the tidy, per-site set of flowline attributes carried into
# # the master table.
# 
# final_hydrorivers_data <- selected_flowlines %>%
#   sf::st_drop_geometry() %>%
#   dplyr::select(-dplyr::any_of(c("Latitude", "Longitude"))) %>%
#   dplyr::left_join(sites_pts %>% 
#                      sf::st_drop_geometry() %>% 
#                      dplyr::select(rowid, Latitude, Longitude),
#                    by = "rowid") %>%
#   dplyr::select(-dplyr::any_of("geometry"))
# 
# saveRDS(final_hydrorivers_data, "data/final_hydrorivers_data.RDS")

final_hydrorivers_data <- readRDS("data/final_hydrorivers_data.RDS")

river_attrs <- final_hydrorivers_data %>%
  dplyr::transmute(Latitude, 
                   Longitude,
                   Status = flowline_status,
                   HYRIV_ID,
                   StreamOrder = ORD_STRA,    # Strahler stream order (analogue of NHD streamorde)
                   DischargeCMS = DIS_AV_CMS,  # mean annual discharge (m3/s)
                   MainRiv = MAIN_RIV) %>%
  dplyr::distinct(Latitude, Longitude, .keep_all = TRUE) %>%
  mutate(Status = case_when(Status == "selected_flowline" ~ "HydroRIVERS feature",
                            Status == "small_no_flowline" ~ "No HydroRIVERS feature (too small)",
                            Status == "big_estuary_or_too_large" ~ "No HydroRIVERS feature (estuary)")) %>%
  mutate(StreamOrder = case_when(Status == "No HydroRIVERS feature (too small)" ~ 1L, 
                                        Status == "No HydroRIVERS feature (estuary)" ~ 10L,
                                        TRUE ~ as.integer(StreamOrder)))
```

## 5.3 Local-basin (HydroBASINS) attributes

Each site falls within one HydroBASINS level-12 polygon. Join those basin
attributes by point-in-polygon and store the nested Pfafstetter codes (2–10).
These are the global equivalent to US HUCs.

```{r basin-attrs}
site_basin_attrs <- sf::st_join(sites_pts,
                                basins %>% dplyr::select(HYBAS_ID, PFAF_ID, SUB_AREA, UP_AREA, NEXT_DOWN, MAIN_BAS),
                                join = sf::st_within) %>%
  sf::st_drop_geometry() %>%
  mutate(pfaf_2  = substr(as.character(PFAF_ID), 1, 2),
                pfaf_4  = substr(as.character(PFAF_ID), 1, 4),
                pfaf_6  = substr(as.character(PFAF_ID), 1, 6),
                pfaf_8  = substr(as.character(PFAF_ID), 1, 8),
                pfaf_10 = substr(as.character(PFAF_ID), 1, 10)) %>%
  dplyr::select(Latitude, Longitude, HYBAS_ID, PFAF_ID,
                pfaf_2, pfaf_4, pfaf_6, pfaf_8, pfaf_10,
                SUB_AREA, UP_AREA) %>%
  dplyr::distinct(Latitude, Longitude, .keep_all = TRUE)
```

## 5.4 Load geospatial source layers

All raster and vector sources used by both the watershed- and site-scale
extractions. Download steps (`geodata::*`) are saved locally and only run once.
Other data was directly downloaded manually.

```{r load-vector-layers}
# Wastewater treatment plants (point outfalls) from HydroSHEDS:
hydrowaste <- read.csv("data/global_watershed_geospatial/hydrowaste/HydroWASTE_v10/HydroWASTE_v10.csv") %>%
  dplyr::select(WASTE_ID, LAT_OUT, LON_OUT, POP_SERVED, WASTE_DIS) %>%
  sf::st_as_sf(coords = c("LON_OUT", "LAT_OUT"), crs = 4326)

# RESOLVE ecoregions / biomes / realms.
# (hub.arcgis.com/datasets/esri::resolve-ecoregions-and-biomes/about)
resolve <- sf::st_read("data/global_watershed_geospatial/resolve/e7c09d3e-4f20-43d9-8873-60c10c28cc7b.gdb/",
                       quiet = TRUE) %>%
  dplyr::select(ECO_NAME, BIOME_NAME, REALM)

# GRanD / GDW dams and reservoirs.
# Lehner et al. 2011, Front. Ecol. Environ. 9(9): 494-502.
grand_dams <- sf::st_read("data/global_watershed_geospatial/grand/GDW_v1_0_shp/GDW_v1_0_shp/GDW_reservoirs_v1_0.shp",
                          quiet = TRUE)
```

```{r load-raster-layers}
# Population density (GPW v4.11, 2015, 30 arc-sec).
# earth.gov/ghgcenter/data-catalog/sedac-popdensity-yeargrid5yr-v4.11
gpw <- terra::rast("data/global_watershed_geospatial/population/gpw_v4_population_density_rev11_2015_30_sec_2015.tif")

# SoilGrids (0-30 cm mean): clay, sand, soil organic carbon, pH.
soil_path <- "data/global_watershed_geospatial/soilgrids"
dir.create(soil_path, showWarnings = FALSE)

# Run once; geodata caches results at soil_path.
geodata::soil_world(var = "clay", depth = 30, stat = "mean", path = soil_path)
geodata::soil_world(var = "sand", depth = 30, stat = "mean", path = soil_path)
geodata::soil_world(var = "soc", depth = 30, stat = "mean", path = soil_path)
geodata::soil_world(var = "phh2o", depth = 30, stat = "mean", path = soil_path)

soil_clay <- terra::rast(list.files(paste0(soil_path, "/soil_world/"), pattern = "clay.*30.*mean",  full.names = TRUE)[1])
soil_sand <- terra::rast(list.files(paste0(soil_path, "/soil_world/"), pattern = "sand.*30.*mean",  full.names = TRUE)[1])
soil_soc <- terra::rast(list.files(paste0(soil_path, "/soil_world/"), pattern = "soc.*30.*mean",   full.names = TRUE)[1])
soil_ph <- terra::rast(list.files(paste0(soil_path, "/soil_world/"), pattern = "phh2o.*30.*mean", full.names = TRUE)[1])

# WorldClim 2.1 climate normals (2.5 arc-min): 12 monthly layers each.
geodata::worldclim_global(var = "prec", res = 2.5, path = raster_dir)
geodata::worldclim_global(var = "tmin", res = 2.5, path = raster_dir)
geodata::worldclim_global(var = "tmax", res = 2.5, path = raster_dir)
geodata::worldclim_global(var = "tavg", res = 2.5, path = raster_dir)

clim_dir <- file.path(raster_dir, "climate", "wc2.1_2.5m")
clim_precip <- terra::rast(list.files(clim_dir, pattern = "prec", full.names = TRUE))
clim_tmin <- terra::rast(list.files(clim_dir, pattern = "tmin", full.names = TRUE))
clim_tmax <- terra::rast(list.files(clim_dir, pattern = "tmax", full.names = TRUE))
clim_tavg <- terra::rast(list.files(clim_dir, pattern = "tavg", full.names = TRUE))

# Elevation (2.5 arc-min).
geodata::elevation_global(res = 2.5, path = raster_dir)
elev_global <- terra::rast(list.files(file.path(raster_dir, "elevation", "wc2.1_2.5m"),
                                      pattern = "elev", full.names = TRUE))

# Aridity index (CGIAR; stored x10^4) and net primary productivity.
aridity <- terra::rast("data/ai_et0.tif")
npp <- terra::rast("data/npp500m.tif")
```

```{r load-landcover}
# ESA-derived landcover fractions, one raster per class.
lc_vars <- c("trees", "grassland", "shrubs", "cropland", "built", "bare",
             "snow", "water", "wetland", "mangroves", "moss")

purrr::walk(lc_vars, ~geodata::landcover(var = .x, path = raster_dir))

landcover_files <- list.files(file.path(raster_dir, "landuse"),
                              full.names = TRUE,
                              pattern = "\\.tif$|\\.grd$|\\.vrt$|\\.nc$",
                              recursive = FALSE)

landcover <- purrr::map(landcover_files, terra::rast)

# Name each layer by its matched class; fall back to the filename stem so no
# layer is silently dropped when the filename doesn't contain an expected var.
names(landcover) <- purrr::map_chr(landcover_files, function(x) {
  stem <- tools::file_path_sans_ext(basename(x))
  matched_var <- lc_vars[stringr::str_detect(stem, lc_vars)]
  if (length(matched_var) == 0) stem else matched_var[1]
})
```

## 5.5 Extraction helpers

`exactextractr` is used for the polygon (watershed) extractions because it is
area-weighted; `terra::extract` is used for the point (site) extractions and for
the count/density vector overlays. The helpers below keep both consistent and
always carry the `id_col` so results can be joined safely (never by just by row 
position).

```{r helpers-polygon}
# Area-weighted mean of every layer of `r` per polygon.
extract_mean_by_polygon <- function(r, polygons, id_col, prefix = NULL) {
  stopifnot(inherits(r, "SpatRaster"), inherits(polygons, "sf"), id_col %in% names(polygons))
  
  if (!is.null(prefix)) names(r) <- paste0(prefix, "_", names(r))
  
  vals <- exactextractr::exact_extract(r, polygons, fun = "mean", progress = TRUE) %>%
    tibble::as_tibble()
  
  dplyr::bind_cols(polygons %>% 
                     sf::st_drop_geometry() %>% 
                     dplyr::select(rowid),
                   vals)
}

# Elevation summary statistics per polygon.
extract_elevation_by_polygon <- function(r, polygons, id_col) {
  stopifnot(inherits(r, "SpatRaster"), inherits(polygons, "sf"), id_col %in% names(polygons))
  
  vals <- exactextractr::exact_extract(r, polygons, 
                                       fun = c("mean", "min", "max", "stdev"), 
                                       progress = TRUE) %>%
    tibble::as_tibble()
  
  names(vals) <- c("elevation_mean_m", "elevation_min_m", "elevation_max_m", "elevation_sd_m")
  
  dplyr::bind_cols(
    polygons %>% sf::st_drop_geometry() %>% dplyr::select(rowid),
    vals
  )
}

# Mean fraction of each landcover class per polygon, returned wide.
extract_landcover_attributes <- function(watersheds, landcover,
                                         id_col = "rowid", prefix = "lc") {
  stopifnot(inherits(watersheds, "sf"), id_col %in% names(watersheds), is.list(landcover))
  
  watersheds <- watersheds %>%
    sf::st_make_valid() %>%
    sf::st_collection_extract("POLYGON") %>%
    sf::st_transform(4326)
  watersheds <- watersheds[!sf::st_is_empty(watersheds), ]
  
  if (is.null(names(landcover)) || any(names(landcover) == "")) {
    names(landcover) <- paste0("landcover_", seq_along(landcover))
  }
  
  landcover_summary <- purrr::imap_dfr(landcover, function(r, landcover_name) {
    message("Extracting landcover layer: ", landcover_name)
    
    # Only coerce if not already a SpatRaster.
    if (!inherits(r, "SpatRaster")) r <- terra::rast(r)
    if (terra::nlyr(r) > 1) r <- r[[1]]
    
    extracted <- exactextractr::exact_extract(r, watersheds, fun = "mean", progress = TRUE)
    
    tibble::tibble("{id_col}" := watersheds[[id_col]],
                   landcover_class = landcover_name,
                   mean_fraction = extracted)
  })
  
  landcover_summary %>%
    mutate(landcover_class = landcover_class %>%
                    stringr::str_replace_all("[^A-Za-z0-9_]+", "_") %>%
                    stringr::str_replace_all("_+$", "") %>%
                    stringr::str_to_lower(),
                  attribute_name = paste0(prefix, "_", landcover_class, "_mean_fraction")) %>%
    dplyr::select(rowid, attribute_name, mean_fraction) %>%
    dplyr::distinct(dplyr::across(rowid), attribute_name, .keep_all = TRUE) %>%
    tidyr::pivot_wider(names_from  = attribute_name,
                       values_from = mean_fraction,
                       values_fn = mean)
}
```

```{r helpers-point}
# Raster cell value(s) at each site point. Every non-ID layer is returned, with
# names prefixed. Works for single- and multi-layer rasters.
terra_at_points <- function(r, points, id_col, prefix) {
  pts <- sf::st_transform(points, terra::crs(r))
  vals <- terra::extract(r, terra::vect(pts)) %>% dplyr::select(-ID)
  if (!is.null(prefix)) names(vals) <- paste0(prefix, "_", names(vals))
  dplyr::bind_cols(points %>% sf::st_drop_geometry() %>% dplyr::select(rowid),
                   tibble::as_tibble(vals))
}
```

## 5.6 Watershed-scale extractions

Each block produces a tibble keyed by the watershed `rowid`. They are reduced
together at the end of this section. `*Ws` suffixes mark watershed-scale values.

### Climate, elevation, landcover (area-weighted)

```{r ws-climate}
prec_extract <- extract_mean_by_polygon(clim_precip, watersheds, id_col)
arrow::write_parquet(prec_extract, file.path(extract_dir, "watershed_worldclim_precip.parquet"))

tavg_extract <- extract_mean_by_polygon(clim_tavg, watersheds, id_col)
arrow::write_parquet(tavg_extract, file.path(extract_dir, "watershed_worldclim_tavg.parquet"))

tmax_extract <- extract_mean_by_polygon(clim_tmax, watersheds, id_col)
arrow::write_parquet(tmax_extract, file.path(extract_dir, "watershed_worldclim_tmax.parquet"))

tmin_extract <- extract_mean_by_polygon(clim_tmin, watersheds, id_col)
arrow::write_parquet(tmin_extract, file.path(extract_dir, "watershed_worldclim_tmin.parquet"))
```

```{r ws-elevation}
elev_extract <- extract_elevation_by_polygon(elev_global, watersheds, id_col)
arrow::write_parquet(elev_extract, file.path(extract_dir, "watershed_elevation.parquet"))
```

```{r ws-landcover}
lc_extract <- extract_landcover_attributes(watersheds = watersheds,
                                           landcover = landcover,
                                           id_col = id_col,
                                           prefix = "lc")
arrow::write_parquet(lc_extract, file.path(extract_dir, "watershed_landcover.parquet"))

purrr::map_dbl(landcover, ~ as.numeric(terra::global(.x, "max", na.rm = TRUE)))

# sanity-check the classes sum appropriately per watershed
# lc_total_pct should sit at ~100 for each watershed...
lc_sanity_check <- lc_extract %>%
  mutate(dplyr::across(-rowid, ~ .x * 100)) %>%
  mutate(lc_total_pct = rowSums(dplyr::across(-rowid), na.rm = TRUE))
# ... they do!
```

### Soil, population, aridity, NPP ({terra} means)

```{r ws-soil}
# Stack the four soil layers (same grid, from soil_world()) so the polygon
# overlay is computed once, and use exactextractr (area-weighted) instead of
# terra::extract. Output is identical in shape to before: rowid + the 4 *Ws cols.
soil_stack <- c(soil_clay, soil_sand, soil_soc, soil_ph)
names(soil_stack) <- c("ClayWs", "SandWs", "SocWs", "PhWs")

ws_soil_crs <- sf::st_transform(watersheds, terra::crs(soil_stack))

soil_ws <- exactextractr::exact_extract(soil_stack, ws_soil_crs, 
                                        fun = "mean", progress = TRUE) %>%
  tibble::as_tibble() %>%
  dplyr::rename_with(~ stringr::str_remove(.x, "^mean\\.")) %>%
  dplyr::bind_cols(watersheds %>% 
                     sf::st_drop_geometry() %>% 
                     dplyr::select(rowid),
                   .) %>%
  dplyr::relocate(rowid)
```

```{r ws-population}
pop_ws <- extract_mean_by_polygon(gpw, watersheds, id_col) %>%
  dplyr::rename(PopDenWs = 2)
```

```{r ws-aridity-npp}
aridity_ws <- extract_mean_by_polygon(aridity, watersheds, id_col) %>%
  dplyr::rename(AridityWs = 2) %>%
  mutate(AridityWs = AridityWs / 10000)   # CGIAR aridity index stored x10^4

npp_ws <- extract_mean_by_polygon(npp, watersheds, id_col) %>%
  dplyr::rename(NPPWs = 2)
```

### Wastewater treatment plants and dams (point-in-polygon counts)

```{r ws-wwtp}
wwtp_ws <- sf::st_join(hydrowaste,
                       watersheds %>% dplyr::select(rowid),
                       join = sf::st_within) %>%
  sf::st_drop_geometry() %>%
  dplyr::filter(!is.na(rowid)) %>%
  dplyr::group_by(rowid) %>%
  dplyr::summarize(WWTPCountWs = dplyr::n(),
                   WWTPPopWs = sum(POP_SERVED, na.rm = TRUE),
                   .groups = "drop")
```

```{r ws-dams}
dam_ws <- sf::st_join(grand_dams,
                      watersheds %>% dplyr::select(rowid),
                      join = sf::st_within) %>%
  sf::st_drop_geometry() %>%
  dplyr::filter(!is.na(rowid)) %>%
  dplyr::group_by(rowid) %>%
  dplyr::summarize(DamCountWs = dplyr::n(),
                   DamStorWs = sum(CAP_MCM, na.rm = TRUE), # total storage (million m3)
                   .groups = "drop")
```

### Dominant ecoregion (largest overlap area)

```{r ws-ecoregion}
eco_ws <- sf::st_intersection(watersheds %>% 
                                dplyr::select(rowid) %>% 
                                sf::st_transform(sf::st_crs(resolve)),
                              resolve) %>%
  mutate(overlap_area = sf::st_area(.)) %>%
  sf::st_drop_geometry() %>%
  dplyr::group_by(rowid) %>%
  dplyr::slice_max(overlap_area, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::select(rowid,
                EcoregionWs = ECO_NAME,
                BiomeWs = BIOME_NAME,
                RealmWs = REALM)
```

### Assemble watershed attributes

Reduce-join every watershed-scale block by `rowid`, attach the canonical lat/long,
and add the area-derived densities. Counts default to 0 where no feature was found.

```{r ws-assemble}
ws_area <- watersheds %>%
  sf::st_drop_geometry() %>%
  dplyr::transmute(rowid, 
                   Latitude, 
                   Longitude,
                   WsAreaSqKm = as.numeric(sf::st_area(watersheds)) / 1e6)

watershed_attributes <- list(ws_area,
                             elev_extract, prec_extract, tmin_extract, tmax_extract, tavg_extract,
                             lc_extract, soil_ws, pop_ws, aridity_ws, npp_ws,
                             wwtp_ws, dam_ws, eco_ws) %>%
  purrr::reduce(dplyr::left_join, by = id_col) %>%
  mutate(WWTPCountWs = tidyr::replace_na(WWTPCountWs, 0),
                WWTPPopWs = tidyr::replace_na(WWTPPopWs, 0),
                DamCountWs = tidyr::replace_na(DamCountWs, 0),
                DamStorWs = tidyr::replace_na(DamStorWs, 0),
                WWTPDensWs = WWTPCountWs / WsAreaSqKm,
                DamDensWs = DamCountWs / WsAreaSqKm)

# Watershed-scale attribute table (one row per watershed).
readr::write_csv(watershed_attributes, file.path(extract_dir, "watershed_all_attributes.csv"))
```

## 5.7 Site-scale (point) extractions

The point equivalent of Section 5.6 for all vars that this applies to: instead
of aggregating over a watershed polygon, take the raster cell value at each site 
location and the vector feature each site falls within. 
`*Site` suffixes mark site-scale values.

### Climate and elevation at the point

```{r site-climate-elev}
prec_site <- terra_at_points(clim_precip, sites_pts, id_col, prefix = "precSite")
tmin_site <- terra_at_points(clim_tmin, sites_pts, id_col, prefix = "tminSite")
tmax_site <- terra_at_points(clim_tmax, sites_pts, id_col, prefix = "tmaxSite")
tavg_site <- terra_at_points(clim_tavg, sites_pts, id_col, prefix = "tavgSite")

elev_site <- terra_at_points(elev_global, sites_pts, id_col, prefix = NULL) %>%
  dplyr::rename(ElevationSite = 2)
```

### Soil at the point

```{r site-soil}
soil_site <- list(terra_at_points(soil_clay, sites_pts, id_col, NULL) %>% dplyr::rename(ClaySite = 2),
                  terra_at_points(soil_sand, sites_pts, id_col, NULL) %>% dplyr::rename(SandSite = 2),
                  terra_at_points(soil_soc, sites_pts, id_col, NULL) %>% dplyr::rename(SocSite = 2),
                  terra_at_points(soil_ph, sites_pts, id_col, NULL) %>% dplyr::rename(PhSite = 2)) %>%
  purrr::reduce(dplyr::left_join, by = id_col)
```

### Population, aridity, NPP at the point

```{r site-pop-aridity-npp}
pop_site <- terra_at_points(gpw, sites_pts, id_col, NULL) %>%
  dplyr::rename(PopDenSite = 2)

aridity_site <- terra_at_points(aridity, sites_pts, id_col, NULL) %>%
  dplyr::rename(AriditySite = 2) %>%
  mutate(AriditySite = AriditySite / 10000)

npp_site <- terra_at_points(npp, sites_pts, id_col, NULL) %>%
  dplyr::rename(NPPSite = 2)
```

### Landcover cell fractions at the point

```{r site-landcover}
lc_site <- purrr::imap(landcover, function(r, nm) {
  if (!inherits(r, "SpatRaster")) r <- terra::rast(r)
  if (terra::nlyr(r) > 1) r <- r[[1]]
  nm_clean <- nm %>%
    stringr::str_replace_all("[^A-Za-z0-9_]+", "_") %>%
    stringr::str_replace_all("_+$", "") %>%
    stringr::str_to_lower()
  terra_at_points(r, sites_pts, id_col, NULL) %>%
    dplyr::rename("lc_{nm_clean}_site_fraction" := 2)
}) %>%
  purrr::reduce(dplyr::left_join, by = id_col)
```

### Ecoregion at the point

```{r site-ecoregion}
eco_site <- sf::st_join(sites_pts %>% sf::st_transform(sf::st_crs(resolve)),
                        resolve,
                        join = sf::st_within,
                        left = TRUE) %>%
  sf::st_drop_geometry() %>%
  dplyr::group_by(rowid) %>%
  dplyr::slice(1) %>%
  dplyr::ungroup() %>%
  dplyr::select(rowid,
                EcoregionSite = ECO_NAME,
                BiomeSite = BIOME_NAME,
                RealmSite = REALM)
```

### Assemble site attributes

```{r site-assemble}
site_latlong <- sites_pts %>%
  sf::st_drop_geometry() %>%
  dplyr::select(rowid, Latitude, Longitude)

site_attributes <- list(site_latlong,
                        elev_site, prec_site, tmin_site, tmax_site, tavg_site,
                        lc_site, soil_site, pop_site, aridity_site, npp_site, eco_site) %>%
  purrr::reduce(dplyr::left_join, by = id_col)

readr::write_csv(site_attributes,
                 file.path(extract_dir, "site_all_attributes.csv"))
```

## 5.8 Final per-sample master table

Join everything back to the original sample table so every GROWdb sample carries
its basin, flowline, watershed-scale, and site-scale attributes. Joins are by
`Latitude`/`Longitude` (the canonical location key shared by all blocks); the
internal `rowid`s are dropped first so the two numbering schemes never collide.

```{r master-table}
ws_join <- watershed_attributes %>% dplyr::select(-rowid)
site_join <- site_attributes %>% dplyr::select(-rowid)

site_master_table <- all_sites %>%
  dplyr::left_join(site_basin_attrs, by = c("Latitude", "Longitude")) %>%
  dplyr::left_join(river_attrs, by = c("Latitude", "Longitude")) %>%
  dplyr::left_join(ws_join, by = c("Latitude", "Longitude")) %>%
  dplyr::left_join(site_join, by = c("Latitude", "Longitude"))

saveRDS(site_master_table, file.path(extract_dir, "site_global_master_table.RDS"))
readr::write_csv(site_master_table, file.path(extract_dir, "site_global_master_table.csv"))
```
