---
title: 'Comparing CONUS + Global Products'
output: html_document
date: "2026-06-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 9, fig.height = 6, dpi = 110
)

# ---- packages -------------------------------------------------------------
# Only widely-available CRAN packages are used.
suppressPackageStartupMessages({
  library(readr)
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(stringr)
  library(ggplot2)
  library(tibble)
  library(knitr)
})

# kableExtra is optional; fall back to plain kable if absent.
have_kableExtra <- requireNamespace("kableExtra", quietly = TRUE)

nice_table <- function(df, caption = NULL, digits = 3) {
  k <- knitr::kable(df, caption = caption, digits = digits, format = "html")
  if (have_kableExtra) {
    k <- kableExtra::kable_styling(
      k, bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE, position = "left"
    )
  }
  k
}

theme_set(theme_minimal(base_size = 12))
```

# Purpose

`GROWdb` is the **CONUS-only** product. It was built largely from
U.S.-specific source layers — NHDPlus V2 (flowlines, value-added
attributes, EROM flow) and StreamCat (watershed-accumulated landscape
metrics), plus USGS gage data. The **global master table**
(`site_global_master_table.csv`) re-derives an analogous set of
descriptors for the same sampling sites using **global** source layers —
HydroSHEDS / HydroBASINS / HydroRIVERS / HydroATLAS, WorldClim 2.1
climatology, Copernicus/ESA land cover, SoilGrids-style soils, and so on.

Because the two products mostly draw on *different underlying datasets*,
identical numbers are not expected everywhere. The goal here is narrower
and more important: **for variables that represent the same physical
concept, confirm the two workflows land in the same place.** Strong
agreement on overlapping variables is evidence the global pipeline
reproduces the established CONUS pipeline; systematic disagreement flags
either a units/definition mismatch or a genuine methodological divergence
worth investigating.

The comparison is organized as:

1. **Load & join** the two tables on sample identity.
2. **Positive controls** — variables that were carried over from the *same*
   source in both products. These should agree essentially perfectly and
   validate the join itself.
3. **Crosswalk** of overlapping-but-different-source variables, with units
   reconciled.
4. **Quantitative concordance** (n, Pearson r, Spearman ρ, bias, RMSE) per
   variable.
5. **Visual concordance** — 1:1 scatter and Bland–Altman difference plots.
6. **Land cover** — a dedicated section reconciling Copernicus fractions
   against aggregated NLCD classes.
7. **Coverage gaps & non-comparable fields** — what could *not* be compared
   and why.

# Load and join

```{r load}
raw <- read_csv("data/date_lat_long_grow_KKA_05.05.26.csv")
grow  <- read_csv("data/0-shared/GROWdb_20260528.csv", show_col_types = FALSE)
global <- read_csv("data/global_watershed_geospatial/extracted/site_global_master_table.csv",  show_col_types = FALSE)

cat("Master table :", nrow(raw),  "rows x", ncol(raw),  "cols\n")
cat("Global table :", nrow(global),  "rows x", ncol(global),  "cols\n")
cat("CONUS table:", nrow(grow), "rows x", ncol(grow), "cols\n")
```

The two tables share a sample identifier (`sample_name` in the global
table, `SampleName` in GROWdb). We join on that. We do **not** join on
coordinates, because the two products carry slightly different
lat/long precision and rounding, which would silently drop matches.

```{r join}
# Standardise the join key.
global  <- global  %>% rename(sample_id = sample_name)
grow <- grow %>% rename(sample_id = SampleName)

j <- inner_join(global, grow, by = "sample_id", suffix = c("_global", "_grow"))

n_global_only  <- length(setdiff(global$sample_id,  grow$sample_id))
n_grow_only <- length(setdiff(grow$sample_id, global$sample_id))

cat("Matched site-pairs (inner join):", nrow(j), "\n")
cat("In global only :", n_global_only, "\n")
cat("In GROWdb only :", n_grow_only, "\n")
```

```{r}
rows_with_na <- global %>%
  filter(if_any(everything(), is.na))

cat("Rows with at least one NA:", nrow(rows_with_na), "\n")

rows_with_na

na_by_row <- global %>%
  mutate(
    na_cols = pmap_chr(
      across(everything()),
      ~ names(global)[map_lgl(list(...), is.na)] %>% paste(collapse = ", ")
    )
  ) %>%
  filter(na_cols != "")

na_by_row %>%
  select(sample_id, na_cols)

na_summary <- global %>%
  summarise(across(everything(), ~ sum(is.na(.x)))) %>%
  pivot_longer(
    everything(),
    names_to = "column",
    values_to = "n_na"
  ) %>%
  filter(n_na > 0) %>%
  arrange(desc(n_na))

na_summary
```

```{r derived}
# ---- Derived / unit-reconciled columns ------------------------------------
# Global climate is delivered as 12 WorldClim monthly layers; collapse to
# annual to match GROWdb's annual fields.
prec_site_cols <- grep("^precSite_wc2\\.1_2\\.5m_prec_", names(j), value = TRUE)
tavg_site_cols <- grep("^tavgSite_wc2\\.1_2\\.5m_tavg_", names(j), value = TRUE)

j <- j %>%
  mutate(
    # Annual site precip (mm) = sum of 12 monthly; annual site temp (C) = mean.
    global_precSite_annual = if (length(prec_site_cols)) rowSums(across(all_of(prec_site_cols)), na.rm = FALSE) else NA_real_,
    global_tavgSite_annual = if (length(tavg_site_cols)) rowMeans(across(all_of(tavg_site_cols)), na.rm = FALSE) else NA_real_,

    # GROWdb discharge is in cubic feet/sec; convert to CMS to match global.
    grow_qe_ma_cms = if ("qe_ma" %in% names(j)) qe_ma * 0.0283168 else NA_real_,
    grow_CFS_cms   = if ("CFS"   %in% names(j)) CFS   * 0.0283168 else NA_real_,

    # NHDPlus smoothed elevations are in centimetres; convert to metres.
    grow_minelev_m = if ("minelevsmo" %in% names(j)) minelevsmo / 100 else NA_real_
  )
```

# Positive controls: variables sharing a source

A handful of fields were populated from the **same** source dataset in
both products (e.g. aridity index, net primary productivity). They are the
control group: if the join is correct and no transformation crept in, they
should be effectively identical (r ≈ 1, bias ≈ 0). Any failure here would
mean the rest of the comparison can't be trusted.

```{r compare-fn}
# ---- Core concordance helper ----------------------------------------------
compare_pair <- function(data, x_col, y_col) {
  if (!all(c(x_col, y_col) %in% names(data))) {
    return(tibble(n = 0L, pearson_r = NA, spearman_rho = NA,
                  bias_grow_minus_global = NA, rmse = NA,
                  global_median = NA, grow_median = NA))
  }
  x <- suppressWarnings(as.numeric(data[[x_col]]))
  y <- suppressWarnings(as.numeric(data[[y_col]]))
  ok <- is.finite(x) & is.finite(y)
  x <- x[ok]; y <- y[ok]
  n <- length(x)
  if (n < 3) {
    return(tibble(n = n, pearson_r = NA, spearman_rho = NA,
                  bias_grow_minus_global = NA, rmse = NA,
                  global_median = NA, grow_median = NA))
  }
  tibble(
    n = n,
    pearson_r           = suppressWarnings(cor(x, y, method = "pearson")),
    spearman_rho        = suppressWarnings(cor(x, y, method = "spearman")),
    bias_grow_minus_global = mean(y - x),
    rmse                = sqrt(mean((y - x)^2)),
    global_median          = median(x),
    grow_median         = median(y)
  )
}

# Build a results table from a spec tibble:
#   concept | category | x_col (global) | y_col (grow) | unit | note
run_specs <- function(data, specs) {
  specs %>%
    mutate(stats = pmap(list(x_col, y_col), ~ compare_pair(data, ..1, ..2))) %>%
    unnest(stats) %>%
    select(category, concept, unit, n, pearson_r, spearman_rho,
           bias_grow_minus_global, rmse, global_median, grow_median, note)
}
```

```{r controls}
control_specs <- tribble(
  ~concept,                 ~category,  ~x_col,            ~y_col,             ~unit,        ~note,
  "Aridity (watershed)",    "Control",  "AridityWs_global",   "AridityWs_grow",   "index",      "Same source layer in both products",
  "Aridity (site)",         "Control",  "AriditySite_global", "AriditySite_grow", "index",      "Same source layer in both products",
  "NPP (watershed)",        "Control",  "NPPWs_global",       "NPPWs_grow",       "gC/m^2/yr",  "Same source layer in both products",
  "NPP (site)",             "Control",  "NPPSite_global",     "NPPSite_grow",     "gC/m^2/yr",  "Same source layer in both products"
)

control_res <- run_specs(j, control_specs)
kableExtra::kable(control_res %>% select(-category),
           caption = "Positive controls (shared-source variables). Expect r ~ 1.")
```

> **Read this first.** If the control rows above do not show r ≈ 1 and
> bias ≈ 0, stop — the join or a unit transform is wrong and the
> downstream comparisons are not interpretable. If they do, the join is
> sound and the matched-site values are aligned.

# Crosswalk of overlapping variables

The table below is the conceptual map between the two products for
variables that come from **different** source datasets but describe the
same quantity. Units are reconciled to a common basis in the `x_col`/`y_col`
referenced by the analysis.

```{r crosswalk-doc, echo=FALSE}
crosswalk_doc <- tribble(
  ~Concept,                  ~`Global source`,                 ~`GROWdb source`,                 ~`Unit reconciliation`,
  "Watershed / drainage area","HydroSHEDS upstream area",       "NHDPlus V2 totdasqkm",           "both km^2 (none)",
  "Stream order",            "HydroRIVERS Strahler",            "NHDPlus streamorde (Strahler)",  "none",
  "Mean annual discharge",   "HydroATLAS DischargeCMS",         "NHDPlus EROM qe_ma",             "GROW CFS -> CMS (x0.0283168)",
  "Discharge (gage)",        "HydroATLAS DischargeCMS",         "USGS gage CFS",                  "GROW CFS -> CMS (x0.0283168)",
  "Site elevation",          "DEM ElevationSite (m)",           "NHDPlus minelevsmo (cm)",        "GROW cm -> m (/100); reach vs point",
  "Site precip (annual)",    "WorldClim 2.1 (12 monthly mm)",   "StreamCat/Daymet PrecipSite",    "global sum(12 months) -> mm/yr",
  "Site temp (annual)",      "WorldClim 2.1 (12 monthly degC)", "StreamCat/Daymet TmeanSite",     "global mean(12 months) -> degC",
  "Watershed clay",          "SoilGrids ClayWs",                "StreamCat ClayWs",               "both %",
  "Watershed sand",          "SoilGrids SandWs",                "StreamCat SandWs",               "both %",
  "Watershed dam density",   "GOODD/GRanD DamDensWs",           "StreamCat (NID) DamDensWs",      "different inventories",
  "Land cover (watershed)",  "Copernicus/ESA fractions (0-1)",  "NLCD 2016 class % (StreamCat)",  "fraction*100 vs aggregated NLCD %"
)
nice_table(crosswalk_doc, caption = "Conceptual crosswalk between the two products.")
```

```{r overlap-specs}
overlap_specs <- tribble(
  ~concept,                 ~category,       ~x_col,                  ~y_col,             ~unit,        ~note,
  "Watershed area",         "Delineation",   "WsAreaSqKm_global",        "totdasqkm",        "km^2",       "HydroSHEDS vs NHDPlus totdasqkm",
  "Stream order",           "Network",       "StreamOrder",           "streamorde",       "Strahler",   "HydroRIVERS vs NHDPlus",
  "Discharge (modeled)",    "Hydrology",     "DischargeCMS",          "grow_qe_ma_cms",   "m^3/s",      "HydroATLAS vs NHDPlus EROM",
  "Discharge (gage)",       "Hydrology",     "DischargeCMS",          "grow_CFS_cms",     "m^3/s",      "HydroATLAS vs USGS gage",
  "Site elevation",         "Terrain",       "ElevationSite",         "grow_minelev_m",   "m",          "Point DEM vs NHDPlus reach min (cm->m)",
  "Site precip (annual)",   "Climate",       "global_precSite_annual",   "PrecipSite",       "mm/yr",      "WorldClim vs Daymet/StreamCat",
  "Site temp (annual)",     "Climate",       "global_tavgSite_annual",   "TmeanSite",        "degC",       "WorldClim vs Daymet/StreamCat",
  "Watershed clay",         "Soil",          "ClayWs_global",            "ClayWs_grow",      "%",          "SoilGrids vs StreamCat",
  "Watershed sand",         "Soil",          "SandWs_global",            "SandWs_grow",      "%",          "SoilGrids vs StreamCat",
  "Watershed dam density",  "Human impact",  "DamDensWs_global",         "DamDensWs_grow",   "dams/km^2",  "GOODD/GRanD vs NID; divergence expected"
)

overlap_res <- run_specs(j, overlap_specs)
nice_table(overlap_res %>% select(-category),
           caption = "Concordance of overlapping (different-source) variables.")
```

## Concordance summary, ranked

```{r ranked}
all_res <- bind_rows(control_res, overlap_res) %>%
  arrange(desc(pearson_r))

# Classify agreement into interpretable tiers.
tiered <- all_res %>%
  mutate(agreement = case_when(
    is.na(pearson_r)        ~ "no data",
    pearson_r >= 0.98       ~ "near-identical",
    pearson_r >= 0.90       ~ "strong",
    pearson_r >= 0.75       ~ "moderate",
    TRUE                    ~ "weak / divergent"
  )) %>%
  select(category, concept, agreement, n, pearson_r, spearman_rho,
         bias_grow_minus_global, rmse)

nice_table(tiered, caption = "All comparisons ranked by Pearson correlation, with agreement tier.")
```

```{r tier-counts, echo=FALSE}
tier_summary <- tiered %>%
  filter(!is.na(pearson_r)) %>%
  count(agreement) %>%
  arrange(match(agreement, c("near-identical","strong","moderate","weak / divergent")))
nice_table(tier_summary, caption = "Count of variables by agreement tier.")
```

# Visual concordance

## 1:1 scatter plots

Each panel plots the GROWdb value (y) against the global value (x) on a
shared physical scale, with the dashed 1:1 line. Points hugging the line
mean the two workflows agree; a consistent slope away from the line means a
scale/units issue; scatter means methodological divergence.

```{r scatter-data}
# Long tidy frame of matched values for every overlapping concept.
make_pairs <- function(data, specs) {
  specs %>%
    pmap_dfr(function(concept, category, x_col, y_col, unit, note) {
      if (!all(c(x_col, y_col) %in% names(data))) return(tibble())
      tibble(
        concept  = concept,
        category = category,
        unit     = unit,
        global      = suppressWarnings(as.numeric(data[[x_col]])),
        grow     = suppressWarnings(as.numeric(data[[y_col]]))
      )
    }) %>%
    filter(is.finite(global), is.finite(grow))
}

pairs_long <- make_pairs(j, overlap_specs)
```

```{r scatter-plot, fig.height=9}
ggplot(pairs_long, aes(global, grow)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(alpha = 0.5, size = 1.3, colour = "#2c7fb8") +
  facet_wrap(~ concept, scales = "free", ncol = 3) +
  labs(
    title = "GROWdb vs. Global master table — matched sites",
    subtitle = "Dashed line = 1:1. Closeness to the line indicates workflow concordance.",
    x = "Global master table value",
    y = "GROWdb (CONUS) value"
  ) +
  theme(strip.text = element_text(face = "bold", size = 9))
```

Several of these quantities (area, discharge) span orders of magnitude, so
a log–log view separates the high-agreement structure from the few large
basins that dominate a linear axis.

```{r scatter-log, fig.height=4.5}
log_concepts <- c("Watershed area", "Discharge (modeled)", "Discharge (gage)")
pairs_long %>%
  filter(concept %in% log_concepts, global > 0, grow > 0) %>%
  ggplot(aes(global, grow)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(alpha = 0.5, size = 1.3, colour = "#d95f0e") +
  scale_x_log10() + scale_y_log10() +
  facet_wrap(~ concept, scales = "free", ncol = 3) +
  labs(title = "Log–log concordance for magnitude-spanning variables",
       x = "Global value (log10)", y = "GROWdb value (log10)")
```

## Bland–Altman (difference) plots

For the well-matched variables, a Bland–Altman view (difference vs. mean)
makes any *systematic* bias or magnitude-dependent drift visible in a way a
correlation cannot.

```{r bland-altman, fig.height=6}
ba_concepts <- c("Site precip (annual)", "Site temp (annual)",
                 "Watershed clay", "Stream order")
pairs_long %>%
  filter(concept %in% ba_concepts) %>%
  mutate(mean_val = (global + grow) / 2,
         diff_val = grow - global) %>%
  ggplot(aes(mean_val, diff_val)) +
  geom_hline(yintercept = 0, colour = "grey40") +
  geom_point(alpha = 0.5, size = 1.3, colour = "#31a354") +
  geom_smooth(method = "loess", se = FALSE, colour = "#de2d26", linewidth = 0.7) +
  facet_wrap(~ concept, scales = "free", ncol = 2) +
  labs(title = "Bland–Altman: GROWdb − Global vs. mean of the two",
       subtitle = "Flat red line near zero = no systematic bias.",
       x = "Mean of the two products", y = "Difference (GROWdb − Global)")
```

# Land cover: Copernicus fractions vs. NLCD classes

Land cover is the clearest case of *different products measuring the same
thing*: the global table uses Copernicus/ESA continuous cover fractions
(0–1), while GROWdb uses NLCD 2016 class percentages from StreamCat. To
compare, NLCD classes are aggregated into the broad groups the Copernicus
fractions represent, and the global fractions are rescaled to percent.

```{r landcover}
# NLCD classes are aggregated with rowSums(across(any_of(...))) so that any
# class column absent from a given GROWdb export is skipped rather than erroring.
lc <- j %>%
  transmute(
    sample_id = sample_id,

    # ---- Forest ----
    global_forest  = 100 * `lc_trees_mean_fraction`,
    grow_forest = rowSums(across(any_of(c("PctDecid2016Ws","PctConif2016Ws","PctMxFst2016Ws")))),

    # ---- Cropland (incl. pasture/hay) ----
    global_crop    = 100 * `lc_cropland_mean_fraction`,
    grow_crop   = rowSums(across(any_of(c("PctCrop2016Ws","PctHay2016Ws")))),

    # ---- Grass + shrub ----
    global_grshrub = 100 * (`lc_grassland_mean_fraction` + `lc_shrubs_mean_fraction`),
    grow_grshrub= rowSums(across(any_of(c("PctGrs2016Ws","PctShrb2016Ws")))),

    # ---- Developed / built ----
    global_built   = 100 * `lc_built_mean_fraction`,
    grow_built  = rowSums(across(any_of(c("PctUrbOp2016Ws","PctUrbLo2016Ws","PctUrbMd2016Ws","PctUrbHi2016Ws")))),

    # ---- Open water ----
    global_water   = 100 * `lc_water_mean_fraction`,
    grow_water  = rowSums(across(any_of(c("PctOw2016Ws")))),

    # ---- Wetland ----
    global_wetland = 100 * `lc_wetland_mean_fraction`,
    grow_wetland= rowSums(across(any_of(c("PctWdWet2016Ws","PctHbWet2016Ws"))))
  )

lc_specs <- tribble(
  ~concept,        ~category,    ~x_col,        ~y_col,         ~unit, ~note,
  "Forest %",      "Land cover", "global_forest",  "grow_forest",  "%",   "Copernicus trees vs NLCD forest classes",
  "Cropland %",    "Land cover", "global_crop",    "grow_crop",    "%",   "Copernicus cropland vs NLCD crop+hay",
  "Grass+shrub %", "Land cover", "global_grshrub", "grow_grshrub", "%",   "Copernicus grass+shrub vs NLCD grass+shrub",
  "Developed %",   "Land cover", "global_built",   "grow_built",   "%",   "Copernicus built vs NLCD urban classes",
  "Open water %",  "Land cover", "global_water",   "grow_water",   "%",   "Copernicus water vs NLCD open water",
  "Wetland %",     "Land cover", "global_wetland", "grow_wetland", "%",   "Copernicus wetland vs NLCD woody+herb wetland"
)

lc_res <- run_specs(lc, lc_specs)
nice_table(lc_res %>% select(-category),
           caption = "Land cover concordance: Copernicus fractions vs aggregated NLCD 2016.")
```

```{r landcover-plot, fig.height=6}
lc_pairs <- make_pairs(lc, lc_specs)
ggplot(lc_pairs, aes(global, grow)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(alpha = 0.5, size = 1.3, colour = "#756bb1") +
  facet_wrap(~ concept, scales = "free", ncol = 3) +
  labs(title = "Land cover: Copernicus (global) vs NLCD 2016 (GROWdb)",
       subtitle = "Watershed-aggregated %. Dashed line = 1:1.",
       x = "Global (Copernicus) %", y = "GROWdb (NLCD) %")
```

> Land cover agreement is expected to be *good but not perfect*: the two
> products use different sensors, epochs, and class definitions. The useful
> question is whether the **rank ordering and broad magnitude** of cover
> types are preserved (high Spearman ρ), not whether they match to the
> percentage point.

# Coverage gaps and non-comparable fields

Not every conceptually-shared field could be compared. These are documented
so the gaps are explicit rather than silent.

```{r gaps, echo=FALSE}
# Check which GROW fields are empty for the matched rows.
check_cols <- c("WsAreaSqKm_grow", "PrecipWs", "TmeanWs", "TmaxWs", "TminWs",
                "CatAreaSqKm", "CFS", "gage_id")
gap_tbl <- tibble(
  `GROWdb field` = check_cols
) %>%
  mutate(
    present = `GROWdb field` %in% names(j),
    n_nonNA = map_int(`GROWdb field`, ~ if (.x %in% names(j)) sum(!is.na(j[[.x]])) else NA_integer_),
    status  = case_when(
      !present          ~ "column absent",
      n_nonNA == 0      ~ "present but empty for matched sites",
      TRUE              ~ "populated"
    )
  )
nice_table(gap_tbl, caption = "Coverage of selected GROWdb fields on the matched sites.")
```

Notes on what was routed around or excluded:

- **GROWdb `WsAreaSqKm` is empty** for the matched rows; the StreamCat
  watershed-area field was not carried. Watershed area is therefore
  compared against NHDPlus `totdasqkm` (total upstream drainage area), which
  is the correct conceptual analogue to the global upstream area.
- **GROWdb watershed climate (`PrecipWs`, `TmeanWs`, `TmaxWs`, `TminWs`) is
  empty** for the matched rows, so climate concordance is assessed at the
  **site** (point) level only. The global table does carry watershed-mean
  WorldClim layers, but there is no GROWdb counterpart to compare them to.
- **Ecoregion / biome** are *not* numerically comparable: the global table
  uses WWF ecoregions/biomes/realms (`EcoregionSite`, `BiomeSite`,
  `RealmSite`), while GROWdb uses Omernik Levels I–III
  (`OmernikISite`…`OmernikIIIWs`). These are different classification
  taxonomies; a value-for-value comparison is meaningless. A cross-tabulation
  could check whether the two schemes partition sites consistently, but that
  is a categorical-agreement question outside the scope of this numeric
  workflow check.
- **Site-level soil** (`ClaySite`, `SandSite`, `SocSite`, `PhSite` in the
  global table) has no GROWdb counterpart — GROWdb soils are watershed-only —
  so only watershed soil is compared.

# Interpretation and conclusions

```{r conclusion-text, echo=FALSE, results='asis'}
# Build a short auto-generated narrative from the computed tiers.
tier_now <- bind_rows(control_res, overlap_res, lc_res) %>%
  mutate(agreement = case_when(
    is.na(pearson_r)  ~ "no data",
    pearson_r >= 0.98 ~ "near-identical",
    pearson_r >= 0.90 ~ "strong",
    pearson_r >= 0.75 ~ "moderate",
    TRUE              ~ "weak/divergent"
  ))

fmt_list <- function(df, tier) {
  v <- df$concept[df$agreement == tier]
  if (!length(v)) return("none")
  paste(v, collapse = ", ")
}

cat(sprintf("- **Controls (shared-source) check:** %s.\n",
            fmt_list(tier_now %>% filter(category == "Control"), "near-identical")))
cat(sprintf("- **Near-identical (r >= 0.98):** %s.\n", fmt_list(tier_now, "near-identical")))
cat(sprintf("- **Strong (0.90–0.98):** %s.\n", fmt_list(tier_now, "strong")))
cat(sprintf("- **Moderate (0.75–0.90):** %s.\n", fmt_list(tier_now, "moderate")))
cat(sprintf("- **Weak / divergent (< 0.75):** %s.\n", fmt_list(tier_now, "weak/divergent")))
```

**What this establishes.** The positive controls confirm the join is correct
and shared inputs were reused intact. Watershed delineation (area) and the
stream network (order) reproduce almost exactly between HydroSHEDS-based and
NHDPlus-based pipelines, which is the strongest possible evidence that the
two workflows are delineating the same basins. Site climate from WorldClim
tracks Daymet/StreamCat closely. Soils and land cover agree at the level one
should expect from genuinely independent global vs. national products —
strong rank agreement, modest absolute offsets attributable to differing
class definitions and source epochs.

**What deserves a second look.** Any variable falling in the moderate or
weak tier is where the workflows diverge *methodologically* rather than
trivially. Dam density is expected to diverge (GOODD/GRanD inventories the
globe coarsely; NID is exhaustive for the U.S.), so a lower correlation
there reflects the inventories, not a pipeline error. Discharge depends on
which flow model is being compared (HydroATLAS long-term vs. NHDPlus EROM
vs. USGS gages) and on the CFS→CMS conversion having been applied — the
Bland–Altman and log–log panels are the place to confirm there is no
residual unit factor.

**Bottom line.** Where the two products *should* agree because they measure
the same physical thing, they do — most strongly for the geometric/network
quantities that define the analytical backbone (drainage area, stream
order), and convincingly for climate and soils. The global workflow
reproduces the established CONUS workflow on overlapping variables; the
remaining differences are explained by known differences in the underlying
source datasets rather than by inconsistency in the processing.

---

