---
title: "Post-process Global Geospatial Master Table"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = FALSE
)
```

# Post-process global geospatial master table

This RMarkdown runs after `4-get-geospatial`.

It fills point-level gaps where the sample point falls in a raster or vector
coverage gap.

Rules:

- For raster-backed **site-scale soil and aridity variables**, fill missing
  values from the nearest non-`NA` grid cell using
  `terra::extract(search_radius = ...)`.
- For RESOLVE ecoregion site variables, fill missing values from the nearest
  RESOLVE polygon.
- Add a `notes` column describing which variables were filled this way.
- Landcover is not changed in this post-processing step.
- Write the updated `site_global_master_table.RDS` and `.csv`.

```{r libraries}
library(tidyverse)
library(sf)
library(terra)
library(geodata)

sf::sf_use_s2(FALSE)
```

## 1. Paths and inputs

```{r paths}
out_dir     <- "data/global_watershed_geospatial"
extract_dir <- file.path(out_dir, "extracted")
soil_path   <- file.path(out_dir, "soilgrids")

master_rds <- file.path(extract_dir, "site_global_master_table.RDS")
master_csv <- file.path(extract_dir, "site_global_master_table.csv")
```

```{r read-master}
if (file.exists(master_rds)) {
  site_master_table <- readRDS(master_rds)
} else if (file.exists(master_csv)) {
  site_master_table <- readr::read_csv(master_csv, show_col_types = FALSE)
} else {
  stop("Could not find site_global_master_table.RDS or site_global_master_table.csv in ", extract_dir)
}

if (!"notes" %in% names(site_master_table)) {
  site_master_table <- site_master_table %>%
    dplyr::mutate(notes = NA_character_)
}

site_master_table <- site_master_table %>%
  dplyr::mutate(notes = as.character(notes))
```

## 2. Load source layers needed for filling gaps

These should already exist locally from the main geospatial workflow. The
`geodata::soil_world()` calls are included so the script can also rebuild the
local SoilGrids cache if needed.

```{r load-source-layers}
# SoilGrids rasters --------------------------------------------------------

dir.create(soil_path, recursive = TRUE, showWarnings = FALSE)

# geodata::soil_world(var = "clay",  depth = 30, stat = "mean", path = soil_path)
# geodata::soil_world(var = "sand",  depth = 30, stat = "mean", path = soil_path)
# geodata::soil_world(var = "soc",   depth = 30, stat = "mean", path = soil_path)
# geodata::soil_world(var = "phh2o", depth = 30, stat = "mean", path = soil_path)

soil_clay <- terra::rast(list.files(file.path(soil_path, "soil_world"), pattern = "clay.*30.*mean",  full.names = TRUE)[1])
soil_sand <- terra::rast(list.files(file.path(soil_path, "soil_world"), pattern = "sand.*30.*mean",  full.names = TRUE)[1])
soil_soc  <- terra::rast(list.files(file.path(soil_path, "soil_world"), pattern = "soc.*30.*mean",   full.names = TRUE)[1])
soil_ph   <- terra::rast(list.files(file.path(soil_path, "soil_world"), pattern = "phh2o.*30.*mean", full.names = TRUE)[1])

# Aridity -----------------------------------------------------------------
aridity <- terra::rast("data/ai_et0.tif")

# RESOLVE ecoregions ------------------------------------------------------
resolve <- sf::st_read(
  "data/global_watershed_geospatial/resolve/e7c09d3e-4f20-43d9-8873-60c10c28cc7b.gdb/",
  quiet = TRUE
) %>%
  dplyr::select(ECO_NAME, BIOME_NAME, REALM)
```

## 3. Helper functions

```{r helper-functions}
append_note <- function(existing, new_note) {
  dplyr::case_when(
    is.na(existing) | existing == "" ~ new_note,
    stringr::str_detect(existing, stringr::fixed(new_note)) ~ existing,
    TRUE ~ paste(existing, new_note, sep = " | ")
  )
}

# Faster raster gap fill:
# terra::extract(search_radius = ...) searches for the nearest non-NA cell
# within search_radius_m when the point itself lands on an NA cell.
fill_nearest_raster_fast <- function(df,
                                     r,
                                     col,
                                     note_text,
                                     search_radius_m = 50000) {

  if (!col %in% names(df)) {
    message("Skipping missing column: ", col)
    return(df)
  }

  if (!inherits(r, "SpatRaster")) {
    r <- terra::rast(r)
  }

  if (terra::nlyr(r) > 1) {
    r <- r[[1]]
  }

  # Robust missing index. Works for numeric and non-numeric columns.
  missing_idx <- which(is.na(df[[col]]))

  if (is.numeric(df[[col]])) {
    missing_idx <- which(is.na(df[[col]]) | is.nan(df[[col]]))
  }

  missing_idx <- missing_idx[!is.na(missing_idx)]

  if (length(missing_idx) == 0) {
    return(df)
  }

  pts_missing <- df[missing_idx, ] %>%
    sf::st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE) %>%
    sf::st_transform(terra::crs(r))

  extracted <- tryCatch(
    terra::extract(
      r,
      terra::vect(pts_missing),
      search_radius = search_radius_m
    ),
    error = function(e) {
      warning("terra::extract failed for ", col, ": ", e$message)
      NULL
    }
  )

  if (is.null(extracted)) {
    return(df)
  }

  vals <- extracted %>%
    dplyr::select(-ID) %>%
    dplyr::pull(1)

  fillable <- !is.na(vals)

  fill_idx <- missing_idx[fillable]
  fill_vals <- vals[fillable]

  # Protect against NA subscripts, NA values, and length mismatches.
  keep <- !is.na(fill_idx) & !is.na(fill_vals)
  fill_idx <- fill_idx[keep]
  fill_vals <- fill_vals[keep]

  if (length(fill_idx) > 0) {
    df[[col]][fill_idx] <- fill_vals

    df$notes[fill_idx] <- append_note(
      df$notes[fill_idx],
      note_text
    )
  }

  if (length(fill_idx) < length(missing_idx)) {
    warning(
      length(missing_idx) - length(fill_idx),
      " missing values could not be filled within ",
      search_radius_m,
      " m for: ",
      col
    )
  }

  df
}

fill_nearest_resolve <- function(df, resolve) {

  resolve_vars <- c("EcoregionSite", "BiomeSite", "RealmSite")
  resolve_vars <- resolve_vars[resolve_vars %in% names(df)]

  if (length(resolve_vars) == 0) {
    return(df)
  }

  missing_idx <- which(rowSums(is.na(df[, resolve_vars, drop = FALSE])) > 0)
  missing_idx <- missing_idx[!is.na(missing_idx)]

  if (length(missing_idx) == 0) {
    return(df)
  }

  pts_missing <- df[missing_idx, ] %>%
    sf::st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE) %>%
    sf::st_transform(sf::st_crs(resolve))

  nearest_id <- sf::st_nearest_feature(pts_missing, resolve)
  nearest_resolve <- resolve[nearest_id, ] %>%
    sf::st_drop_geometry()

  if ("EcoregionSite" %in% names(df)) {
    df$EcoregionSite[missing_idx] <- dplyr::coalesce(
      df$EcoregionSite[missing_idx],
      nearest_resolve$ECO_NAME
    )
  }

  if ("BiomeSite" %in% names(df)) {
    df$BiomeSite[missing_idx] <- dplyr::coalesce(
      df$BiomeSite[missing_idx],
      nearest_resolve$BIOME_NAME
    )
  }

  if ("RealmSite" %in% names(df)) {
    df$RealmSite[missing_idx] <- dplyr::coalesce(
      df$RealmSite[missing_idx],
      nearest_resolve$REALM
    )
  }

  df$notes[missing_idx] <- append_note(
    df$notes[missing_idx],
    "RESOLVE site fields filled from nearest RESOLVE polygon because the sample point fell in a vector coverage gap."
  )

  df
}
```

## 4. Fill soil site variables from nearest non-NA SoilGrids cells

```{r fill-soil}
site_master_table <- site_master_table %>%
  fill_nearest_raster_fast(
    r = soil_clay,
    col = "ClaySite",
    note_text = "ClaySite filled from nearest non-NA SoilGrids clay cell because the sample point overlaid a raster gap.",
    search_radius_m = 50000
  ) %>%
  fill_nearest_raster_fast(
    r = soil_sand,
    col = "SandSite",
    note_text = "SandSite filled from nearest non-NA SoilGrids sand cell because the sample point overlaid a raster gap.",
    search_radius_m = 50000
  ) %>%
  fill_nearest_raster_fast(
    r = soil_soc,
    col = "SocSite",
    note_text = "SocSite filled from nearest non-NA SoilGrids soil organic carbon cell because the sample point overlaid a raster gap.",
    search_radius_m = 50000
  ) %>%
  fill_nearest_raster_fast(
    r = soil_ph,
    col = "PhSite",
    note_text = "PhSite filled from nearest non-NA SoilGrids pH cell because the sample point overlaid a raster gap.",
    search_radius_m = 50000
  )
```

## 5. Fill aridity site variable from nearest non-NA aridity raster cell

```{r fill-aridity}
site_master_table <- site_master_table %>%
  fill_nearest_raster_fast(
    r = aridity,
    col = "AriditySite",
    note_text = "AriditySite filled from nearest non-NA aridity raster cell because the sample point overlaid a raster gap.",
    search_radius_m = 100000
  ) %>%
  dplyr::mutate(
    # CGIAR aridity index is stored as x10^4 in the source raster.
    # This keeps filled raw raster values on the same scale as the rest of the table.
    AriditySite = dplyr::if_else(
      !is.na(AriditySite) & AriditySite > 100,
      AriditySite / 10000,
      AriditySite
    )
  )
```

## 6. Fill RESOLVE site variables from nearest polygon

```{r fill-resolve}
site_master_table <- site_master_table %>%
  fill_nearest_resolve(resolve)
```

## 7. Add annual climate summaries and write outputs

```{r add-annual-climate-summaries}
site_master_table <- site_master_table %>%
  dplyr::mutate(
    # Watershed-scale WorldClim annual summaries ---------------------------
    tminWs_annual_mean_c = rowMeans(
      dplyr::across(dplyr::matches("^mean\\.wc2\\.1_2\\.5m_tmin_\\d{2}$")),
      na.rm = TRUE
    ),
    tmaxWs_annual_mean_c = rowMeans(
      dplyr::across(dplyr::matches("^mean\\.wc2\\.1_2\\.5m_tmax_\\d{2}$")),
      na.rm = TRUE
    ),
    tmeanWs_annual_mean_c = rowMeans(
      dplyr::across(dplyr::matches("^mean\\.wc2\\.1_2\\.5m_tavg_\\d{2}$")),
      na.rm = TRUE
    ),
    precipWs_annual_sum_mm = rowSums(
      dplyr::across(dplyr::matches("^mean\\.wc2\\.1_2\\.5m_prec_\\d{2}$")),
      na.rm = TRUE
    ),

    # Site-scale WorldClim annual summaries --------------------------------
    tminSite_annual_mean_c = rowMeans(
      dplyr::across(dplyr::matches("^tminSite_wc2\\.1_2\\.5m_tmin_\\d{2}$")),
      na.rm = TRUE
    ),
    tmaxSite_annual_mean_c = rowMeans(
      dplyr::across(dplyr::matches("^tmaxSite_wc2\\.1_2\\.5m_tmax_\\d{2}$")),
      na.rm = TRUE
    ),
    tmeanSite_annual_mean_c = rowMeans(
      dplyr::across(dplyr::matches("^tavgSite_wc2\\.1_2\\.5m_tavg_\\d{2}$")),
      na.rm = TRUE
    ),
    precipSite_annual_sum_mm = rowSums(
      dplyr::across(dplyr::matches("^precSite_wc2\\.1_2\\.5m_prec_\\d{2}$")),
      na.rm = TRUE
    )
  ) %>%
  dplyr::mutate(
    # rowMeans/rowSums with na.rm = TRUE returns NaN or 0 when all months are NA.
    # Convert those all-missing annual summaries back to NA.
    dplyr::across(
      c(
        tminWs_annual_mean_c,
        tmaxWs_annual_mean_c,
        tmeanWs_annual_mean_c,
        precipWs_annual_sum_mm,
        tminSite_annual_mean_c,
        tmaxSite_annual_mean_c,
        tmeanSite_annual_mean_c,
        precipSite_annual_sum_mm
      ),
      ~ dplyr::na_if(.x, NaN)
    )
  )
```
## 7. Final cleanup and write outputs

```{r write-outputs}
# Keep notes near the end.
site_master_table <- site_master_table %>%
  dplyr::relocate(notes, .after = dplyr::last_col())

saveRDS(site_master_table, file.path(extract_dir, "site_global_master_table.RDS"))
readr::write_csv(site_master_table, file.path(extract_dir, "site_global_master_table.csv"))
```

## 8. Summarize remaining missing values

```{r summarize-na}
summarize_na <- site_master_table %>%
  dplyr::summarise(dplyr::across(dplyr::everything(), ~ sum(is.na(.x)))) %>%
  tidyr::pivot_longer(
    dplyr::everything(),
    names_to = "column",
    values_to = "n_na"
  ) %>%
  dplyr::filter(n_na > 0) %>%
  dplyr::arrange(dplyr::desc(n_na))

summarize_na
```

## 9. Inspect rows with notes

```{r inspect-notes}
site_master_table %>%
  dplyr::filter(!is.na(notes), notes != "") %>%
  dplyr::select(
    sample_name,
    Latitude,
    Longitude,
    ClaySite,
    SandSite,
    SocSite,
    PhSite,
    AriditySite,
    EcoregionSite,
    BiomeSite,
    RealmSite,
    notes
  )
```

```{r}
source("src/review_outputs.R")

site_master_table_reviewed <- review_outputs(
  site_master_table = site_master_table,
  flowlines = hydro_rivers,
  river_buffer_m = 10000
)

saveRDS(
  site_master_table_reviewed,
  file.path(extract_dir, "site_global_master_table_reviewed.RDS")
)

readr::write_csv(
  site_master_table_reviewed,
  file.path(extract_dir, "site_global_master_table_reviewed.csv")
)
```

