---
title: "Comparing CONUS + Global Products"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 6,
  dpi = 110
)

suppressPackageStartupMessages({
  library(readr)
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(stringr)
  library(ggplot2)
  library(tibble)
  library(knitr)
})

have_kableExtra <- requireNamespace("kableExtra", quietly = TRUE)

nice_table <- function(df, caption = NULL, digits = 3) {
  k <- knitr::kable(df, caption = caption, digits = digits, format = "html")
  if (have_kableExtra) {
    k <- kableExtra::kable_styling(
      k,
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "left"
    )
  }
  k
}

theme_set(theme_minimal(base_size = 12))
```

# Purpose

`GROWdb` is the **CONUS-only** product. It was built largely from
U.S.-specific source layers, including NHDPlus V2, StreamCat, and USGS gage
data. The **global master table** (`site_global_master_table.csv`) re-derives
an analogous set of descriptors for the same sampling sites using global
source layers, including HydroSHEDS / HydroBASINS / HydroRIVERS / HydroATLAS,
WorldClim 2.1, global land cover, SoilGrids-style soils, and related global
products.

Because the two products draw on different underlying datasets, identical
numbers are not expected everywhere. The goal here is narrower: for variables
that represent the same physical concept, check whether the global workflow
lands in the same general place as the established CONUS workflow.

This version assumes the global post-processing workflow has already added
annual climate summaries to `site_global_master_table`, including:

- `tminSite_annual_mean_c`, `tmaxSite_annual_mean_c`, `tmeanSite_annual_mean_c`, `precipSite_annual_sum_mm`
- `tminWs_annual_mean_c`, `tmaxWs_annual_mean_c`, `tmeanWs_annual_mean_c`, `precipWs_annual_sum_mm`

NPP is intentionally excluded because it is no longer part of the global
workflow.

# Load and join

```{r load}
raw <- read_csv("data/date_lat_long_grow_KKA_05.05.26.csv", show_col_types = FALSE)
grow <- read_csv("data/0-shared/GROWdb_20260528.csv", show_col_types = FALSE)
global <- read_csv("data/global_watershed_geospatial/extracted/site_global_master_table.csv", show_col_types = FALSE)

cat("Master table:", nrow(raw), "rows x", ncol(raw), "cols\n")
cat("Global table:", nrow(global), "rows x", ncol(global), "cols\n")
cat("CONUS table:", nrow(grow), "rows x", ncol(grow), "cols\n")
```

The two tables share a sample identifier: `sample_name` in the global table
and `SampleName` in GROWdb. We join on that identifier rather than coordinates,
because the two products may carry slightly different coordinate precision.

```{r join}
global <- global %>%
  rename(sample_id = sample_name)

grow <- grow %>%
  rename(sample_id = SampleName)

j <- inner_join(global, grow, by = "sample_id", suffix = c("_global", "_grow"))

n_global_only <- length(setdiff(global$sample_id, grow$sample_id))
n_grow_only <- length(setdiff(grow$sample_id, global$sample_id))

cat("Matched site-pairs:", nrow(j), "\n")
cat("In global only:", n_global_only, "\n")
cat("In GROWdb only:", n_grow_only, "\n")
```

# Helper functions

These helpers make the analysis robust to post-join suffixes. If a field exists
in both input tables, it is usually renamed to `_global` and `_grow`. If it
exists in only one table, it usually remains unsuffixed. The helpers below check
all plausible candidate names and report the actual column used.

```{r helpers}
first_present <- function(data, candidates) {
  candidates <- unlist(candidates, use.names = FALSE)
  hits <- candidates[candidates %in% names(data)]
  if (length(hits) == 0) NA_character_ else hits[1]
}

as_numeric_vector <- function(x, n_expected = NULL) {
  # readr/dplyr joins should produce atomic vectors, but this keeps the workflow
  # from breaking if a previous step accidentally created a one-value list column.
  if (is.data.frame(x)) {
    x <- x[[1]]
  }

  if (is.list(x) && !is.data.frame(x)) {
    x <- purrr::map_dbl(x, function(z) {
      if (length(z) == 0 || all(is.na(z))) {
        NA_real_
      } else {
        suppressWarnings(as.numeric(z[[1]]))
      }
    })
  }

  x <- suppressWarnings(as.numeric(x))

  if (!is.null(n_expected) && length(x) != n_expected) {
    out <- rep(NA_real_, n_expected)
    n_copy <- min(length(x), n_expected)
    out[seq_len(n_copy)] <- x[seq_len(n_copy)]
    x <- out
  }

  x
}

pick_numeric <- function(data, candidates, transform = identity) {
  candidates <- unlist(candidates, use.names = FALSE)
  hits <- candidates[candidates %in% names(data)]

  if (length(hits) == 0) {
    return(rep(NA_real_, nrow(data)))
  }

  transform <- purrr::as_mapper(transform)

  vals <- purrr::map(hits, ~ as_numeric_vector(data[[.x]], n_expected = nrow(data)))
  n_finite <- purrr::map_int(vals, ~ sum(is.finite(.x)))

  best_i <- if (any(n_finite > 0)) {
    which(n_finite > 0)[1]
  } else {
    1
  }

  as_numeric_vector(transform(vals[[best_i]]), n_expected = nrow(data))
}

row_sum_candidates <- function(data, candidates) {
  hits <- candidates[candidates %in% names(data)]

  if (length(hits) == 0) {
    return(rep(NA_real_, nrow(data)))
  }

  mat <- data %>%
    dplyr::select(dplyr::all_of(hits)) %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), ~ suppressWarnings(as.numeric(.x)))) %>%
    as.data.frame()

  out <- rowSums(mat, na.rm = TRUE)
  all_missing <- rowSums(!is.na(mat)) == 0
  out[all_missing] <- NA_real_
  out
}

count_non_na <- function(data, candidates) {
  hits <- candidates[candidates %in% names(data)]

  if (length(hits) == 0) {
    return(tibble(
      present = FALSE,
      column_used = NA_character_,
      n_nonNA = NA_integer_
    ))
  }

  counts <- purrr::map_int(hits, ~ sum(!is.na(data[[.x]])))

  best_i <- if (any(counts > 0)) {
    which(counts > 0)[1]
  } else {
    1
  }

  tibble(
    present = TRUE,
    column_used = hits[best_i],
    n_nonNA = counts[best_i]
  )
}

empty_stats <- function(n = 0L) {
  tibble(
    n = as.integer(n),
    pearson_r = NA_real_,
    spearman_rho = NA_real_,
    bias_grow_minus_global = NA_real_,
    rmse = NA_real_,
    global_median = NA_real_,
    grow_median = NA_real_
  )
}

safe_cor <- function(x, y, method) {
  if (length(x) < 3 || stats::sd(x) == 0 || stats::sd(y) == 0) {
    return(NA_real_)
  }
  suppressWarnings(stats::cor(x, y, method = method))
}

compare_pair <- function(data, x_col, y_col) {
  if (!x_col %in% names(data) || !y_col %in% names(data)) {
    return(empty_stats(0L))
  }

  x <- as_numeric_vector(data[[x_col]], n_expected = nrow(data))
  y <- as_numeric_vector(data[[y_col]], n_expected = nrow(data))
  ok <- is.finite(x) & is.finite(y)
  x <- x[ok]
  y <- y[ok]
  n <- length(x)

  if (n == 0) {
    return(empty_stats(0L))
  }

  tibble(
    n = as.integer(n),
    pearson_r = safe_cor(x, y, method = "pearson"),
    spearman_rho = safe_cor(x, y, method = "spearman"),
    bias_grow_minus_global = mean(y - x),
    rmse = sqrt(mean((y - x)^2)),
    global_median = median(x),
    grow_median = median(y)
  )
}

run_specs <- function(data, specs) {
  required_cols <- c("concept", "category", "x_col", "y_col", "unit", "note")
  missing_cols <- setdiff(required_cols, names(specs))

  if (length(missing_cols) > 0) {
    stop("Specs table is missing required columns: ", paste(missing_cols, collapse = ", "))
  }

  specs %>%
    dplyr::mutate(
      x_present = x_col %in% names(data),
      y_present = y_col %in% names(data),
      stats = purrr::map2(x_col, y_col, ~ compare_pair(data, .x, .y))
    ) %>%
    tidyr::unnest(cols = c(stats)) %>%
    dplyr::select(
      category, concept, unit, n, pearson_r, spearman_rho,
      bias_grow_minus_global, rmse, global_median, grow_median,
      x_col, y_col, x_present, y_present, note
    )
}

make_pairs <- function(data, specs) {
  purrr::pmap_dfr(
    specs,
    function(concept, category, x_col, y_col, unit, note, ...) {
      if (!x_col %in% names(data) || !y_col %in% names(data)) {
        return(tibble())
      }

      tibble(
        concept = concept,
        category = category,
        unit = unit,
        global = as_numeric_vector(data[[x_col]], n_expected = nrow(data)),
        grow = as_numeric_vector(data[[y_col]], n_expected = nrow(data))
      )
    }
  ) %>%
    dplyr::filter(is.finite(global), is.finite(grow))
}
```

# Build analysis columns

This step creates stable analysis columns with explicit names. The rest of the
Rmd uses only these stable names, so the comparison does not break if the join
creates `_global` / `_grow` suffixes differently than expected.

```{r derived}
j <- j %>%
  mutate(
    # Global-side aliases ---------------------------------------------------
    global_aridity_ws = pick_numeric(., c("AridityWs_global", "AridityWs")),
    global_aridity_site = pick_numeric(., c("AriditySite_global", "AriditySite")),
    global_area_km2 = pick_numeric(., c("AreaSqKmWs_global", "AreaSqKmWs", "WsAreaSqKm_global", "WsAreaSqKm")),
    global_stream_order = pick_numeric(., c("StreamOrder_global", "StreamOrder")),
    global_discharge_cms = pick_numeric(., c("DischargeCMS_global", "DischargeCMS")),
    global_elevation_site_m = pick_numeric(., c("ElevationSite_global", "ElevationSite")),

    global_precip_site_mm = pick_numeric(., c("precipSite_annual_sum_mm_global", "precipSite_annual_sum_mm")),
    global_tmean_site_c = pick_numeric(., c("tmeanSite_annual_mean_c_global", "tmeanSite_annual_mean_c")),
    global_tmax_site_c = pick_numeric(., c("tmaxSite_annual_mean_c_global", "tmaxSite_annual_mean_c")),
    global_tmin_site_c = pick_numeric(., c("tminSite_annual_mean_c_global", "tminSite_annual_mean_c")),

    global_precip_ws_mm = pick_numeric(., c("precipWs_annual_sum_mm_global", "precipWs_annual_sum_mm")),
    global_tmean_ws_c = pick_numeric(., c("tmeanWs_annual_mean_c_global", "tmeanWs_annual_mean_c")),
    global_tmax_ws_c = pick_numeric(., c("tmaxWs_annual_mean_c_global", "tmaxWs_annual_mean_c")),
    global_tmin_ws_c = pick_numeric(., c("tminWs_annual_mean_c_global", "tminWs_annual_mean_c")),

    global_clay_ws = pick_numeric(., c("ClayWs_global", "ClayWs")),
    global_sand_ws = pick_numeric(., c("SandWs_global", "SandWs")),
    global_dam_dens_ws = pick_numeric(., c("DamDensWs_global", "DamDensWs")),

    # GROWdb-side aliases ---------------------------------------------------
    grow_aridity_ws = pick_numeric(., c("AridityWs_grow", "AridityWs")),
    grow_aridity_site = pick_numeric(., c("AriditySite_grow", "AriditySite")),
    grow_area_km2 = pick_numeric(., c("AreaSqKmWs_grow", "WsAreaSqKm_grow", "totdasqkm", "AreaSqKmWs", "WsAreaSqKm")),
    grow_stream_order = pick_numeric(., c("streamorde_grow", "streamorde")),
    grow_qe_ma_cms = pick_numeric(., c("qe_ma_grow", "qe_ma"), transform = ~ .x * 0.0283168),
    grow_cfs_cms = pick_numeric(., c("CFS_grow", "CFS"), transform = ~ .x * 0.0283168),
    grow_minelev_m = pick_numeric(., c("minelevsmo_grow", "minelevsmo"), transform = ~ .x / 100),

    grow_precip_site_mm = pick_numeric(., c("PrecipSite_grow", "PrecipSite")),
    grow_tmean_site_c = pick_numeric(., c("TmeanSite_grow", "TmeanSite")),
    grow_tmax_site_c = pick_numeric(., c("TmaxSite_grow", "TmaxSite")),
    grow_tmin_site_c = pick_numeric(., c("TminSite_grow", "TminSite")),

    grow_precip_ws_mm = pick_numeric(., c("PrecipWs_grow", "PrecipWs")),
    grow_tmean_ws_c = pick_numeric(., c("TmeanWs_grow", "TmeanWs")),
    grow_tmax_ws_c = pick_numeric(., c("TmaxWs_grow", "TmaxWs")),
    grow_tmin_ws_c = pick_numeric(., c("TminWs_grow", "TminWs")),

    grow_clay_ws = pick_numeric(., c("ClayWs_grow", "ClayWs")),
    grow_sand_ws = pick_numeric(., c("SandWs_grow", "SandWs")),
    grow_dam_dens_ws = pick_numeric(., c("DamDensWs_grow", "DamDensWs"))
  )
```

```{r column-resolution, echo=FALSE}
resolution_specs <- tribble(
  ~analysis_field, ~candidate_columns,
  "global_area_km2", list(c("AreaSqKmWs_global", "AreaSqKmWs", "WsAreaSqKm_global", "WsAreaSqKm")),
  "grow_area_km2", list(c("AreaSqKmWs_grow", "WsAreaSqKm_grow", "totdasqkm", "AreaSqKmWs", "WsAreaSqKm")),
  "global_precip_ws_mm", list(c("precipWs_annual_sum_mm_global", "precipWs_annual_sum_mm")),
  "grow_precip_ws_mm", list(c("PrecipWs_grow", "PrecipWs")),
  "global_tmean_ws_c", list(c("tmeanWs_annual_mean_c_global", "tmeanWs_annual_mean_c")),
  "grow_tmean_ws_c", list(c("TmeanWs_grow", "TmeanWs")),
  "global_tmax_ws_c", list(c("tmaxWs_annual_mean_c_global", "tmaxWs_annual_mean_c")),
  "grow_tmax_ws_c", list(c("TmaxWs_grow", "TmaxWs")),
  "global_tmin_ws_c", list(c("tminWs_annual_mean_c_global", "tminWs_annual_mean_c")),
  "grow_tmin_ws_c", list(c("TminWs_grow", "TminWs"))
)

resolution_tbl <- resolution_specs %>%
  mutate(info = map(candidate_columns, ~ count_non_na(j, .x))) %>%
  unnest(info) %>%
  select(analysis_field, column_used, present, n_nonNA)

nice_table(resolution_tbl, caption = "Resolved columns used for selected analysis fields.")
```

# Missing value summary for the global table

```{r missing-global}
na_summary_global <- global %>%
  summarise(across(everything(), ~ sum(is.na(.x)))) %>%
  pivot_longer(
    everything(),
    names_to = "column",
    values_to = "n_na"
  ) %>%
  filter(n_na > 0) %>%
  arrange(desc(n_na))

nice_table(na_summary_global, caption = "Remaining missing values in the global master table.")
```

# Positive controls: shared-source variables

Aridity is kept as the main positive control. NPP is excluded because it is no
longer used in this workflow.

```{r controls}
control_specs <- tribble(
  ~concept, ~category, ~x_col, ~y_col, ~unit, ~note,
  "Aridity (watershed)", "Control", "global_aridity_ws", "grow_aridity_ws", "index", "Shared-source or directly comparable aridity field",
  "Aridity (site)", "Control", "global_aridity_site", "grow_aridity_site", "index", "Shared-source or directly comparable aridity field"
)

control_res <- run_specs(j, control_specs)

nice_table(
  control_res %>% dplyr::select(-category, -x_col, -y_col, -x_present, -y_present),
  caption = "Positive controls. Expect near-identical results if the join and units are correct."
)
```

# Crosswalk of overlapping variables

```{r crosswalk-doc, echo=FALSE}
crosswalk_doc <- tribble(
  ~Concept, ~`Global source`, ~`GROWdb source`, ~`Unit reconciliation`,
  "Watershed / drainage area", "HydroSHEDS / HydroBASINS upstream area", "NHDPlus V2 total drainage area", "both km^2",
  "Stream order", "HydroRIVERS Strahler", "NHDPlus stream order", "none",
  "Mean annual discharge", "HydroATLAS DischargeCMS", "NHDPlus EROM qe_ma", "GROW CFS -> CMS (x0.0283168)",
  "Discharge at gage", "HydroATLAS DischargeCMS", "USGS gage CFS", "GROW CFS -> CMS (x0.0283168)",
  "Site elevation", "DEM point elevation", "NHDPlus smoothed minimum elevation", "GROW cm -> m (/100)",
  "Site annual precip", "WorldClim 2.1 annual summary", "GROWdb / StreamCat site precip", "both mm/yr",
  "Site annual tmean", "WorldClim 2.1 annual summary", "GROWdb / StreamCat site tmean", "both degC",
  "Site annual tmax", "WorldClim 2.1 annual summary", "GROWdb / StreamCat site tmax", "both degC",
  "Site annual tmin", "WorldClim 2.1 annual summary", "GROWdb / StreamCat site tmin", "both degC",
  "Watershed annual precip", "WorldClim 2.1 watershed annual summary", "GROWdb / StreamCat watershed precip", "both mm/yr",
  "Watershed annual tmean", "WorldClim 2.1 watershed annual summary", "GROWdb / StreamCat watershed tmean", "both degC",
  "Watershed annual tmax", "WorldClim 2.1 watershed annual summary", "GROWdb / StreamCat watershed tmax", "both degC",
  "Watershed annual tmin", "WorldClim 2.1 watershed annual summary", "GROWdb / StreamCat watershed tmin", "both degC",
  "Watershed clay", "SoilGrids-style global soil", "StreamCat soil", "both %",
  "Watershed sand", "SoilGrids-style global soil", "StreamCat soil", "both %",
  "Watershed dam density", "GOODD / GRanD-style global inventories", "NID / StreamCat", "different inventories; divergence expected",
  "Land cover", "Copernicus/ESA fractions", "NLCD 2016 StreamCat percentages", "global fraction * 100 vs aggregated NLCD %"
)

nice_table(crosswalk_doc, caption = "Conceptual crosswalk between the two products.")
```

```{r overlap-specs}
overlap_specs <- tribble(
  ~concept, ~category, ~x_col, ~y_col, ~unit, ~note,
  "Watershed area", "Delineation", "global_area_km2", "grow_area_km2", "km^2", "HydroSHEDS/HydroBASINS vs NHDPlus total drainage area",
  "Stream order", "Network", "global_stream_order", "grow_stream_order", "Strahler", "HydroRIVERS vs NHDPlus",
  "Discharge (modeled)", "Hydrology", "global_discharge_cms", "grow_qe_ma_cms", "m^3/s", "HydroATLAS vs NHDPlus EROM",
  "Discharge (gage)", "Hydrology", "global_discharge_cms", "grow_cfs_cms", "m^3/s", "HydroATLAS vs USGS gage",
  "Site elevation", "Terrain", "global_elevation_site_m", "grow_minelev_m", "m", "Point DEM vs NHDPlus reach elevation",

  "Site precip (annual)", "Climate", "global_precip_site_mm", "grow_precip_site_mm", "mm/yr", "WorldClim site annual precip vs GROWdb site precip",
  "Site tmean (annual)", "Climate", "global_tmean_site_c", "grow_tmean_site_c", "degC", "WorldClim site annual tmean vs GROWdb site tmean",
  "Site tmax (annual)", "Climate", "global_tmax_site_c", "grow_tmax_site_c", "degC", "WorldClim site annual tmax vs GROWdb site tmax",
  "Site tmin (annual)", "Climate", "global_tmin_site_c", "grow_tmin_site_c", "degC", "WorldClim site annual tmin vs GROWdb site tmin",

  "Watershed precip (annual)", "Climate", "global_precip_ws_mm", "grow_precip_ws_mm", "mm/yr", "WorldClim watershed annual precip vs GROWdb watershed precip",
  "Watershed tmean (annual)", "Climate", "global_tmean_ws_c", "grow_tmean_ws_c", "degC", "WorldClim watershed annual tmean vs GROWdb watershed tmean",
  "Watershed tmax (annual)", "Climate", "global_tmax_ws_c", "grow_tmax_ws_c", "degC", "WorldClim watershed annual tmax vs GROWdb watershed tmax",
  "Watershed tmin (annual)", "Climate", "global_tmin_ws_c", "grow_tmin_ws_c", "degC", "WorldClim watershed annual tmin vs GROWdb watershed tmin",

  "Watershed clay", "Soil", "global_clay_ws", "grow_clay_ws", "%", "SoilGrids vs StreamCat",
  "Watershed sand", "Soil", "global_sand_ws", "grow_sand_ws", "%", "SoilGrids vs StreamCat",
  "Watershed dam density", "Human impact", "global_dam_dens_ws", "grow_dam_dens_ws", "dams/km^2", "Different dam inventories; divergence expected"
)

overlap_res <- run_specs(j, overlap_specs)

overlap_skipped <- overlap_res %>%
  dplyr::filter(n == 0 | !x_present | !y_present) %>%
  dplyr::select(concept, x_col, y_col, x_present, y_present, n, note)

nice_table(
  overlap_res %>%
    dplyr::select(-category, -x_col, -y_col, -x_present, -y_present),
  caption = "Concordance of overlapping variables."
)

if (nrow(overlap_skipped) > 0) {
  nice_table(
    overlap_skipped,
    caption = "Comparisons skipped or unresolved because one side was missing or had no paired finite values."
  )
}
```

# Concordance summary

```{r ranked}
all_res <- bind_rows(control_res, overlap_res) %>%
  arrange(desc(pearson_r))

tiered <- all_res %>%
  mutate(agreement = case_when(
    is.na(pearson_r) ~ "no data",
    pearson_r >= 0.98 ~ "near-identical",
    pearson_r >= 0.90 ~ "strong",
    pearson_r >= 0.75 ~ "moderate",
    TRUE ~ "weak / divergent"
  )) %>%
  select(category, concept, agreement, n, pearson_r, spearman_rho,
         bias_grow_minus_global, rmse)

nice_table(tiered, caption = "All comparisons ranked by agreement tier.")
```

```{r tier-counts, echo=FALSE}
tier_summary <- tiered %>%
  filter(!is.na(pearson_r)) %>%
  count(agreement) %>%
  arrange(match(agreement, c("near-identical", "strong", "moderate", "weak / divergent")))

nice_table(tier_summary, caption = "Count of variables by agreement tier.")
```

# Visual concordance

## 1:1 scatter plots

```{r scatter-data}
pairs_long <- make_pairs(j, overlap_specs)
```

```{r scatter-plot, fig.height=10}
ggplot(pairs_long, aes(global, grow)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(alpha = 0.5, size = 1.3, colour = "#2c7fb8") +
  facet_wrap(~ concept, scales = "free", ncol = 4) +
  labs(
    title = "GROWdb vs. global master table — matched samples",
    subtitle = "Dashed line = 1:1. Closeness to the line indicates workflow concordance.",
    x = "Global master table value",
    y = "GROWdb value"
  ) +
  theme(strip.text = element_text(face = "bold", size = 8))
```

## Log-log view for magnitude-spanning variables

```{r scatter-log, fig.height=4.5}
log_concepts <- c("Watershed area", "Discharge (modeled)", "Discharge (gage)")

pairs_long %>%
  filter(concept %in% log_concepts, global > 0, grow > 0) %>%
  ggplot(aes(global, grow)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(alpha = 0.5, size = 1.3, colour = "#d95f0e") +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~ concept, scales = "free", ncol = 3) +
  labs(
    title = "Log-log concordance for magnitude-spanning variables",
    x = "Global value (log10)",
    y = "GROWdb value (log10)"
  )
```

## Bland-Altman plots

```{r bland-altman, fig.height=9}
ba_concepts <- c(
  "Site precip (annual)",
  "Site tmean (annual)",
  "Site tmax (annual)",
  "Site tmin (annual)",
  "Watershed precip (annual)",
  "Watershed tmean (annual)",
  "Watershed tmax (annual)",
  "Watershed tmin (annual)",
  "Watershed clay",
  "Stream order"
)

pairs_long %>%
  filter(concept %in% ba_concepts) %>%
  mutate(
    mean_val = (global + grow) / 2,
    diff_val = grow - global
  ) %>%
  ggplot(aes(mean_val, diff_val)) +
  geom_hline(yintercept = 0, colour = "grey40") +
  geom_point(alpha = 0.5, size = 1.3, colour = "#31a354") +
  geom_smooth(method = "loess", se = FALSE, colour = "#de2d26", linewidth = 0.7) +
  facet_wrap(~ concept, scales = "free", ncol = 3) +
  labs(
    title = "Bland-Altman: GROWdb - Global vs. mean of the two",
    subtitle = "Flat red line near zero = no systematic bias.",
    x = "Mean of the two products",
    y = "Difference (GROWdb - Global)"
  )
```

# Land cover: Copernicus fractions vs. NLCD classes

The global table uses continuous land-cover fractions. GROWdb uses NLCD 2016
class percentages from StreamCat. To compare them, NLCD classes are aggregated
into broad groups and global fractions are converted to percentages.

```{r landcover}
lc <- j %>%
  transmute(
    sample_id = sample_id,

    global_forest = 100 * pick_numeric(., c("lc_trees_mean_fraction_global", "lc_trees_mean_fraction")),
    grow_forest = row_sum_candidates(., c("PctDecid2016Ws", "PctConif2016Ws", "PctMxFst2016Ws")),

    global_crop = 100 * pick_numeric(., c("lc_cropland_mean_fraction_global", "lc_cropland_mean_fraction")),
    grow_crop = row_sum_candidates(., c("PctCrop2016Ws", "PctHay2016Ws")),

    global_grshrub = 100 * (
      pick_numeric(., c("lc_grassland_mean_fraction_global", "lc_grassland_mean_fraction")) +
        pick_numeric(., c("lc_shrubs_mean_fraction_global", "lc_shrubs_mean_fraction"))
    ),
    grow_grshrub = row_sum_candidates(., c("PctGrs2016Ws", "PctShrb2016Ws")),

    global_built = 100 * pick_numeric(., c("lc_built_mean_fraction_global", "lc_built_mean_fraction")),
    grow_built = row_sum_candidates(., c("PctUrbOp2016Ws", "PctUrbLo2016Ws", "PctUrbMd2016Ws", "PctUrbHi2016Ws")),

    global_water = 100 * pick_numeric(., c("lc_water_mean_fraction_global", "lc_water_mean_fraction")),
    grow_water = row_sum_candidates(., c("PctOw2016Ws")),

    global_wetland = 100 * pick_numeric(., c("lc_wetland_mean_fraction_global", "lc_wetland_mean_fraction")),
    grow_wetland = row_sum_candidates(., c("PctWdWet2016Ws", "PctHbWet2016Ws"))
  )

lc_specs <- tribble(
  ~concept, ~category, ~x_col, ~y_col, ~unit, ~note,
  "Forest %", "Land cover", "global_forest", "grow_forest", "%", "Copernicus trees vs NLCD forest classes",
  "Cropland %", "Land cover", "global_crop", "grow_crop", "%", "Copernicus cropland vs NLCD crop + hay",
  "Grass+shrub %", "Land cover", "global_grshrub", "grow_grshrub", "%", "Copernicus grass + shrub vs NLCD grass + shrub",
  "Developed %", "Land cover", "global_built", "grow_built", "%", "Copernicus built vs NLCD developed classes",
  "Open water %", "Land cover", "global_water", "grow_water", "%", "Copernicus water vs NLCD open water",
  "Wetland %", "Land cover", "global_wetland", "grow_wetland", "%", "Copernicus wetland vs NLCD wetland classes"
)

lc_res <- run_specs(lc, lc_specs)

nice_table(
  lc_res %>% dplyr::select(-category, -x_col, -y_col, -x_present, -y_present),
  caption = "Land cover concordance: Copernicus fractions vs aggregated NLCD 2016."
)
```

```{r landcover-plot, fig.height=6}
lc_pairs <- make_pairs(lc, lc_specs)

ggplot(lc_pairs, aes(global, grow)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(alpha = 0.5, size = 1.3, colour = "#756bb1") +
  facet_wrap(~ concept, scales = "free", ncol = 3) +
  labs(
    title = "Land cover: Copernicus global product vs NLCD 2016 GROWdb product",
    subtitle = "Watershed-aggregated %. Dashed line = 1:1.",
    x = "Global / Copernicus %",
    y = "GROWdb / NLCD %"
  )
```

# Coverage gaps and non-comparable fields

This section now checks all plausible post-join column names. It should no
longer falsely report watershed climate fields as empty when they are actually
present under an unsuffixed or suffixed name.

```{r gaps, echo=FALSE}
gap_specs <- tribble(
  ~field_label, ~candidates,
  "Watershed area", list(c("AreaSqKmWs_grow", "WsAreaSqKm_grow", "totdasqkm", "AreaSqKmWs", "WsAreaSqKm")),
  "PrecipWs", list(c("PrecipWs_grow", "PrecipWs")),
  "TmeanWs", list(c("TmeanWs_grow", "TmeanWs")),
  "TmaxWs", list(c("TmaxWs_grow", "TmaxWs")),
  "TminWs", list(c("TminWs_grow", "TminWs")),
  "CatAreaSqKm", list(c("CatAreaSqKm_grow", "CatAreaSqKm")),
  "CFS", list(c("CFS_grow", "CFS")),
  "gage_id", list(c("gage_id_grow", "gage_id")),
  "Global annual watershed precip", list(c("precipWs_annual_sum_mm_global", "precipWs_annual_sum_mm")),
  "Global annual watershed tmean", list(c("tmeanWs_annual_mean_c_global", "tmeanWs_annual_mean_c")),
  "Global annual watershed tmax", list(c("tmaxWs_annual_mean_c_global", "tmaxWs_annual_mean_c")),
  "Global annual watershed tmin", list(c("tminWs_annual_mean_c_global", "tminWs_annual_mean_c"))
)

gap_tbl <- gap_specs %>%
  mutate(info = map(candidates, ~ count_non_na(j, .x))) %>%
  unnest(info) %>%
  mutate(
    status = case_when(
      !present ~ "column absent after join",
      n_nonNA == 0 ~ "present but empty for matched sites",
      TRUE ~ "populated"
    )
  ) %>%
  select(
    `Field checked` = field_label,
    column_used,
    present,
    n_nonNA,
    status
  )

nice_table(gap_tbl, caption = "Coverage of selected fields on the matched sites.")
```

Notes on routed-around or non-comparable fields:

- **Watershed climate is compared when populated.** The script now compares
  `PrecipWs`, `TmeanWs`, `TmaxWs`, and `TminWs` to the newly added global annual
  watershed climate summaries.
- **Column suffixes are handled explicitly.** A field may appear as `PrecipWs`,
  `PrecipWs_grow`, or another resolved alias depending on whether the same name
  also exists in the global table.
- **Ecoregion / biome are not numerically comparable.** The global table uses
  WWF ecoregions/biomes/realms, while GROWdb uses Omernik levels. These are
  different classification systems.
- **Site-level soil** in the global table has no direct GROWdb counterpart, so
  only watershed soil variables are compared.
- **NPP is excluded** from all summaries, tables, figures, and conclusions.

# Interpretation and conclusions

```{r conclusion-text, echo=FALSE, results='asis'}
tier_now <- bind_rows(control_res, overlap_res, lc_res) %>%
  mutate(agreement = case_when(
    is.na(pearson_r) ~ "no data",
    pearson_r >= 0.98 ~ "near-identical",
    pearson_r >= 0.90 ~ "strong",
    pearson_r >= 0.75 ~ "moderate",
    TRUE ~ "weak/divergent"
  ))

fmt_list <- function(df, tier) {
  v <- df$concept[df$agreement == tier]
  if (!length(v)) {
    return("none")
  }
  paste(v, collapse = ", ")
}

cat(sprintf("- **Controls, near-identical:** %s.\n",
            fmt_list(tier_now %>% filter(category == "Control"), "near-identical")))
cat(sprintf("- **Near-identical overall:** %s.\n", fmt_list(tier_now, "near-identical")))
cat(sprintf("- **Strong:** %s.\n", fmt_list(tier_now, "strong")))
cat(sprintf("- **Moderate:** %s.\n", fmt_list(tier_now, "moderate")))
cat(sprintf("- **Weak / divergent:** %s.\n", fmt_list(tier_now, "weak/divergent")))
```


```{r}
sites_spatial <- site_master_table %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) 

mapview(sites_spatial, zcol = "tminWs_annual_mean_c")
```


**What this establishes.** Agreement on the positive controls supports the join
and unit handling. Strong agreement on area, stream order, climate, and soils
supports the global workflow as a reasonable analogue to the CONUS workflow for
shared concepts. Lower agreement for dam density and discharge can reflect real
methodological differences among inventories and hydrologic products rather
than a processing error.

**What to inspect first.** If a variable unexpectedly falls into the weak or
moderate tier, inspect the resolved column table first. That table shows exactly
which post-join columns were used, which avoids the earlier problem where
populated fields were mistakenly labeled empty.
