---
title: "GROWdb International Sites — Watershed Delineation and Variable Extraction"
author: "Katie Willi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = FALSE)

library(tidyverse)
library(sf)
library(terra)
library(geodata)    # WorldClim, SoilGrids
library(mapview)
library(nngeo)
library(units)

# Source helper functions (includes getXYWatersheds_intl, build_hydrobasins_adj)
list.files("src/", full.names = TRUE) %>% walk(~source(.))
```

---

# 1. Identify International Sites

All GROWdb samples not represented in `conus_watersheds.RDS` are treated as international.

```{r identify_intl}
all_sites        <- read_csv("data/date_lat_long_grow_KKA_05.05.26.csv", show_col_types = FALSE)
conus_watersheds <- readRDS("data/conus_watersheds.RDS")

intl_sites_all <- all_sites %>%
  filter(!sample_name %in% conus_watersheds$sample_name)

# Unique locations only — same deduplication logic as the CONUS workflow
intl_unique <- intl_sites_all %>%
  distinct(Latitude, Longitude) %>%
  rowid_to_column() %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE)

cat("Unique international locations to delineate:", nrow(intl_unique), "\n")
mapview(intl_unique)
```

---

# 2. Load HydroBASINS Level 12

HydroBASINS provides pre-delineated watersheds globally at 8 Pfafstetter scale levels.
Level 12 is the finest available (~250 km² mean area) and is used here for delineation.

**Download**: https://www.hydrosheds.org/products/hydrobasins
Download the level-12 shapefiles for every continent containing your sites.
File naming: `hybas_{continent}_lev12_v1c.shp` where continent codes are:
`af` (Africa), `ar` (Arctic), `as` (Asia), `au` (Australia/Oceania),
`eu` (Europe), `na` (North America), `sa` (South America), `si` (Siberia/Central Asia)

```{r load_hydrobasins}
# Update this list to match which continents your sites fall in.
# GROWdb international sites span North America, Europe, and Africa at minimum.
basin_files <- list.files("data/hydrobasins/", pattern = "hybas.*lev12.*\\.shp$",
                          full.names = TRUE, recursive = TRUE)

basins <- map(basin_files, st_read, quiet = TRUE) %>%
  bind_rows() %>%
  st_make_valid()

cat("HydroBASINS level-12 polygons loaded:", nrow(basins), "\n")
cat("Columns:", paste(names(basins), collapse = ", "), "\n")

# Build the upstream adjacency list once — passed to every delineation call
# so it is not recomputed per site.
adj <- build_hydrobasins_adj(basins)
```

---

# 3. Watershed Delineation

For each unique international site, find the containing level-12 basin then traverse
upstream via `NEXT_DOWN` relationships to collect the full contributing watershed.

No DEM download or raster processing required — this is entirely polygon-based.

```{r create_output_dir}
dir.create("data/intl_watersheds", showWarnings = FALSE)
```

```{r delineate, eval=FALSE}
intl_unique %>%
  split(seq_len(nrow(.))) %>%
  walk(~try(
    getXYWatersheds_intl(
      sf        = .x,
      basins    = basins,
      adj       = adj,
      file_path = "data/intl_watersheds/"
    )
  ))
```

Load all delineated watersheds and check for failures:

```{r load_watersheds}
intl_watersheds <- list.files("data/intl_watersheds/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x))

failed_sites <- intl_unique %>%
  filter(!rowid %in% intl_watersheds$rowid)

cat("Delineated:", nrow(intl_watersheds), "of", nrow(intl_unique), "sites\n")
cat("Failed:", nrow(failed_sites), "\n")
print(failed_sites)
```

Visual QA — review watersheds against pour points:

```{r review_watersheds, eval=FALSE}
mapview(intl_watersheds, zcol = "watershed") +
  mapview(intl_unique, col.regions = "red", cex = 3)
```

Join watersheds back to all samples sharing each unique location:

```{r join_samples}
intl_watersheds_all <- intl_sites_all %>%
  inner_join(
    intl_watersheds %>% st_drop_geometry(),
    by = c("Latitude", "Longitude")
  ) %>%
  left_join(
    intl_watersheds %>% select(rowid, geometry),
    by = "rowid"
  ) %>%
  st_as_sf()

saveRDS(intl_watersheds_all, "data/intl_watersheds.RDS")
```

---

# 4. HydroBASINS Sub-Basin Hierarchy

Since delineation is done with HydroBASINS, Pfafstetter codes are already available
from the containing level-12 basin. These are the global analogue of HUC codes.

```{r pfafstetter}
# The HYBAS_ID of the pour point's local basin was stored during delineation.
# Re-join the site basin attributes here.
site_basin_attrs <- sf::st_join(
  intl_unique %>% st_transform(st_crs(basins)),
  basins %>% select(HYBAS_ID, PFAF_ID, SUB_AREA, UP_AREA, NEXT_DOWN, MAIN_BAS),
  join = sf::st_within
) %>%
  st_drop_geometry() %>%
  # Fall back for any points missed by st_within
  group_by(rowid) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
    pfaf_2  = substr(as.character(PFAF_ID), 1, 2),
    pfaf_4  = substr(as.character(PFAF_ID), 1, 4),
    pfaf_6  = substr(as.character(PFAF_ID), 1, 6),
    pfaf_8  = substr(as.character(PFAF_ID), 1, 8),
    pfaf_10 = substr(as.character(PFAF_ID), 1, 10)
  ) %>%
  select(rowid, Latitude, Longitude, HYBAS_ID, PFAF_ID,
         pfaf_2, pfaf_4, pfaf_6, pfaf_8, pfaf_10,
         SUB_AREA, UP_AREA)
```

---

# 5. River Network Attributes (HydroRIVERS)

HydroRIVERS provides stream order, estimated mean annual discharge, and upstream
drainage area globally — the analogue of NHD flowline metadata.

**Download**: https://www.hydrosheds.org/products/hydrorivers
Same continental file structure as HydroBASINS.

```{r hydrorivers}

hydro_rivers <- st_read("data/hydrorivers/HydroRIVERS_v10.gdb/HydroRIVERS_v10.gdb")

# Nearest HydroRIVERS segment per unique site
nearest_idx <- sf::st_nearest_feature(
  intl_unique %>% st_transform(st_crs(hydro_rivers)),
  hydro_rivers
)

river_attrs <- hydro_rivers[nearest_idx, ] %>%
  st_drop_geometry() %>%
  select(
    HYRIV_ID,
    StreamOrder = ORD_STRA,   # Strahler stream order (analogue of NHD streamorde)
    FlowOrder   = ORD_FLOW,   # flow-based order
    DischargeCMS = DIS_AV_CMS, # mean annual discharge (m³/s)
    UplandSqKm  = UPLAND_SKM, # upstream drainage area (km²)
    LengthKm    = LENGTH_KM,
    MainRiv     = MAIN_RIV
  )

intl_hydro <- bind_cols(
  intl_unique %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude),
  river_attrs
)
```

---

# 6. Climate — WorldClim v2.1 (30-year normals, 1970–2000)

Replaces PRISM (CONUS-only). WorldClim provides global 30-year normals at 10 arcmin
(~18 km). Each variable is a 12-layer SpatRaster (one per month).

```{r worldclim_download, eval=FALSE}
wc_path <- "data/worldclim"
dir.create(wc_path, showWarnings = FALSE)

# Run once; geodata caches the downloaded files at wc_path
geodata::worldclim_global(var = "tavg", res = 10, path = wc_path)
geodata::worldclim_global(var = "tmax",  res = 10, path = wc_path)
geodata::worldclim_global(var = "tmin",  res = 10, path = wc_path)
geodata::worldclim_global(var = "prec",  res = 10, path = wc_path)
```

```{r worldclim_load}
wc_tmean <- terra::rast(list.files("data/worldclim", pattern = "wc2.1_10m_tavg", full.names = TRUE, recursive = TRUE))
wc_tmax  <- terra::rast(list.files("data/worldclim", pattern = "wc2.1_10m_tmax", full.names = TRUE, recursive = TRUE))
wc_tmin  <- terra::rast(list.files("data/worldclim", pattern = "wc2.1_10m_tmin", full.names = TRUE, recursive = TRUE))
wc_prec  <- terra::rast(list.files("data/worldclim", pattern = "wc2.1_10m_prec", full.names = TRUE, recursive = TRUE))

# Collapse 12 monthly layers to annual summaries
wc_tmean_ann <- mean(wc_tmean)     # mean annual temperature (°C)
wc_tmax_ann  <- max(wc_tmax)       # mean temperature of warmest month (°C)
wc_tmin_ann  <- min(wc_tmin)       # mean temperature of coldest month (°C)
wc_prec_ann  <- sum(wc_prec)       # total annual precipitation (mm)
```

### 6a. Site-level point extraction

```{r worldclim_point}
intl_pts_4326 <- intl_unique %>% st_transform(st_crs(wc_tmean_ann))

point_tmean <- terra::extract(wc_tmean_ann, intl_pts_4326, df = TRUE) %>% select(TmeanSite = 2)
point_tmax  <- terra::extract(wc_tmax_ann,  intl_pts_4326, df = TRUE) %>% select(TmaxSite  = 2)
point_tmin  <- terra::extract(wc_tmin_ann,  intl_pts_4326, df = TRUE) %>% select(TminSite  = 2)
point_prec  <- terra::extract(wc_prec_ann,  intl_pts_4326, df = TRUE) %>% select(PrecipSite = 2)

intl_points_climate <- bind_cols(
  intl_unique %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude),
  point_tmean, point_tmax, point_tmin, point_prec
)
```

### 6b. Watershed-level mean extraction

```{r worldclim_watershed}
intl_ws_4326 <- intl_watersheds %>% st_transform(st_crs(wc_tmean_ann))

ws_tmean <- terra::extract(wc_tmean_ann, intl_ws_4326, mean, na.rm = TRUE) %>% select(TmeanWs = 2)
ws_tmax  <- terra::extract(wc_tmax_ann,  intl_ws_4326, mean, na.rm = TRUE) %>% select(TmaxWs  = 2)
ws_tmin  <- terra::extract(wc_tmin_ann,  intl_ws_4326, mean, na.rm = TRUE) %>% select(TminWs  = 2)
ws_prec  <- terra::extract(wc_prec_ann,  intl_ws_4326, mean, na.rm = TRUE) %>% select(PrecipWs = 2)

intl_ws_climate <- cbind(
  intl_watersheds %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude),
  ws_tmean, ws_tmax, ws_tmin, ws_prec
)
```

---

# 7. Global Aridity Index

Already global — reuse `data/ai_et0.tif` directly (same file as CONUS workflow).

```{r aridity}
aridity <- terra::rast("data/ai_et0.tif")

point_aridity <- terra::extract(aridity, intl_pts_4326, df = TRUE) %>%
  mutate(AriditySite = ai_et0 / 10000) %>%
  select(AriditySite)

ws_aridity <- terra::extract(aridity, intl_ws_4326, mean, na.rm = TRUE) %>%
  mutate(AridityWs = ai_et0 / 10000) %>%
  select(AridityWs)
```

---

# 8. MODIS Net Primary Production (NPP)

Already global — reuse `data/npp500m.tif` directly (same file as CONUS workflow).

```{r npp}
npp <- terra::rast("data/npp500m.tif")

point_npp <- terra::extract(npp, intl_pts_4326 %>% st_transform(st_crs(npp)), df = TRUE) %>%
  select(NPPSite = 2)

ws_npp <- terra::extract(npp, intl_ws_4326 %>% st_transform(st_crs(npp)), mean, na.rm = TRUE) %>%
  select(NPPWs = 2)
```

---

# 9. Land Cover — ESA WorldCover 2021

Global 10 m land cover with 11 classes; replaces NLCD for international sites.

**Download**: https://zenodo.org/record/7254221
Files are 3° × 3° tiles; download only the tiles covering your sites, or use the
global mosaic if available. File pattern: `ESA_WorldCover_10m_2021_v200_*_Map.tif`

```{r worldcover_load}
# Load tiles — terra handles multi-file mosaics via sprc() + merge() or vrt()
wc_tiles <- list.files("data/worldcover/", pattern = "\\.tif$", full.names = TRUE,
                       recursive = TRUE)
worldcover <- terra::vrt(wc_tiles)  # virtual raster mosaic; no data copied to disk
```

```{r worldcover_extract}
# ESA WorldCover class values:
# 10 = Tree cover       20 = Shrubland      30 = Grassland     40 = Cropland
# 50 = Built-up         60 = Bare/sparse     70 = Snow/ice      80 = Permanent water
# 90 = Herbaceous wetland  95 = Mangroves   100 = Moss/lichen
lc_class_vals <- c(10, 20, 30, 40, 50, 60, 70, 80, 90, 95, 100)
lc_class_nms  <- c("PctTree2021Ws", "PctShrub2021Ws", "PctGrass2021Ws", "PctCrop2021Ws",
                   "PctBuilt2021Ws", "PctBare2021Ws", "PctIce2021Ws", "PctWater2021Ws",
                   "PctHbWet2021Ws", "PctMang2021Ws", "PctOther2021Ws")

intl_ws_wc <- intl_watersheds %>% st_transform(st_crs(worldcover))

ws_lc_raw <- terra::extract(worldcover, intl_ws_wc) %>%
  as_tibble() %>%
  rename(lc_class = 2) %>%
  group_by(ID) %>%
  mutate(n_total = n()) %>%
  group_by(ID, lc_class) %>%
  summarize(pct = n() / first(n_total) * 100, .groups = "drop") %>%
  filter(lc_class %in% lc_class_vals) %>%
  mutate(varname = lc_class_nms[match(lc_class, lc_class_vals)]) %>%
  select(ID, varname, pct) %>%
  pivot_wider(names_from = varname, values_from = pct, values_fill = 0)

intl_ws_landcover <- bind_cols(
  intl_watersheds %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude),
  ws_lc_raw %>% select(-ID)
) %>%
  mutate(WatershedAreaSqKm = as.numeric(st_area(intl_ws_4326)) / 1e6)
```

---

# 10. Ecoregions — Resolve Ecoregions 2017

Replaces North American Omernik ecoregions. Resolve provides biome and ecoregion
names globally (847 ecoregions, 14 biomes).

**Download**: https://zenodo.org/record/3551663

```{r resolve_ecoregions}
resolve <- sf::st_read("data/resolve_ecoregions/Ecoregions2017.shp", quiet = TRUE) %>%
  select(ECO_NAME, BIOME_NAME, REALM)

# Site-level: which ecoregion does the pour point fall in?
eco_site <- sf::st_join(
  intl_unique %>% st_transform(st_crs(resolve)),
  resolve,
  join = sf::st_within,
  left = TRUE
) %>%
  st_drop_geometry() %>%
  group_by(rowid) %>%
  slice(1) %>%
  ungroup() %>%
  select(rowid, Latitude, Longitude,
         EcoregionSite = ECO_NAME,
         BiomeSite     = BIOME_NAME,
         RealmSite     = REALM)

# Watershed dominant ecoregion (largest overlap area)
intl_ws_id <- intl_watersheds %>%
  st_transform(st_crs(resolve)) %>%
  mutate(ws_id = row_number())

ws_eco <- sf::st_intersection(
  intl_ws_id %>% select(ws_id),
  resolve
) %>%
  mutate(overlap_area = sf::st_area(.)) %>%
  st_drop_geometry() %>%
  group_by(ws_id) %>%
  slice_max(overlap_area, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(ws_id) %>%
  select(EcoregionWs = ECO_NAME, BiomeWs = BIOME_NAME, RealmWs = REALM)
```

---

# 11. Soil Properties — SoilGrids (via `geodata`)

Replaces StreamCat soil variables (ClayWs, SandWs, OmWs, etc.).
`geodata::soil_world()` downloads global SoilGrids rasters at 5 km resolution.

```{r soil_download, eval=FALSE}
soil_path <- "data/soilgrids"
dir.create(soil_path, showWarnings = FALSE)

# Run once; geodata caches results at soil_path
geodata::soil_world(var = "clay",  depth = 30, stat = "mean", path = soil_path)
geodata::soil_world(var = "sand",  depth = 30, stat = "mean", path = soil_path)
geodata::soil_world(var = "soc",   depth = 30, stat = "mean", path = soil_path)  # soil organic carbon
geodata::soil_world(var = "phh2o", depth = 30, stat = "mean", path = soil_path)  # pH
```

```{r soil_extract}
soil_clay <- terra::rast(list.files("data/soilgrids", pattern = "clay.*30.*mean", full.names = TRUE)[1])
soil_sand <- terra::rast(list.files("data/soilgrids", pattern = "sand.*30.*mean", full.names = TRUE)[1])
soil_soc  <- terra::rast(list.files("data/soilgrids", pattern = "soc.*30.*mean",  full.names = TRUE)[1])
soil_ph   <- terra::rast(list.files("data/soilgrids", pattern = "phh2o.*30.*mean",full.names = TRUE)[1])

intl_ws_soil_crs <- intl_watersheds %>% st_transform(st_crs(soil_clay))

ws_clay <- terra::extract(soil_clay, intl_ws_soil_crs, mean, na.rm = TRUE) %>% select(ClayWs = 2)
ws_sand <- terra::extract(soil_sand, intl_ws_soil_crs, mean, na.rm = TRUE) %>% select(SandWs = 2)
ws_soc  <- terra::extract(soil_soc,  intl_ws_soil_crs, mean, na.rm = TRUE) %>% select(SocWs  = 2)
ws_ph   <- terra::extract(soil_ph,   intl_ws_soil_crs, mean, na.rm = TRUE) %>% select(PhWs   = 2)

intl_ws_soil_df <- cbind(
  intl_watersheds %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude),
  ws_clay, ws_sand, ws_soc, ws_ph
)
```

---

# 12. Dam Data — GRanD (Global Reservoir and Dam Database)

Replaces StreamCat `DamDensWs` / `DamNrmStorWs`. GRanD v1.3 contains ~7,000 dams.

**Download**: https://www.globaldamwatch.org/grand (free, requires registration)

```{r grand_dams}
grand_dams <- sf::st_read("data/GRanD_Version_1_3/GRanD_dams_v1_3.shp", quiet = TRUE) %>%
  select(GRAND_ID, DAM_NAME, CAP_MCM, YEAR)

intl_ws_dam_crs <- intl_watersheds %>% st_transform(st_crs(grand_dams))

dams_in_ws <- sf::st_join(
  grand_dams,
  intl_ws_dam_crs %>% select(rowid),
  join = sf::st_within
) %>%
  st_drop_geometry() %>%
  filter(!is.na(rowid)) %>%
  group_by(rowid) %>%
  summarize(
    DamCountWs = n(),
    DamStorWs  = sum(CAP_MCM, na.rm = TRUE),   # total storage (million m³)
    .groups = "drop"
  )

intl_ws_dams <- intl_watersheds %>%
  st_drop_geometry() %>%
  select(rowid, Latitude, Longitude) %>%
  left_join(dams_in_ws, by = "rowid") %>%
  mutate(
    DamCountWs = replace_na(DamCountWs, 0),
    DamStorWs  = replace_na(DamStorWs, 0),
    WsAreaSqKm = as.numeric(st_area(intl_ws_4326)) / 1e6,
    DamDensWs  = DamCountWs / WsAreaSqKm
  )
```

---

# 13. Population Density — GPW v4

Replaces StreamCat `PopDen2010Ws`. GPW v4 (rev 11) at 2.5 arcmin (~5 km).

**Download**: https://sedac.ciesin.columbia.edu/data/collection/gpw-v4
Select "Population Density, v4 Revision 11" → year 2010 or 2020.

```{r population}
gpw <- terra::rast("data/gpw/gpw_v4_population_density_rev11_2010_2pt5_min.tif")

ws_popden <- terra::extract(gpw, intl_ws_4326 %>% st_transform(st_crs(gpw)),
                            mean, na.rm = TRUE) %>%
  select(PopDenWs = 2)

intl_ws_population <- cbind(
  intl_watersheds %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude),
  ws_popden
)
```

---

# 14. Wastewater Treatment Plants — HydroWASTE

Replaces StreamCat `WWTPAllDensWs`. HydroWASTE contains ~58,000 WWTPs globally.

**Download**: https://www.hydrosheds.org/products/hydrowaste (no registration required)

```{r hydrowaste}
hydrowaste <- sf::st_read("data/HydroWASTE_v10/HydroWASTE_v10.shp", quiet = TRUE) %>%
  select(WWTP_ID, COUNTRY, POP_SERVED, WASTE_DIS)

intl_ws_hw_crs <- intl_watersheds %>% st_transform(st_crs(hydrowaste))

wwtp_in_ws <- sf::st_join(
  hydrowaste,
  intl_ws_hw_crs %>% select(rowid),
  join = sf::st_within
) %>%
  st_drop_geometry() %>%
  filter(!is.na(rowid)) %>%
  group_by(rowid) %>%
  summarize(
    WWTPCountWs = n(),
    WWTPPopWs   = sum(POP_SERVED, na.rm = TRUE),
    .groups = "drop"
  )

intl_ws_wwtp <- intl_watersheds %>%
  st_drop_geometry() %>%
  select(rowid, Latitude, Longitude) %>%
  left_join(wwtp_in_ws, by = "rowid") %>%
  mutate(
    WWTPCountWs = replace_na(WWTPCountWs, 0),
    WWTPPopWs   = replace_na(WWTPPopWs, 0),
    WsAreaSqKm  = as.numeric(st_area(intl_ws_4326)) / 1e6,
    WWTPDensWs  = WWTPCountWs / WsAreaSqKm
  )
```

---

# 15. Flow Data

No global equivalent of USGS NWIS exists with a clean R API. Options by region:

| Region       | Source                        | R package / access |
|--------------|-------------------------------|-------------------|
| Canada       | WSC Hydat                     | `tidyhydat`       |
| UK           | NRFA                          | `rnrfa`           |
| Europe       | Copernicus / ECA&D            | `ecad` / manual   |
| Global (modelled) | GloFAS reanalysis v4.0   | `ecmwfr` (CDS)    |
| Global (obs) | GRDC                          | Manual request at portal.grdc.bafg.de |

```{r flow_canada, eval=FALSE}
library(tidyhydat)
tidyhydat::download_hydat()  # one-time download of full Hydat database

canada_sites <- intl_unique %>%
  filter(Latitude > 49 & Latitude < 85 & Longitude > -141 & Longitude < -52)

wsc_stations <- tidyhydat::hy_stations() %>%
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326)

nearest_wsc_idx <- sf::st_nearest_feature(canada_sites, wsc_stations)

canada_flow_sites <- canada_sites %>%
  st_drop_geometry() %>%
  mutate(STATION_NUMBER = wsc_stations$STATION_NUMBER[nearest_wsc_idx])

canada_daily <- tidyhydat::hy_daily_flows(
  station_number = unique(canada_flow_sites$STATION_NUMBER),
  start_date     = "1979-10-01",
  end_date       = "2020-09-30"
)
```

```{r flow_uk, eval=FALSE}
library(rnrfa)

uk_sites <- intl_unique %>%
  filter(Latitude > 49 & Latitude < 61 & Longitude > -8 & Longitude < 2)

nrfa_cat <- rnrfa::catalogue()
# Match sites to nearest NRFA gauge via coordinates, then:
# uk_flow <- rnrfa::gdf(id = matched_gauge_id)
```

```{r flow_glofas, eval=FALSE}
# GloFAS reanalysis v4.0: modelled daily discharge globally at 0.05° (~5 km), 1979–present.
# Requires free Copernicus CDS account: https://cds.climate.copernicus.eu
library(ecmwfr)

# Set CDS credentials once:
# wf_set_key(user = "your_uid", key = "your_api_key", service = "cds")

glofas_request <- list(
  dataset            = "cems-glofas-historical",
  system_version     = "version_4_0",
  format             = "netcdf",
  variable           = "river_discharge_in_the_last_24_hours",
  hydrological_model = "lisflood",
  product_type       = "consolidated",
  year               = as.character(1979:2020),
  month              = sprintf("%02d", 1:12),
  day                = sprintf("%02d", 1:31),
  target             = "data/glofas/glofas_discharge.nc"
)

# wf_request(user = "your_uid", request = glofas_request, transfer = TRUE, path = ".")
```

---

# 16. Final Join

Combine all variable sets into a single international dataset. Structure mirrors
`conus_watersheds.RDS` as closely as possible.

```{r final_join}
intl_site_vars <- intl_unique %>%
  st_drop_geometry() %>%
  select(rowid, Latitude, Longitude) %>%
  left_join(site_basin_attrs,    by = c("rowid", "Latitude", "Longitude")) %>%
  left_join(intl_hydro,          by = c("rowid", "Latitude", "Longitude")) %>%
  left_join(intl_points_climate, by = c("rowid", "Latitude", "Longitude")) %>%
  left_join(
    bind_cols(
      intl_unique %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude),
      point_aridity
    ),
    by = c("rowid", "Latitude", "Longitude")
  ) %>%
  left_join(
    bind_cols(
      intl_unique %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude),
      point_npp
    ),
    by = c("rowid", "Latitude", "Longitude")
  ) %>%
  left_join(eco_site, by = c("rowid", "Latitude", "Longitude"))

intl_ws_vars <- intl_watersheds %>%
  st_drop_geometry() %>%
  select(rowid, Latitude, Longitude) %>%
  left_join(intl_ws_climate,    by = c("rowid", "Latitude", "Longitude")) %>%
  left_join(
    cbind(intl_watersheds %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude), ws_aridity),
    by = c("rowid", "Latitude", "Longitude")
  ) %>%
  left_join(
    cbind(intl_watersheds %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude), ws_npp),
    by = c("rowid", "Latitude", "Longitude")
  ) %>%
  left_join(intl_ws_landcover, by = c("rowid", "Latitude", "Longitude")) %>%
  left_join(
    bind_cols(
      intl_watersheds %>% st_drop_geometry() %>% select(rowid, Latitude, Longitude),
      ws_eco
    ),
    by = c("rowid", "Latitude", "Longitude")
  ) %>%
  left_join(intl_ws_soil_df,   by = c("rowid", "Latitude", "Longitude")) %>%
  left_join(intl_ws_dams,       by = c("rowid", "Latitude", "Longitude")) %>%
  left_join(intl_ws_population, by = c("rowid", "Latitude", "Longitude")) %>%
  left_join(intl_ws_wwtp,       by = c("rowid", "Latitude", "Longitude"))

intl_complete <- intl_watersheds_all %>%
  st_drop_geometry() %>%
  left_join(intl_site_vars, by = c("rowid", "Latitude", "Longitude")) %>%
  left_join(intl_ws_vars,   by = c("rowid", "Latitude", "Longitude")) %>%
  left_join(
    intl_watersheds %>% select(rowid, geometry),
    by = "rowid"
  ) %>%
  st_as_sf() %>%
  st_make_valid()

saveRDS(intl_complete, "data/intl_complete.RDS")
write_csv(intl_complete %>% st_drop_geometry(), "data/intl_complete.csv")

cat("Final international dataset:", nrow(intl_complete), "samples\n")
cat("Variables:", ncol(intl_complete), "\n")
```

---

# Data Download Checklist

```{r download_checklist}
datasets <- tribble(
  ~dataset,              ~path,                                                           ~notes,
  "HydroBASINS L12",    "data/hydrobasins/",                                             "hydrosheds.org — 1 shp per continent",
  "HydroRIVERS",        "data/hydrorivers/",                                             "hydrosheds.org — 1 shp per continent",
  "WorldClim tavg",     "data/worldclim/wc2.1_10m_tavg_01.tif",                         "auto via geodata::worldclim_global()",
  "ESA WorldCover",     "data/worldcover/",                                              "zenodo.org/record/7254221 — tiled",
  "Resolve Ecoregions", "data/resolve_ecoregions/Ecoregions2017.shp",                   "zenodo.org/record/3551663",
  "SoilGrids clay",     "data/soilgrids/",                                               "auto via geodata::soil_world()",
  "GRanD dams",         "data/GRanD_Version_1_3/GRanD_dams_v1_3.shp",                  "globaldamwatch.org/grand",
  "GPW pop density",    "data/gpw/gpw_v4_population_density_rev11_2010_2pt5_min.tif",   "sedac.ciesin.columbia.edu",
  "HydroWASTE",         "data/HydroWASTE_v10/HydroWASTE_v10.shp",                      "hydrosheds.org/products/hydrowaste",
  "Global Aridity",     "data/ai_et0.tif",                                               "already present",
  "MODIS NPP",          "data/npp500m.tif",                                              "already present"
)

datasets %>%
  mutate(
    path_exists = case_when(
      grepl("/$", path) ~ dir.exists(path) & length(list.files(path)) > 0,
      TRUE              ~ file.exists(path)
    )
  ) %>%
  select(dataset, path_exists, notes) %>%
  rename(present = path_exists) %>%
  print(n = Inf)
```
