---
title: "GROWdb International Sites — HydroBASINS / HydroRIVERS Watershed Workflow"
author: "Katie Willi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = FALSE)
```

```{r libraries}
library(tidyverse)
library(sf)
library(terra)
library(elevatr)
library(lwgeom)
library(nngeo)
library(mapview)
library(leaflet)
library(units)
library(geodata)

list.files("src", full.names = TRUE) %>%
  walk(~source(.))
```

# 1. Setup

```{r paths}
site_csv <- "data/date_lat_long_grow_KKA_05.05.26.csv"
conus_watershed_rds <- "data/conus_watersheds.RDS"

hydrobasins_dir <- "data/hydrobasins"
hydrorivers_dir <- "data/hydrorivers"
```

# 2. Identify international sites

All GROWdb samples not represented in the CONUS watershed object are routed through this international workflow.

```{r identify-international-sites}
all_sites <- readr::read_csv(site_csv, show_col_types = FALSE)
conus_watersheds <- readRDS(conus_watershed_rds)

# If conus_watersheds has sample_name, use that directly. If not, fall back to
# filtering by coordinates already present in the CONUS watershed table.
if ("sample_name" %in% names(conus_watersheds)) {
  intl_sites_all <- all_sites %>%
    filter(!sample_name %in% conus_watersheds$sample_name)
} else {
  conus_coords <- conus_watersheds %>%
    sf::st_drop_geometry() %>%
    distinct(Latitude, Longitude) %>%
    mutate(coord_key = paste(round(Latitude, 6), round(Longitude, 6)))
  
  intl_sites_all <- all_sites %>%
    mutate(coord_key = paste(round(Latitude, 6), round(Longitude, 6))) %>%
    filter(!coord_key %in% conus_coords$coord_key) %>%
    select(-coord_key)
}

# Unique pour-point locations only. All samples sharing the same coordinate are
# joined back after watershed delineation/review.
intl_unique <- intl_sites_all %>%
  distinct(Latitude, Longitude) %>%
  arrange(desc(Latitude), Longitude) %>%
  rowid_to_column("rowid") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE)

cat("International samples:", nrow(intl_sites_all), "\n")
cat("Unique international locations:", nrow(intl_unique), "\n")

mapview(intl_unique, col.regions = "red", cex = 4)
```

# 3. Load HydroBASINS level 12

HydroBASINS is used to get the full upstream watershed for each international pour point.
Level 12 is the finest HydroBASINS level and is the closest global analogue to a local watershed unit.

Download HydroBASINS from HydroSHEDS and unzip all continent-level files needed for the international sites.
Expected shapefile names usually look like `hybas_na_lev12_v1c.shp`, `hybas_sa_lev12_v1c.shp`, etc.

```{r load-hydrobasins}
basin_files <- list.files(hydrobasins_dir,
                          pattern = "hybas.*lev12.*\\.shp$",
                          full.names = TRUE,
                          recursive = TRUE)

#stopifnot(length(basin_files) > 0)

basins <- purrr::map(basin_files, ~sf::st_read(.x, quiet = TRUE)) %>%
  dplyr::bind_rows() %>%
  sf::st_transform(4326) %>%
  sf::st_make_valid()

# Build the upstream-child lookup once. This makes repeated watershed tracing much
# faster because getXYWatersheds_intl() does not need to rebuild the graph per site.
# adj <- build_hydrobasins_adj(basins)
```

# 4. Load HydroRIVERS

HydroRIVERS is used in two ways:

1. As a review-map overlay.
2. To split the local HydroBASINS polygon at the pour point and remove the downstream area below the sample location.

Download HydroRIVERS from HydroSHEDS and unzip all continent-level files needed for the international sites.
Expected shapefile names usually look like `HydroRIVERS_v10_na.shp`, `HydroRIVERS_v10_sa.shp`, etc.

```{r load-hydrorivers}
hydro_rivers <- st_read("data/hydrorivers/HydroRIVERS_v10.gdb/HydroRIVERS_v10.gdb")
```

Keep only HydroRIVERS segments near the international sites for faster maps and split calls.

```{r trim-hydrorivers}
intl_sites_buf <- intl_unique %>%
  sf::st_transform(3857) %>%
  sf::st_buffer(50000) %>%
  sf::st_transform(4326)

hydro_rivers_review <- hydro_rivers[intl_sites_buf, ]
```


```{r}
flowline_selections <- select_flowline_global(
  watersheds    = basins %>%
    inner_join(intl_unique %>% st_drop_geometry(), by = "rowid"),
  sites         = intl_unique,
  rivers        = hydro_rivers_review,
  river_buffer_m = 10000,
  id_col        = "rowid"
)

saveRDS(flowline_selections, "data/flowline_selections.RDS")
```


```{r}
# Step 1 — select flowlines interactively (once for all sites)
flowline_selections <- select_flowline_intl(
  watersheds = intl_watersheds_full %>%
    inner_join(intl_unique %>% st_drop_geometry(), by = "rowid"),
  sites  = intl_unique,
  rivers = hydro_rivers_review,
  id_col = "rowid"
)

```

# 8. Join watersheds back to all samples

The delineation/review is done once per unique coordinate. This section joins the reviewed watershed geometry back to every sample that shares the same latitude/longitude.

```{r join-back-to-samples}
intl_watersheds_all <- intl_sites_all %>%
  inner_join(
    intl_watersheds_reviewed %>%
      sf::st_drop_geometry(),
    by = c("Latitude", "Longitude")
  ) %>%
  left_join(
    intl_watersheds_reviewed %>%
      select(rowid, geometry),
    by = "rowid"
  ) %>%
  sf::st_as_sf() %>%
  sf::st_make_valid()

saveRDS(intl_watersheds_all, final_out)

cat("International samples with reviewed watersheds:", nrow(intl_watersheds_all), "\n")
```

# 9. HydroBASINS hierarchy attributes

Pfafstetter codes provide a global HydroBASINS hierarchy analogous to HUC-style watershed nesting.

```{r hydrobasins-hierarchy}
site_basin_attrs <- sf::st_join(
  intl_unique %>% sf::st_transform(sf::st_crs(basins)),
  basins %>%
    select(HYBAS_ID, PFAF_ID, SUB_AREA, UP_AREA, NEXT_DOWN, MAIN_BAS),
  join = sf::st_within
) %>%
  sf::st_drop_geometry() %>%
  group_by(rowid) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
    pfaf_2  = substr(as.character(PFAF_ID), 1, 2),
    pfaf_4  = substr(as.character(PFAF_ID), 1, 4),
    pfaf_6  = substr(as.character(PFAF_ID), 1, 6),
    pfaf_8  = substr(as.character(PFAF_ID), 1, 8),
    pfaf_10 = substr(as.character(PFAF_ID), 1, 10)
  ) %>%
  select(rowid, Latitude, Longitude, HYBAS_ID, PFAF_ID,
         pfaf_2, pfaf_4, pfaf_6, pfaf_8, pfaf_10,
         SUB_AREA, UP_AREA, NEXT_DOWN, MAIN_BAS)

site_basin_attrs
```

# 10. HydroRIVERS attributes

Nearest HydroRIVERS attributes provide a global analogue to flowline metadata.

```{r hydrorivers-attributes}
nearest_idx <- sf::st_nearest_feature(
  intl_unique %>% sf::st_transform(sf::st_crs(hydro_rivers)),
  hydro_rivers
)

river_attrs <- hydro_rivers[nearest_idx, ] %>%
  sf::st_drop_geometry()

# Keep these only if present because HydroRIVERS versions/download formats can
# differ slightly in column names.
river_keep <- c(
  "HYRIV_ID",
  "NEXT_DOWN",
  "MAIN_RIV",
  "LENGTH_KM",
  "DIST_DN_KM",
  "DIST_UP_KM",
  "CATCH_SKM",
  "UPLAND_SKM",
  "DIS_AV_CMS",
  "ORD_STRA",
  "ORD_CLAS",
  "ORD_FLOW",
  "HYBAS_L12"
)

river_attrs <- river_attrs %>%
  select(any_of(river_keep))

intl_hydro_attrs <- bind_cols(
  intl_unique %>%
    sf::st_drop_geometry() %>%
    select(rowid, Latitude, Longitude),
  river_attrs
)

intl_hydro_attrs
```

# 11. Optional: WorldClim variables

This section is unchanged conceptually from the original international workflow. WorldClim replaces PRISM because PRISM is CONUS-only.

```{r worldclim-download, eval=FALSE}
wc_path <- "data/worldclim"
dir.create(wc_path, showWarnings = FALSE, recursive = TRUE)

geodata::worldclim_global(var = "tavg", res = 10, path = wc_path)
geodata::worldclim_global(var = "tmax", res = 10, path = wc_path)
geodata::worldclim_global(var = "tmin", res = 10, path = wc_path)
geodata::worldclim_global(var = "prec", res = 10, path = wc_path)
```

```{r worldclim-load-extract, eval=FALSE}
wc_tmean <- terra::rast(list.files("data/worldclim", pattern = "wc2.1_10m_tavg", full.names = TRUE, recursive = TRUE))
wc_tmax  <- terra::rast(list.files("data/worldclim", pattern = "wc2.1_10m_tmax",  full.names = TRUE, recursive = TRUE))
wc_tmin  <- terra::rast(list.files("data/worldclim", pattern = "wc2.1_10m_tmin",  full.names = TRUE, recursive = TRUE))
wc_prec  <- terra::rast(list.files("data/worldclim", pattern = "wc2.1_10m_prec",  full.names = TRUE, recursive = TRUE))

wc_tmean_ann <- mean(wc_tmean)
wc_tmax_ann  <- max(wc_tmax)
wc_tmin_ann  <- min(wc_tmin)
wc_prec_ann  <- sum(wc_prec)

intl_vect <- terra::vect(intl_watersheds_reviewed)

climate_attrs <- tibble(
  rowid = intl_watersheds_reviewed$rowid,
  tmean_ann_c = terra::extract(wc_tmean_ann, intl_vect, fun = mean, na.rm = TRUE)[, 2],
  tmax_warmest_month_c = terra::extract(wc_tmax_ann, intl_vect, fun = mean, na.rm = TRUE)[, 2],
  tmin_coldest_month_c = terra::extract(wc_tmin_ann, intl_vect, fun = mean, na.rm = TRUE)[, 2],
  precip_ann_mm = terra::extract(wc_prec_ann, intl_vect, fun = mean, na.rm = TRUE)[, 2]
)

climate_attrs
```

# 12. Final reviewed watershed object

```{r final-summary}
cat("Unique reviewed/refined international watersheds:", nrow(intl_watersheds_reviewed), "\n")
cat("All international samples joined to watersheds:", nrow(intl_watersheds_all), "\n")

intl_watersheds_reviewed %>%
  sf::st_drop_geometry() %>%
  select(rowid, Latitude, Longitude, HYBAS_ID, watershed,
         any_of(c("split_status", "split_method", "snap_dist_m", "review"))) %>%
  arrange(rowid)
```
