---
title: "GROWdb_2026_Update"
author: "Katie Willi"
date: "2026-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(mapview)
library(nngeo)
library(nhdplusTools)
library(sf)

list.files("src/", full.names = TRUE) %>%
  walk(~source(.))
```

Pull in all sites in GROW 2026:

```{r}
sites_2026 <- read_csv("data/lat_long_grow_KKA_03.25.26.csv") %>%
  #filter(grepl("_sed_|_bw_", sample_name, ignore.case = TRUE)) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  .[tigris::states() %>% filter(!STUSPS %in% c("HI", "PR", "AK")) %>% st_buffer(10000),]

international_sites <- sites_2026 %>%
  filter(sample_name %in% c("columbia_2019_sw_WHONDRS-S19S_0036", "columbia_2019_sw_WHONDRS-S19S_0074",
                            "columbiariver1_2019_sw_WHONDRS-S19S_0075", "columbiariver2_2019_sw_WHONDRS-S19S_0076",
                            "NASQAN2016_178", "NASQAN2016_179", "NASQAN2014_141", "NASQAN2014_144", "NASQAN2015_031", "NASQAN2016_164")) %>%
  distinct(Latitude, Longitude)

# small_sites <- sites_2026 %>%
#   filter(sample_name %in% c("providencecreek_2019_sw_WHONDRS-S19S_0007", "redbuttecreek_2019_sw_WHONDRS-S19S_0029",
#     "redbuttecreek_2019_sw_WHONDRS-S19S_0086", "watershed3_2019_sw_WHONDRS-S19S_0084",
#     "sawmillbrook_2019_sw_WHONDRS-S19S_0072", "muddycreek_2019_sw_WHONDRS-S19S_0082")) %>%
#   distinct(Latitude, Longitude)
# 
# wrong_comid <- sites_2026 %>%
#   filter(sample_name %in% c("NASQAN2011_429", "NASQAN2011_433", "NASQAN2011_435", "teakettle2_2019_sw_WHONDRS-S19S_0032", "farmington_2019_sw_WHONDRS-S19S_0053", "NASQAN2011_188", "NASQAN2011_194","NASQAN2011_198", "southforkpolouse_2019_sw_WHONDRS-S19S_0055")) %>%
#   distinct(Latitude, Longitude)
```

By site name, these are the ones missing:
```{r}
sites_2024 <- readRDS("data/archive/GROWdb_CONTUS.RDS") 

missing <- sites_2026 %>%
  anti_join(., sites_2024, by = c("sample_name" = "SampleName"))

# These are the ones that were DROPPED from 2026:
missing_2024 <- sites_2024 %>%
  anti_join(., sites_2026, by = c("SampleName" = "sample_name")) %>%
  pull(SampleName)
```


Grab pre-developed watersheds

```{r}
# remove sites at the same location for watershed delineation
# to reduce processing time, and those found to be international
new_dupes_removed <- sites_2026 %>%
  select(Longitude, Latitude) %>%
  distinct(Longitude, Latitude, .keep_all = TRUE) %>%
  filter(!Longitude %in% international_sites$Longitude & !Latitude %in% international_sites$Latitude) %>%
  rowid_to_column() %>% 
  st_as_sf()
```

Delineate site watersheds: 

```{r}
new_dupes_removed %>%
  split(seq_len(nrow(.))) %>%
  walk(~try(getXYWatersheds(sf = .x, snap = TRUE, file_path = "data/grow_watersheds_all/")))
```

Pull in the delineated watersheds:

```{r}
watersheds <- list.files("data/grow_watersheds_all/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) 
```

Retry sites whose watersheds didn't delineate (errored out):

```{r}
new_dupes_removed %>%
  filter(!rowid %in% watersheds$rowid)  %>%
  split(seq_len(nrow(.))) %>%
  walk(~try(getXYWatersheds(sf = .x, snap = TRUE, file_path = "data/grow_watersheds_all/")))

watersheds <- list.files("data/grow_watersheds_all/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) 

broken_sites <- new_dupes_removed %>%
 filter(!rowid %in% watersheds$rowid) 
```


Review the delineated watersheds, flag watersheds that weren't delineated correctly:

```{r}
flowlines <- get_nhdplus(comid = watersheds %>% filter(!is.na(comid)) %>% pull(comid))

review <- review_watersheds(watersheds = watersheds, flowlines = flowlines, sites = new_dupes_removed)

saveRDS(review, "data/review_all.RDS")
```

Subset watersheds as either good, bad "small" (site too small for NHD), bad "other" (some other reason the watershed isn't delineated correctly)

```{r}
review <- readRDS("data/review_all.RDS")

good <- review %>%
  filter(review == "Y")

small <- review %>% 
  filter(review == "S")

weird <- review %>%
  filter(review == "B")
```

Retry for bad small ones with snap = FALSE

```{r}
new_dupes_removed %>%
  filter(rowid %in% small$rowid) %>%
  split(seq_len(nrow(.))) %>%
  walk(~try(getXYWatersheds(sf = .x, snap = FALSE)))

small_review <- list.files("data/grow_watersheds_all/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) %>%
  filter(rowid %in% small$rowid) %>%
  review_watersheds(watersheds = ., sites = new_dupes_removed, flowlines = flowlines)
```


# Put all watersheds together:

```{r}
previous_combined <- new_with_normal_delineations %>%
  bind_rows(new_with_small_delineations %>% st_as_sf() %>% st_transform(4326)) %>%
  st_as_sf() %>%
  left_join(., archive_ws %>% st_drop_geometry() %>% select(sample_name, old_comid = comid), by = c("sample_name")) %>%
  left_join(., archive_ws %>% st_drop_geometry() %>% select(Latitude, Longitude, old_comid = comid), by = c("Latitude", "Longitude")) %>%
  mutate(comid = if_else(is.na(old_comid.x), old_comid.y, old_comid.x)) %>%
  select(-c(old_comid.x, old_comid.y)) %>%
  distinct()

new_combined <- list.files("data/grow_watersheds/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) %>%
  inner_join(sites_2026 %>% st_drop_geometry(), by = c("Longitude", "Latitude"))

all_watersheds <- previous_combined %>%
  bind_rows(new_combined %>% st_transform(4326)) 

saveRDS(all_watersheds, "data/all_watersheds.RDS")

# are any sites missing that we needed to delineate?

test <- sites_2026 %>%
  anti_join(all_watersheds %>% st_drop_geometry(), by = "sample_name") 


nrow(test) != 0
```

