---
title: "GROWdb_Update"
author: "Katie Willi"
date: "2026-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(mapview)
library(nngeo)
library(nhdplusTools)
library(sf)
  library(leaflet)

list.files("src/", full.names = TRUE) %>%
  walk(~source(.))
```

### Review GROW

Pull in all sites in updated GROW:
```{r}
conus_sites_pre <- read_csv("data/lat_long_grow_KKA_03.25.26.csv") %>%
  mutate(Longitude = ifelse(sample_name == "marshallgulch_2019_sw_WHONDRS-S19S_0035", -110.75784, Longitude),
         Latitude = ifelse(sample_name == "marshallgulch_2019_sw_WHONDRS-S19S_0035", 32.42735, Latitude)) %>%
  #filter(grepl("_sed_|_bw_", sample_name, ignore.case = TRUE)) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  .[tigris::states() %>% filter(!STUSPS %in% c("HI", "PR", "AK")) %>% st_buffer(10000),]

conus_sites_update <- read_csv("data/date_lat_long_grow_KKA_05.05.26.csv") %>%
  mutate(Longitude = ifelse(sample_name == "marshallgulch_2019_sw_WHONDRS-S19S_0035", -110.75784, Longitude),
         Latitude = ifelse(sample_name == "marshallgulch_2019_sw_WHONDRS-S19S_0035", 32.42735, Latitude)) %>%
  #filter(grepl("_sed_|_bw_", sample_name, ignore.case = TRUE)) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  .[tigris::states() %>% filter(!STUSPS %in% c("HI", "PR", "AK")) %>% st_buffer(10000),]

# Find rows where coordinates changed between versions
coords_changed <- conus_sites_update %>%
  st_drop_geometry() %>%
  select(sample_name, Latitude_update = Latitude, Longitude_update = Longitude) %>%
  left_join(
    conus_sites_pre %>%
      st_drop_geometry() %>%
      select(sample_name, Latitude_pre = Latitude, Longitude_pre = Longitude),
    by = "sample_name") %>%
  filter(
    !is.na(Latitude_pre) &
    (
      Latitude_update != Latitude_pre |
      Longitude_update != Longitude_pre
    )
  )

# looks good!
coords_changed

```

Which sites are new, are any sites dropped?
```{r}
sites_previous <- read_csv("data/archive/GROWdb_with_vars_20230519.csv") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  .[tigris::states() %>% filter(!STUSPS %in% c("HI", "PR", "AK")) %>% st_buffer(10000),] %>%
  st_drop_geometry()

# These are the ones that are NEW in latest update:
new_update <- conus_sites_update %>%
  anti_join(., sites_previous, by = c("sample_name" = "SampleName"))

new_previous_coords_dif <- conus_sites_update %>%
  filter(sample_name %in% sites_previous$SampleName) %>%
  anti_join(., sites_previous, by = c("Latitude", "Longitude"))

old_points <- read_csv("data/archive/GROWdb_with_vars_20230519.csv") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  filter(SampleName %in% new_previous_coords_dif$sample_name) %>%
  select(SampleName, Longitude, Latitude)

# new_previous_coords_dif_reviewed <- review_point_pairs(new_points = new_previous_coords_dif,
#                                                        old_points = old_points)

#... Marshall Gulch changed substantially, and it looks like previous coordinates are more appropriate. Updated. 

# These are the ones that were DROPPED from previous GROW:
missing_previous <- sites_previous %>%
  anti_join(., conus_sites_update, by = c("SampleName" = "sample_name")) 

```

# Delineate all GROW site watersheds:
```{r}
international_sites <- conus_sites_update %>%
  filter(sample_name %in% c("columbia_2019_sw_WHONDRS-S19S_0036", "columbia_2019_sw_WHONDRS-S19S_0074",
                            "columbiariver1_2019_sw_WHONDRS-S19S_0075", "columbiariver2_2019_sw_WHONDRS-S19S_0076",
                            "NASQAN2016_178", "NASQAN2016_179", "NASQAN2014_141", "NASQAN2014_144", "NASQAN2015_031",
                            "NASQAN2016_164")) %>%
  distinct(Latitude, Longitude)

# remove sites at the same location for watershed delineation
# to reduce processing time
new_dupes_removed <- conus_sites_update %>%
  select(Longitude, Latitude) %>%
  distinct(Longitude, Latitude, .keep_all = TRUE) %>%
  # and those found to be international
  filter(!Longitude %in% international_sites$Longitude & !Latitude %in% international_sites$Latitude) %>%
  rowid_to_column() %>% 
  st_as_sf()
```

Delineate site watersheds: 
```{r eval = FALSE}
new_dupes_removed %>%
  split(seq_len(nrow(.))) %>%
  walk(~try(getXYWatersheds(sf = .x, snap = TRUE, file_path = "data/grow_watersheds_all/")))
```

Pull in the delineated watersheds:
```{r}
watersheds <- list.files("data/grow_watersheds_all/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) #%>%
  #filter(Latitude %in% coords_changed$Latitude_update & Longitude %in% coords_changed$Longitude_update)
```

Retry sites whose watersheds didn't delineate (if any errored out):

```{r, eval = FALSE}
new_dupes_removed %>%
  filter(!rowid %in% watersheds$rowid)  %>%
  split(seq_len(nrow(.))) %>%
  walk(~try(getXYWatersheds(sf = .x, snap = TRUE, file_path = "data/grow_watersheds_all/")))

watersheds <- list.files("data/grow_watersheds_all/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) 

broken_sites <- new_dupes_removed %>%
  filter(!rowid %in% watersheds$rowid) 
```

Review the delineated watersheds, flag watersheds that weren't delineated correctly:

```{r, eval = FALSE}
flowlines <- get_nhdplus(comid = watersheds %>% filter(!is.na(comid)) %>% pull(comid))

review <- review_watersheds(watersheds = watersheds, flowlines = flowlines, sites = new_dupes_removed)

saveRDS(review, "data/review_all.RDS")
```

Subset watersheds as either good, bad "small" (site too small for NHD), bad "other" (some other reason the watershed isn't delineated correctly)

```{r, eval = FALSE}
# loop - when batched together, does not return all flowlines 
comids <- watersheds %>% filter(!is.na(comid)) %>% .$comid %>% unique()

flowlines <- vector("list", length = length(comids))

for(i in 1:length(comids)){
  
  flowlines[[i]] <- get_nhdplus(comid = comids[i]) %>% select(comid)
  
}

flowlines <- bind_rows(flowlines)

review <- readRDS("data/review_all.RDS")

good <- review %>%
  filter(review == "Y")

# remaining small are: rowid 44, 51, 62
small <- review %>% 
  filter(review == "S")

weird <- review %>%
  filter(review == "B")
```

... still several that are strange

Review the "bizarre" ones, one at a time:
```{r}
# mapview(new_dupes_removed[44,]) # 
# mapview(new_dupes_removed[51,]) # 
# mapview(new_dupes_removed[62,]) # 
```

## Bind all watersheds together
```{r}
conus_watersheds <- conus_sites_update %>%
  st_drop_geometry() %>%
  inner_join(., list.files("data/grow_watersheds_all/", full.names = TRUE) %>%
               map_dfr(~readRDS(.x)), by = c("Latitude", "Longitude")) %>%
  mutate(update = if_else(sample_name %in% new_update$sample_name, "new", "previous"))

saveRDS(conus_watersheds, "data/conus_watersheds.RDS")

outside_conus <- read_csv("data/date_lat_long_grow_KKA_05.05.26.csv") %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  filter(!sample_name %in% conus_watersheds$sample_name)

mapview(outside_conus) #... looks correct!
```

## Add back in the sites that were dropped ?

```{r}
international_sites <- conus_sites_update %>%
  filter(sample_name %in% c("columbia_2019_sw_WHONDRS-S19S_0036", "columbia_2019_sw_WHONDRS-S19S_0074",
                            "columbiariver1_2019_sw_WHONDRS-S19S_0075", "columbiariver2_2019_sw_WHONDRS-S19S_0076",
                            "NASQAN2016_178", "NASQAN2016_179", "NASQAN2014_141", "NASQAN2014_144", "NASQAN2015_031",
                            "NASQAN2016_164")) %>%
  distinct(Latitude, Longitude)

# remove sites at the same location for watershed delineation
# to reduce processing time
missing_dupes_removed <- missing_previous %>%
  select(Longitude, Latitude) %>%
  distinct(Longitude, Latitude, .keep_all = TRUE) %>%
  # and those found to be international
  filter(!Longitude %in% international_sites$Longitude & !Latitude %in% international_sites$Latitude) %>%
  rowid_to_column() %>%  
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269, remove = FALSE) %>%
  .[tigris::states() %>% filter(!STUSPS %in% c("HI", "PR", "AK")) %>% st_buffer(10000),]

missing_dupes_removed %>%
  split(seq_len(nrow(.))) %>%
  walk(~try(getXYWatersheds(sf = .x, snap = TRUE, file_path = "data/grow_watersheds_missing/")))

watersheds_missing <- list.files("data/grow_watersheds_missing/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) 

flowlines <- get_nhdplus(comid = watersheds_missing %>% filter(!is.na(comid)) %>% pull(comid))

review <- review_watersheds(watersheds = watersheds_missing, flowlines = flowlines, sites = missing_dupes_removed)

saveRDS(review, "data/review_missing.RDS")
```

```{r}
conus_watersheds <- list.files("data/grow_watersheds_all/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) %>%
  inner_join(., conus_sites_update %>% st_drop_geometry(), by = c("Latitude", "Longitude")) %>%
  mutate(update = if_else(sample_name %in% new_update$sample_name, "new", "previous")) %>%
  st_make_valid()

conus_missing_watersheds <- list.files("data/grow_watersheds_missing/", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) %>%
  inner_join(., missing_previous %>% st_drop_geometry(), by = c("Latitude", "Longitude")) %>%
  mutate(update = "missing") %>%
  select(sample_name = SampleName, Collection_Date = Date, comid = comid.y, rowid, Latitude, Longitude, watershed, update) %>%
  st_make_valid()

conus_complete <- conus_watersheds %>%
  bind_rows(conus_missing_watersheds)  %>%
  select(-rowid) %>%
  distinct() %>%
  st_make_valid()

saveRDS(conus_complete, "data/conus_watersheds.RDS")
```

